### 2019-11-01 notes-automata-logic.tex
.

\documentclass{article}
\usepackage{graphicx,xcolor}
\usepackage{algorithmicx,algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amssymb,amsmath}

\newcommand{\jana}[1]{{\color{magenta} #1}}

\newcommand{\Real}{\mathbb{R}}
\newcommand{\obs}{\mathrm{obs}}
\newcommand{\free}{\mathrm{free}}
\newcommand{\AP}{\Sigma}

\begin{document}

\subsection{Problem Setup}

Let $X$ be a bounded open subset of $\Real^n$ and let $X_\obs \subset X$ be the obstacle region. $X_\free = X \setminus X_\obs$ is called the obstacle-free space.

A few more things to define in the standard way:
\begin{itemize}
\item A path
\item A collision-free path
\item A ball
\end{itemize}

Consider a labeling $L: X \rightarrow 2^\AP$ \jana{TODO formalize and assume the labeling induces a partitioning into a finite number of regions}.

Consider a spec given as an \jana{LTL/sc-LTL?} formula over $\AP$.

The problem is to find \jana{a path/an optimal path?} that satisifes the spec. The problem can be solved using straightforward extension to existing algorithms. \jana{Our contribution is a more efficient procedure that ???}

\subsection{Solution/Contribution draft}

We get a simulation, we project it to the automaton, see what went wrong (is violated or not satisfied yet), in the next try we aim at spotting where the wrong item can be fixed by a different decision.

\begin{itemize}
	\item Learning a small model from the traces, solving that one, lifting the strategy to the original model.
	\item Looking at what remained to be fullfiled and where in the learnt model this is possible.
	\item For unbounded properties, we have to do this hiearchically and check the progress for subformulas. We can do that, using the labelling from Rabinizer.
\end{itemize}

\begin{algorithm}
	\caption{Basic algorithm}\label{alg:general}
	\begin{algorithmic}[1]
		\State build trivial automaton
		\Repeat\Comment{Start new simulation}
		\Repeat\Comment{Next step in simulation}
		\State \Call{ask}{} automaton for advice which Ap to satisfy next
		\State sample a bit to get that one satisfied
		\Until{sampled all the way to satisfaction of the property or too long}
		\State \Call{learn}{} the automaton using also the new sampled paths
		\Until{sampled all the way to satisfaction of the property (fast enough)}

		\Procedure{ask}{}
		\State consider product of the learnt (system) automaton (DFA) and of the property automaton ($\omega$-automaton)
		\State for DFA use shortest path to target
		\State for DBA too
		\State for general ones we might need to non-deterministically choose: which Rabin pair of the DRA? which direction in the NBA?
		\EndProcedure
		\Procedure{learn}{}
		\State automata learning or even probabilistic automata/MDP learning
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\paragraph{Example 1}
Formulae beyond safety: things like $\varphi=GF(a\wedge Xb)\wedge FG(c\wedge dUe)$ do not even have a deterministic B\"uchi automaton, not to mention a DFA, so how to measure progress is not clear form classical approaches. 
(In DFA one can measure the distance to the accepting state, extensible for B\"uchi - not sure this is done, beyond B\"uchi hard.)
Rabinizer tracks the evolution of both parenthesised expressions and after an $a$ this advises us to go for $b$ althought requirement to be satisfied is still exactly $\varphi$.

\begin{figure}[h]
	\includegraphics[scale=0.4]{auto}\caption{Models for the second automata example}
\end{figure}

\paragraph{Example 2}
Write $\alpha\to\beta$ for $\alpha\wedge F\beta$
Within safety, but taking into account the model: $\psi=F(a\to b\to c)\vee F(d\to e)$.
Classical technique (get closer to the accepting state) prefers $d$ 9get to distance 1) over $a$ (distance still 2). 
However, if we see that nothing reasonable comes after $d$, in particular not even an $a$ then we want to try out the first option, too.
To this end we can build a model from learnt traces to see that taking $d$ does not allow for a winning trace.

Similarly, for $\wedge$ instead of $\vee$, getting $d$ is bad at first (no $a$ after $d$, better to stay on spot than to shorten the distance badly).
Seen orders of subsets of $\{a,b,c,d,e\}$ induce a model.
We can make a product with the automaton and then decide there, where to go.

Jana's notes from the call: 
\begin{itemize}
\item State space; partitioned -- we don't know the labels of regions apriori, but if we have a concrete sample, I am able to test whether an atomic proposition holds or not; make sure we formalize this as an assumption; needed: an atomic proposition does not change infinitely many times during a finite time horizon (zeno formulation??)
\jana{Hmm, how can we not have labels of regions apriori, but be able to test samples at all??}
\item Fast finding of A path that satisfies a complex mission. Pre-processing step to Informed RRT* or something like that. Prove that the important properties are maintained beyond reachability.
\jana{Hmm, how can this further guide informed RRT*? I have logically constrained stuff, so the distance is not the metric to guide it now. What is the metric?}
\end{itemize}

\newcommand{\must}{\Rightarrow}
\newcommand{\may}{\rightarrow}
\newcommand{\maynot}{\nrightarrow}
\newcommand{\mustnot}{\nRightarrow}

\begin{algorithm}
	\caption{one simple instantiation}\label{alg:learn}
	\begin{algorithmic}[1]
		\Require ``multi-modal'' transition system $(S,s_0,\must,\may,\maynot,\mustnot)$, $S\subseteq 2^{Ap}$, transition relations disjsoint (generally any quantity $S\times S\to \mathit{OrderedDomain}$, for instance a counter; here just $\{0,1,2,3\}$)
		\Procedure{learn}{transition from $s$ to $s'$}
		\State $t:=(s_1,s_2):=$(Labelling(s),Labelling(s'))
		\If{$t\notin\must$}
		\State add $t$ to $\must$
		\State domainOf Changes$:=\{a\in Ap\mid s_1(a)\oplus s_2(a)\}$
		\For{$s\in S$ agreeing with $s_1$ on domainOf Changes}
		\State \Comment or any other ``neighbourhood'' criterion / one can learn planning-style rules
		\State \Comment(and under some further conditions so that we don't overwrite $\may,\maynot$ back and forth)
		\State add (s,s changed to $s_2$ on domainOfChanges) to $\may$
		\EndFor
		\EndIf
		\State $counter(s_1)++$
		\If {$counter(s_1)>threshold$}
		\For{$s\in S: (s_1,s)\notin$ any relation}
		\State add $(s_1,s)$ to $\maynot$
		\EndFor
		\EndIf
		\EndProcedure
		\State\Comment and $\mustnot$ requires some really good reason to believe there is no transition
		\Procedure Ask{s}
		\State try using the best $\must$-successor or sample long to get an even better $\may$-successor
		\EndProcedure
		\newline
		\State Comments: 
		\State (1) $\maynot$ can be rather written as a $\may$ into ``?'' from where we can go anywhere; ASK tries to win with as few ?'s as possible
		\State (2) $\mustnot$ could be written as $\must$ into a sink ``:-(''
	\end{algorithmic}
\end{algorithm}

\end{document}

### 2019-11-04 learning Logic/draft.tex
.
Case study updated

--- 2019-11-01 notes-automata-logic.tex
@@
 
-\subsection{Problem Setup}
+\section{Problem Setup}
 
@@
 
-\subsection{Solution/Contribution draft}
+\section{Solution/Contribution draft}
 
@@
 \begin{figure}[h]
-	\includegraphics[scale=0.4]{auto}\caption{Models for the second automata example}
+	\includegraphics[scale=0.4]{Fig/auto}\caption{Models for the second automata example}
 \end{figure}
@@
 
+\subsection{A mini case study}
+
+Let us consider an office space with several rooms as illustrated in Fig.
+\ref{fig:office}.
+There are different objects in the office rooms.
+For simplicity, let us assume that each of the objects is either a desk (illustrated as a rectangle), or a chair (illustrated as a square), or a waste bin (illustrated as a circle).
+Consider a fully actuated robot with the initial position $x_0 \in \mathbb{R}^2$.
+We assume that the robot can recognize if it is close to a certain type of object or landmark in the environment; by close, we mean that the robot's state intersect with the object's/landmark's $\epsilon$-inflation (illustrated as dashed and dotted lines).
+
+The set of atomic propositions is as follows:
+\begin{itemize}
+\item $\mathit{Room}_1,\ldots,\mathit{Room}_6, \mathit{Hall}$, stand for being in a particular room or in a hallway
+\item $\mathit{Desk}$ stands for being close to a desk
+\item $\mathit{Chair}$ stands for being close to a chair
+\item $\mathit{Bin}$ stands for being close to a bin
+\item $\mathit{Corner}$ stands for being close to a corner
+\item $\mathit{Wall}$ stands for being close to a wall
+\end{itemize}
+
+The specification for the robot's behavior is given by the LTL formula:
+$$G F (\mathit{Room}_1 \wedge \mathit{Bin}) \wedge \ldots \wedge G F (\mathit{Room}_6 \wedge \mathit{Bin}).$$
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{Fig/small-office} \caption{A mini case study office environment illustration.}
+\label{fig:office}
+\end{figure}
+
 \end{document}

### 2020-04-09 learning Logic/draft.tex
a

@@
 \usepackage{amssymb,amsmath}
+\usepackage{tikz}
+\usetikzlibrary{automata, positioning, arrows}
+
+\tikzset{
+	->, % makes the edges directed
+	>=stealth', % makes the arrow heads bold
+	node distance=3cm, % specifies the minimum distance between two nodes. Change if necessary.
+	every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
+	initial text=$ $, % sets the text that appears on the start arrow
+}
 
@@
 
+\begin{algorithm}
+	\caption{Current Implementation: MAIN}\label{alg:current_main}
+	\begin{algorithmic}[1]
+		\State build trivial automaton
+		\Repeat
+		\State $adviceBDD$=\Call{ask}{$fromStates$} \Comment{Set of good transitions}
+		\State sample a bit to find a transition in $adviceBDD$ ($p \rightarrow q$)
+		\State \Call{learn}{$p,q$}
+		\State $fromStates = fromStates\vee q$
+		\Until{sampled all the way to satisfaction of the property}
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Current Implementation: LEARN}\label{alg:current_learn}
+	\begin{algorithmic}[1]
+		\Procedure{learn}{$s_1,s_2$}
+		\If{$\Call{getLevel}{s_1,s_2}$ $!=3$}\Comment{Possible levels: 0,1,2,3}
+		\State $\Call{setLevel}{s_1,s_2,3}$
+		\State $DOC^c=\{a\in APSystem\mid \neg(s_1(a)\oplus s_2(a))\}$\Comment{Used for $restrict$}
+		\State $R=restrict(s_1,DOC^c)$\Comment{states smilar to $s_1$}
+		\State $newTransitions=R \wedge \neg R' \bigwedge\limits_{a\in DOC^c} a\leftrightarrow a' \wedge level=2$
+		\State $productAutomaton $ $\vee= (newTransitions \wedge propertyBDD)$
+		\EndIf
+		\State $counter(s_1)++$
+		\If {$counter(s_1)>threshold$}
+		\For{$s\in S: (s_1,s) \wedge productAutomataBDD = false$}
+		\State \Call{addTransition} {$s_1,s,1$}
+		\EndFor
+		\EndIf
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Current Implementation: ASK}\label{alg:current_ask}
+	\begin{algorithmic}[1]
+		\Procedure{ask}{$fromStates$}
+		\State $finalStates=$\Call{$findFinalStates$}{$productAutomataBDD\wedge fromStates$}
+		\State $reachableStates.add(finalStates)$ 
+		\State $reachableStates.add(preFinalStates)$\Comment{required for transition acceptance}
+		\State $allReachableStates=currentStates[0]$
+		\State $i=1$
+		\While{$! (\neg allReachableStates \wedge \Call{pre}{allReachableStates}).isZero()$}
+			\State $reachableStates.add(\Call{pre}{reachableStates[i]})$
+			\State\Comment{add $(i+1)$th element}
+			\State $allReachableStates$ $\vee =reachableStates[i+1]$
+			\State $i++$
+
+		\EndWhile
+		\State $fromStates$ $\wedge=allReachableStates$
+		\State$toState=\Call{img}{fromStates}\wedge reachableStates[i]$
+		\State \Return{$getTransitions(fromStates, toStates)$}
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
 \subsection{A mini case study}
@@
 
+\subsection{A sample run of the algorithm}
+
 \end{document}

### 2020-04-09 MotionPlanning/src/Planning.java

@@ -2,21 +2,11 @@ import net.sf.javabdd.BDD;
 import net.sf.javabdd.BDDFactory;
 import net.sf.javabdd.BuDDyFactory;
 
-import jhoafparser.consumer.HOAConsumer;
-import jhoafparser.parser.HOAFParser;
-import jhoafparser.consumer.HOAIntermediate;
-
-import transitionSystem.TS;
+import modules.DefaultExperiment;
+import transitionSystem.Initialize;
 import transitionSystem.MarkovChain;
-import transitionSystem.MultiModalTS;
+import transitionSystem.ProductAutomaton;
 import transitionSystem.TSparser.MarkovChainParser;
-import transitionSystem.TSparser.BuchiAutomataParser;
-
-import java.io.BufferedReader;
-import java.io.BufferedWriter;
-import java.io.FileReader;
-import java.io.FileWriter;
-import java.io.InputStreamReader;
 
 import java.util.ArrayList;
 
@@ -28,108 +18,39 @@ public class Planning{
 
 	public static void main(String[] args) throws Exception{
 		
-		boolean verbose=true;
-		int cacheSizeForBDDFactory=1000;
-		int levelOfTransitions=2;
+		final boolean verbose=true;
+		int levelOfTransitions=4;
+		int threshold=20;
+		final int cacheSizeForBDDFactory=10000;
+		BDDFactory factory= BuDDyFactory.init(20,cacheSizeForBDDFactory);
 		
 		MarkovChain mc=MarkovChainParser.markovChainParser();
 		mc.checkMarkovChain();
-		ArrayList<String> apList=mc.getApList();
-
-		BDDFactory factory= BuDDyFactory.init(20,1000);
 		
-		
-		
+		// Creates propertyBDD and initializes productAutomaton
+        Initialize initialize= new Initialize(factory, threshold, levelOfTransitions, mc.getApList());
+        DefaultExperiment exper1=new DefaultExperiment(initialize.getProductAutomaton());
+        
+		int numVars[]=ProductAutomaton.numVars;
+        
+        BDD currentStates=mc.getBDDForState(factory, mc.getInitialState(), numVars[0]+numVars[1]+numVars[2]);
 
-		
-		
-		
-		BufferedReader propertyReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/property.pr"));
-        String propertyString=propertyReader.readLine();
-        propertyReader.close();
-        System.out.println("\nProperty to satisfty: "+propertyString);
-        
[…]

### 2020-04-09 MotionPlanning/src/modules/Ask.java

@@ -1,14 +1,0 @@
-package modules;
-
-import net.sf.javabdd.BDD;
-import net.sf.javabdd.BDDFactory;
-
-public class Ask
-{
-
-	public Ask(BDDFactory factory, BDD bdd)
-	{
-		
-	}
-
-}

### 2020-04-09 MotionPlanning/src/modules/DefaultExperiment.java

@@ -1,0 +1,136 @@
+package modules;
+
+
+import java.util.ArrayList;
+import java.util.List;
+
+import net.sf.javabdd.BDD;
+import net.sf.javabdd.BDDFactory;
+import transitionSystem.ProductAutomaton;
+
+/**
+ * <p> Original algo for learn and ask procedures</p>
+ * @author kush
+ *
+ */
+
+
+public class DefaultExperiment implements Laboratory{
+	
+	ProductAutomaton productAutomaton;
+	BDDFactory factory;
+	BDD productAutomatonBDD, propertyBDD;
+	public DefaultExperiment(BDD propertyBDD, 
+			int[] numVars, 
+			ArrayList<String> apListProperty,
+			ArrayList<String> apListSystem)
+	{
+		this.productAutomaton=new ProductAutomaton(propertyBDD);
+		this.factory=productAutomaton.getBDD().getFactory();
+		this.productAutomatonBDD=productAutomaton.getBDD();
+	}
+	
+	public DefaultExperiment(ProductAutomaton productAutomaton) {
+		this.productAutomaton=productAutomaton;
+		this.factory=productAutomaton.getBDD().getFactory();
+		this.productAutomatonBDD=productAutomaton.getBDD();
+	}
+
+	@Override
+	public void learn(BDD fromState, BDD toState) throws Exception {
+		if(productAutomaton.getLevel(fromState, toState)==-1) {
+			productAutomaton.addTransition(fromState, toState, 3);	
+		}
+		else if(productAutomaton.getLevel(fromState, toState)!=3) {
+			productAutomaton.setLevel(fromState, toState, 3);
+		}
+		
+//		debug
+//		productAutomaton.createDot(0);
+		BDD complementDomainOfChanges=allExceptDomainOfChanges(fromState,toState);
+		
+		BDD fromStateSimilar=fromState.exist(complementDomainOfChanges);
+		BDD toStateSimilar=factory.one();
+		for(int i=0;i<ProductAutomaton.numAPSystem;i++) {
+			if(fromStateSimilar.and(ProductAutomaton.ithVarSystemPre(i)).isZero()) {
+				toStateSimilar.andWith((ProductAutomaton.ithVarSystemPre(i)));
+			}
+			else if(fromStateSimilar.and(ProductAutomaton.ithVarSystemPre(i).not()).isZero()) {
[…]

### 2020-04-09 MotionPlanning/src/modules/Laboratory.java

@@ -1,0 +1,12 @@
+package modules;
+
+import net.sf.javabdd.BDD;
+
+public interface Laboratory
+{
+//	ProductAutomaton productAutomaton;
+
+	public void learn(BDD fromState, BDD toState) throws Exception;
+	
+	public BDD ask(BDD startingStates);
+}

### 2020-04-09 MotionPlanning/src/printing/PrintProductAutomaton.java

@@ -1,0 +1,118 @@
+package printing;
+
+import java.io.BufferedWriter;
+import java.io.FileWriter;
+import java.io.IOException;
+
+import net.sf.javabdd.BDD;
+import net.sf.javabdd.BDD.BDDIterator;
+import transitionSystem.ProductAutomaton;
+import transitionSystem.extraExceptions.stateException;
+
+public class PrintProductAutomaton
+{
+	ProductAutomaton productAutomaton;
+	BufferedWriter writer;
+	boolean flag[];
+	public PrintProductAutomaton(ProductAutomaton productAutomaton,int num) throws IOException, stateException{
+		flag=new boolean[productAutomaton.numberOfStates()];
+		writer = new BufferedWriter(new FileWriter("/home/kush/Projects/robotmotionplanning/MotionPlanning/temp/productAutomaton"+num+".dot"));
+		
+		this.productAutomaton=productAutomaton;
+		BDD productAutomatonBDD=productAutomaton.getBDD();
+		
+		int[] numVars=ProductAutomaton.numVars;
+		int totalVars=0;
+		for(int i=0;i<numVars.length;i++) {
+			totalVars+=numVars[i];
+		}
+		BDD allVars=productAutomatonBDD.getFactory().one();
+		for(int i=0;i<totalVars;i++) {
+			allVars.andWith(productAutomatonBDD.getFactory().ithVar(i));
+		}
+		
+		writer.write("digraph G {\n");
+		BDDIterator iterator=productAutomatonBDD.iterator(allVars);
+		while(iterator.hasNext()) {
+			BDD transition=(BDD) iterator.next();
+			addTransitionToDotFile(transition);
+		}
+		writer.append("}");
+		
+		writer.close();
+	}
+
+	private void addTransitionToDotFile(BDD transition) throws IOException, stateException{
+		
+		BDD fromState=productAutomaton.removeAllExceptPreVars(transition);
+		BDD toState=productAutomaton.removeAllExceptPostVars(transition);
+		toState=productAutomaton.changePostVarsToPreVars(toState);
+		int fromStateID=productAutomaton.getStateID(fromState);
+		int toStateID=productAutomaton.getStateID(toState);
+		int acceptingSet=transition.scanVar(ProductAutomaton.acceptingSetDomain()).intValue();
+
+		if(!flag[fromStateID]) {
+			flag[fromStateID]=true;
[…]

### 2020-04-09 MotionPlanning/src/printing/package-info.java

@@ -1,0 +1,1 @@
+package printing;

### 2020-04-09 MotionPlanning/src/transitionSystem/BuchiAutomaton.java

@@ -1,19 +1,0 @@
-package transitionSystem;
-
-import java.util.ArrayList;
-
-import net.sf.javabdd.BDDFactory;
-
-public class BuchiAutomaton extends TS
-{
-
-	public BuchiAutomaton(BDDFactory factory, 
-							int cacheSize, 
-							ArrayList<String> apList, 
-							ArrayList<ArrayList<String>> labelling){
-
-		super(factory, cacheSize, apList, labelling, 2);
-
-	}
-
-}

### 2020-04-09 MotionPlanning/src/transitionSystem/Initialize.java

@@ -1,0 +1,89 @@
+package transitionSystem;
+
+import java.io.BufferedReader;
+import java.io.BufferedWriter;
+import java.io.FileReader;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.io.InputStreamReader;
+import java.util.ArrayList;
+
+import jhoafparser.parser.generated.ParseException;
+import net.sf.javabdd.BDDFactory;
+import transitionSystem.TSparser.BuchiAutomataParser;
+
+/**
+ * <p>Creates the BDD of the property and Initializes the product automata</p>
+ * @see ProductAutomaton
+ * @author Kush Grover
+ */
+
+public class Initialize
+{
+	ProductAutomaton productAutomaton;
+	
+	public Initialize(BDDFactory factory, 
+			int threshold, 
+			int levelOfTransitions, 
+			ArrayList<String> apListSystem) throws IOException, ParseException{
+		
+		BufferedReader propertyReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/Examples/Example 1/property.pr"));
+        String propertyString=propertyReader.readLine();
+        propertyReader.close();
+        System.out.println("\nProperty to satisfty: "+propertyString);
+        
+//		String command="/home/kush/Projects/robotmotionplanning/spot-2.8.5/bin/ltl2tgba";
+        String command="/home/kush/Projects/robotmotionplanning/owl/build/install/owl/bin/ltl2ldba";
+		
+		
+//      ProcessBuilder builder1 = new ProcessBuilder(command,"--deterministic",propertyString);
+        ProcessBuilder builder1 = new ProcessBuilder(command,propertyString);
+        builder1.redirectErrorStream(true);
+        Process p1 = builder1.start();
+        
+//      ProcessBuilder builder2 = new ProcessBuilder(command,"--deterministic",propertyString);
+        ProcessBuilder builder2 = new ProcessBuilder(command,propertyString);
+        builder2.redirectErrorStream(true);
+        Process p2 = builder2.start();
+        
+//      ProcessBuilder builder3 = new ProcessBuilder(command,"--deterministic",propertyString);
+        ProcessBuilder builder3 = new ProcessBuilder(command,propertyString);
+        builder3.redirectErrorStream(true);
+        Process p3 = builder2.start();
+        
[…]

### 2020-04-09 MotionPlanning/src/transitionSystem/MarkovChain.java

@@ -3,6 +3,9 @@ package transitionSystem;
 import java.util.ArrayList;
 import java.util.List;
 
+import net.sf.javabdd.BDD;
+import net.sf.javabdd.BDDFactory;
+
 
 public class MarkovChain{
 	private ArrayList<String> apList;
@@ -47,7 +50,7 @@ public class MarkovChain{
 		return size;
 	}
 	
-	public int getInitial() {
+	public int getInitialState() {
 		return initial;
 	}
 	
@@ -98,8 +101,9 @@ public class MarkovChain{
 			for(int j=0;j<currentEdges.size();j++) {
 				totalProb+=currentEdges.get(j).prob;
 			}
-			if(totalProb!=1.0) {
-				throw new Exception("Error: MarkovChainNotComplete");
+			if(totalProb-1.0>0.00001 || 1-totalProb>0.00001) {
+				System.out.println(totalProb);
+				throw new Exception("Error: MarkovChainNotComplete: state "+i);
 			}
 		}
 	}
@@ -133,4 +137,29 @@ public class MarkovChain{
 		return apForStateArray;
 	}
 
+	public BDD getBDDForState(BDDFactory factory, int state, int varsBeforeSystemVars) {
+		int[] toStateAP=findAPForState(state);
+		BDD toStateBDD=factory.one();
+        for(int i=0;i<apList.size();i++) {
+        	if(arraySearch(toStateAP,i)) {
+        		toStateBDD.andWith(factory.ithVar(i+varsBeforeSystemVars));
+        	}
+        	else {
+        		toStateBDD.andWith(factory.ithVar(i+varsBeforeSystemVars).not());
+        	}
+        }
+        return toStateBDD;
+	}
+
+
+
+	private boolean arraySearch(int[] array, int i)
+	{
+		for(int j=0;j<array.length;j++) {
+			if(array[j]==i) {
+				return true;
+			}
+		}
+		return false;
+	}
 }

### 2020-04-09 MotionPlanning/src/transitionSystem/MultiModalTS.java

@@ -1,26 +1,0 @@
-package transitionSystem;
-
-import java.util.ArrayList;
-
-import net.sf.javabdd.BDD;
-import net.sf.javabdd.BDDFactory;
-
-import transitionSystem.TS;
-
-public class MultiModalTS extends TS
-{
-	ArrayList<String> apList;
-	int size;
-	BDD bdd;
-//	public MultiModalTS(ArrayList<String> apList, int cacheSize, int level, BDDFactory factory){
-//		super(apList, cacheSize, level, factory);	
-//	}
-	public MultiModalTS(BDDFactory factory, 
-						int cacheSize, 
-						ArrayList<String> apList, 
-						ArrayList<ArrayList<String>> labelling, 
-						int levelOfTransitions){
-		
-		super(factory, cacheSize, apList, labelling, levelOfTransitions);
-	}
-}

### 2020-04-09 MotionPlanning/src/transitionSystem/ProductAutomaton.java

@@ -1,0 +1,574 @@
+package transitionSystem;
+
+import net.sf.javabdd.BDD;
+import net.sf.javabdd.BDD.BDDIterator;
+import net.sf.javabdd.BDDDomain;
+import net.sf.javabdd.BDDFactory;
+import net.sf.javabdd.BDDPairing;
+import printing.PrintProductAutomaton;
+import transitionSystem.extraExceptions.stateException;
+import transitionSystem.extraExceptions.transitionException;
+
+import java.io.IOException;
+import java.util.ArrayList;
+
+/**
+ * <p> Stores the product automaton and does all the operations on it</p>
+ * @author Kush Grover
+ *
+ */
+
+public class ProductAutomaton{
+	/**
+	 * <p>propertyBDD stores the property automaton
+	 * numVars[0] = number of vars used in property automaton
+	 * numVars[1] = number of vars used in transition level
+	 * numVars[2] = number of vars used in label of transitions in property automaton=number of atomic props used in property automaton
+	 * numVars[3] = number of vars used in system automaton = 2 * numAPSystem
+	 * threshold and counter are to add level 1 transitions
+	 * apListProperty and apListSystem are lists of atomic props used in property and automaton resp.</p>
+	 */
+	BDD productAutomatonBDD, propertyBDD;
+	public static BDDFactory factory;
+	public static int numVars[];
+	public static int numAPSystem, threshold, levelOfTransitions;
+	public static ArrayList<String> apListProperty, apListSystem;
+	static BDDPairing oldToNewPairing, newToOldPairing, statesToLabelPairing;
+	int[] counter;
+	
+	
+	/**
+	 * <p>Initializes Product Automaton to an empty automaton</p>
+	 * @param propertyBDD BDD for the property automaton
+	 */
+	public ProductAutomaton(BDD propertyBDD){	
+		this.propertyBDD=propertyBDD;
+		this.productAutomatonBDD=factory.zero();
+		initializeCounter();
+		oldToNewPairing=oldVarToNewVarPairing();
+		newToOldPairing=newVarToOldVarPairing();
+		statesToLabelPairing=statesToLabelPairing();
+	}
+
+	/**
+	 * <p>Creating the counter array and initializing all values to 0</p>
+	 */
[…]

### 2020-04-09 MotionPlanning/src/transitionSystem/TS.java

@@ -26,6 +26,7 @@ public class TS{
 //	private BDDDomain[] bddDomainForLearnedAutomata;
 	private int initial;
 	private ArrayList<ArrayList<String>> labelling;
+	private int varsUsedInBuchiAutomata;
 	
 	public TS(BDDFactory factory, 
 				int cacheSize, 
@@ -63,19 +64,23 @@ public class TS{
 		this.labelling=labelling;
 	}
 	
-	public void addTransitionFromListOfAP(int[] currentState, int[] nextState, int varsBuchi){
+	public void setVarsUsedInBuchiAutomata(int varsUsedInBuchiAutomata) {
+		this.varsUsedInBuchiAutomata=varsUsedInBuchiAutomata;
+	}
+	
+	public void addTransitionFromListOfAP(int[] currentState, int[] nextState){
 		
 		BDD currentStateBDD=factory.one();
 		BDD nextStateBDD=factory.one();
-		System.out.println("vasrBuhci= "+varsBuchi);
+//		System.out.println("varsBuhci= "+varsUsedInBuchiAutomata);
 		
 		for(int i=0;i<currentState.length;i++) {
-			currentStateBDD=currentStateBDD.and(factory.ithVar(currentState[i]+varsBuchi));
+			currentStateBDD=currentStateBDD.and(factory.ithVar(currentState[i]+varsUsedInBuchiAutomata));
 //			System.out.println("currentState "+i+" "+currentState[i]);
 //			currentStateBDD.andWith(bddDomainForLearnedAutomata[0].ithVar(currentState[i]));
 		}
 		for(int i=0;i<nextState.length;i++) {
-			nextStateBDD=nextStateBDD.and(factory.ithVar(size+nextState[i]+varsBuchi));
+			nextStateBDD=nextStateBDD.and(factory.ithVar(size+nextState[i]+varsUsedInBuchiAutomata));
 //			System.out.println("nextState "+i+" "+nextState[i]);
 //			nextStateBDD.andWith(bddDomainForLearnedAutomata[1].ithVar(nextState[i]));
 		}

### 2020-04-09 MotionPlanning/src/transitionSystem/TSparser/BuchiAutomataParser.java

@@ -3,19 +3,14 @@ package transitionSystem.TSparser;
 //package jhoafparser.examples;
 
 import java.io.InputStream;
+import java.util.ArrayList;
 import java.util.List;
 
 import jhoafparser.ast.AtomLabel;
 import jhoafparser.ast.BooleanExpression;
-import jhoafparser.consumer.HOAConsumer;
-import jhoafparser.consumer.HOAConsumerException;
-import jhoafparser.consumer.HOAConsumerNull;
 import jhoafparser.consumer.HOAConsumerStore;
-import jhoafparser.consumer.HOAIntermediate;
 import jhoafparser.parser.HOAFParser;
 import jhoafparser.parser.generated.ParseException;
-import jhoafparser.storage.StoredAutomaton;
-import jhoafparser.storage.StoredEdgeWithLabel;
 import net.sf.javabdd.BDD;
 import net.sf.javabdd.BDDDomain;
 import net.sf.javabdd.BDDFactory;
@@ -27,9 +22,10 @@ public class BuchiAutomataParser{
 	BDDFactory factory;
 	int numberOfStates=0;
 	int acceptingSets=0;
-	int[] varArray=new int[3];
-	int numberOfVarsUsed;
+	int numVars[]=new int[4];
 	BDDDomain[] bddDomain;
+	ArrayList<String> apListSystem, apListProperty;
+	int levelOfTransitions;
 
 	public class addEdgeToBDD extends HOAConsumerStore{
 		public addEdgeToBDD() {
@@ -42,23 +38,23 @@ public class BuchiAutomataParser{
 									java.util.List<java.lang.Integer> conjSuccessors, 
 									java.util.List<java.lang.Integer> accSignature) {
 
-			BDD temp=factory.one();
-			
+			BDD newEdge=factory.one();
 			for(Integer i : conjSuccessors) {
-				temp=factory.one();
-				temp.andWith(bddDomain[0].ithVar(stateId));
-				temp.andWith(bddDomain[1].ithVar(i));
+				newEdge=factory.one();
+				newEdge.andWith(bddDomain[0].ithVar(stateId));
+				newEdge.andWith(bddDomain[1].ithVar(i));
 				
 				try {
-					temp.andWith(bddDomain[2].ithVar(accSignature.get(i)+1));
-					System.out.println("added edge in buchi from state "+stateId+" ----"+labelExpr.toString()+"----> "+i+" {"+accSignature.get(i)+"}");
[…]

### 2020-04-09 MotionPlanning/src/transitionSystem/TSparser/MarkovChainParser.java

@@ -9,7 +9,7 @@ import java.util.StringTokenizer;
 import transitionSystem.MarkovChain;
 
 /**
- * @author kush
+ * @author Kush Grover
  *
  */
 public class MarkovChainParser
@@ -20,7 +20,7 @@ public class MarkovChainParser
 		MarkovChain mc=new MarkovChain();
 		
 		
-		BufferedReader apFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/atomicPropositions.ap"));
+		BufferedReader apFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/Examples/Example 1/atomicPropositions.ap"));
         String line=apFileReader.readLine();
 		
 		while(line!=null) {
@@ -30,7 +30,7 @@ public class MarkovChainParser
 		apFileReader.close();
 		
 		
-        BufferedReader statesFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/markovChainStates.st"));
+        BufferedReader statesFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/Examples/Example 1/markovChainStates.st"));
         line = statesFileReader.readLine(); //read 1st line, number of states
         mc.setSize(Integer.parseInt(line));
        
@@ -58,7 +58,7 @@ public class MarkovChainParser
         
         
         
-        BufferedReader transitionsFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/markovChainTransitions.tr"));
+        BufferedReader transitionsFileReader = new BufferedReader(new FileReader("/home/kush/Projects/robotmotionplanning/MotionPlanning/Examples/Example 1/markovChainTransitions.tr"));
         line = transitionsFileReader.readLine();
         int numOfTransitions=Integer.parseInt(line);
         int firstState,secondState;

### 2020-04-09 MotionPlanning/src/transitionSystem/extraExceptions/package-info.java

@@ -1,0 +1,1 @@
+package transitionSystem.extraExceptions;

### 2020-04-09 MotionPlanning/src/transitionSystem/extraExceptions/stateException.java

@@ -1,0 +1,40 @@
+package transitionSystem.extraExceptions;
+
+public class stateException extends Exception
+{
+
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 1L;
+
+	public stateException()
+	{
+		// TODO Auto-generated constructor stub
+	}
+
+	public stateException(String message)
+	{
+		super(message);
+		// TODO Auto-generated constructor stub
+	}
+
+	public stateException(Throwable cause)
+	{
+		super(cause);
+		// TODO Auto-generated constructor stub
+	}
+
+	public stateException(String message, Throwable cause)
+	{
+		super(message, cause);
+		// TODO Auto-generated constructor stub
+	}
+
+	public stateException(String message, Throwable cause, boolean enableSuppression, boolean writableStackTrace)
+	{
+		super(message, cause, enableSuppression, writableStackTrace);
+		// TODO Auto-generated constructor stub
+	}
+
+}

### 2020-04-09 MotionPlanning/src/transitionSystem/extraExceptions/transitionException.java

@@ -1,0 +1,40 @@
+package transitionSystem.extraExceptions;
+
+public class transitionException extends Exception
+{
+
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 1L;
+
+	public transitionException()
+	{
+		// TODO Auto-generated constructor stub
+	}
+
+	public transitionException(String message)
+	{
+		super(message);
+		// TODO Auto-generated constructor stub
+	}
+
+	public transitionException(Throwable cause)
+	{
+		super(cause);
+		// TODO Auto-generated constructor stub
+	}
+
+	public transitionException(String message, Throwable cause)
+	{
+		super(message, cause);
+		// TODO Auto-generated constructor stub
+	}
+
+	public transitionException(String message, Throwable cause, boolean enableSuppression, boolean writableStackTrace)
+	{
+		super(message, cause, enableSuppression, writableStackTrace);
+		// TODO Auto-generated constructor stub
+	}
+
+}

### 2020-05-05 learning Logic/draft.tex
a

@@
 		\State build trivial automaton
-		\Repeat
-		\State $adviceBDD$=\Call{ask}{$fromStates$} \Comment{Set of good transitions}
-		\State sample a bit to find a transition in $adviceBDD$ ($p \rightarrow q$)
-		\State \Call{learn}{$p,q$}
-		\State $fromStates = fromStates\vee q$
-		\Until{sampled all the way to satisfaction of the property}
+		\State $currentStates=initState$
+		\While{true}
+			\State $reachableStates=\Call{ask}{currentStates}$ 
+			\State \Comment{array of states backward reachable from final states}
+			\State $pathLength=1$
+			\While{$pathLength<reachableStates.size()$}
+				\State $fromStates=currentStates\wedge reachableStates[pathLength]$
+				\State $toStates=reachableStates[pathLength-1]$
+				\State$t:=(p,q')=\Call{Sample}{fromStates, toStates}$
+				\If{$t$ $!=null$}
+					\State break
+				\EndIf 
+				\State $pathLength++$
+			\EndWhile
+		\If{$t==null$}
+		\State $t=\Call{Sample}{currentStates}$
+		\EndIf
+		\State $currentStates$ $\vee= q$
+		\State $\Call{learn}{p,q}$
+		\If{accepting path exists}
+		\State \Return{$path$}
+		\EndIf
+		\EndWhile
 	\end{algorithmic}
@@
 		\Procedure{learn}{$s_1,s_2$}
-		\If{$\Call{getLevel}{s_1,s_2}$ $!=3$}\Comment{Possible levels: 0,1,2,3}
-		\State $\Call{setLevel}{s_1,s_2,3}$
+		\State $t=s_1\wedge s_2'$
+		\State $sampledTransitions\ \vee = t$
+		\State $removeTransitions(t)$\Comment{remove transition of smaller level}
+		\State $t=\Call{addTransitions}{t,3}$	\Comment{$t$ will be converted to product automaton transitions}
+
 		\State $DOC^c=\{a\in APSystem\mid \neg(s_1(a)\oplus s_2(a))\}$\Comment{Used for $restrict$}
-		\State $R=restrict(s_1,DOC^c)$\Comment{states smilar to $s_1$}
+		\State $R=remove(s_1,DOC^c)$\Comment{states smilar to $s_1$}
 		\State $newTransitions=R \wedge \neg R' \bigwedge\limits_{a\in DOC^c} a\leftrightarrow a' \wedge level=2$
+		\State\Comment{$\neg R':= \bigwedge\limits_{a \text{ in } R} \neg a'$}
+		\State $newTransitions$ $\wedge=\neg sampledTransitions$
 		\State $productAutomaton $ $\vee= (newTransitions \wedge propertyBDD)$
-		\EndIf
 		\State $counter(s_1)++$
 		\If {$counter(s_1)>threshold$}
-		\For{$s\in S: (s_1,s) \wedge productAutomataBDD = false$}
+		\For{$s\in S: (s_1,s) \wedge sampledTransitions = false$}
+			\State \Call{removeTransitions}{$s_1,s_s$}
 			\State \Call{addTransition} {$s_1,s,1$}
@@
 		\State $reachableStates.add(preFinalStates)$\Comment{required for transition acceptance}
-		\State $allReachableStates=currentStates[0]$
 		\State $i=1$
-		\While{$! (\neg allReachableStates \wedge \Call{pre}{allReachableStates}).isZero()$}
-			\State $reachableStates.add(\Call{pre}{reachableStates[i]})$
-			\State\Comment{add $(i+1)$th element}
-			\State $allReachableStates$ $\vee =reachableStates[i+1]$
+		\While{$\neg allReachableStates \wedge \Call{pre}{reachableStates[i]}\ !=\bot$}
+			\State $reachableStates.add(reachableStates[i]\vee \Call{pre}{reachableStates[i]})$
+			\State $allReachableStates\ \vee=reachableStates[i]$
 			\State $i++$
-
 		\EndWhile
-		\State $fromStates$ $\wedge=allReachableStates$
-		\State$toState=\Call{img}{fromStates}\wedge reachableStates[i]$
-		\State \Return{$getTransitions(fromStates, toStates)$}
+		\State \Return{$reachableStates$}
 		\EndProcedure

### 2020-05-05 MotionPlanning/src/Planning.java

@@ -40,9 +40,9 @@ public class Planning{
         productAutomaton.setInitState(productAutomaton.getInitStates().and(initStateSystem));
         BDD currentStates=initStateSystem.id();
         
-        productAutomaton.getBDD().andWith(exper.addFilters());
+//        productAutomaton.getBDD().andWith(exper.addFilters());
         
-        double timeForSampling=0, preTimeSampling=0, postTimeSampling=0;
+        double timeForSampling=0, preTimeSampling=0, postTimeSampling=0, pathTime=0;
         int iterationNumber=0;
         while(true) { //until the property is satisfied
 
@@ -74,19 +74,19 @@ public class Planning{
         	}
         	currentStates=currentStates.or(productAutomaton.getSecondStateSystem(transition));
         	transition=exper.learn(productAutomaton.getFirstStateSystem(transition), productAutomaton.getSecondStateSystem(transition));
-//        	transition.printDot();
-        	if(productAutomaton.findAcceptingPath()) {
+        	double pathStartTime=System.nanoTime();
+        	ArrayList<BDD> path=productAutomaton.findAcceptingPath();
+        	double pathEndTime=System.nanoTime();
+        	pathTime+=pathEndTime-pathStartTime;
+        	if(path!=null) {
+        		productAutomaton.printPath(path);
         		System.out.println("Yay!!!!!");
         		break;
         	}
-        	
-        	
         	initialize.getProductAutomaton().createDot(iterationNumber);
         	iterationNumber++;
         }
-
-		
-		
+        
         factory.done();
 		
 		
@@ -95,9 +95,11 @@ public class Planning{
 		System.out.print("\n\nTotal time taken (in ms):");
         System.out.println((endTime-startTime)/1000000);
         System.out.print("Time taken sampling (in ms):");
-        System.out.println((timeForSampling)/1000000);
+        System.out.println(timeForSampling/1000000);
         System.out.print("Time taken other than sampling (in ms):");
         System.out.println((endTime-startTime-timeForSampling)/1000000);
+        System.out.print("Path Time (in ms):");
+        System.out.println(pathTime/1000000);
 
 	}
 

### 2020-05-05 MotionPlanning/src/modules/DefaultExperiment.java

@@ -3,7 +3,6 @@ package modules;
 
 import java.util.ArrayList;
 import net.sf.javabdd.BDD;
-import net.sf.javabdd.BDD.BDDIterator;
 import net.sf.javabdd.BDDFactory;
 import transitionSystem.ProductAutomaton;
 import transitionSystem.extraExceptions.learningException;
@@ -43,6 +42,7 @@ public class DefaultExperiment implements Experiments{
 			productAutomaton.sampledTransitions=productAutomaton.sampledTransitions.or(transition);
 			productAutomaton.removeTransition(fromState, toState);
 			transition=productAutomaton.addTransition(fromState, toState, 3);
+			productAutomaton.sampledProductTransitions=productAutomaton.sampledProductTransitions.or(transition);
 		}
 		else {
 			throw new learningException("Transition already learned");
@@ -52,17 +52,13 @@ public class DefaultExperiment implements Experiments{
 		
 		
 		int counter=productAutomaton.increaseCounter(productAutomaton.getFirstState(fromState));
+		
+		
+		
+		// never tested this
 		if(counter>ProductAutomaton.threshold) {
 			productAutomaton.setLevel(fromState,1);
 		}
-//		BDDIterator iterator=productAutomaton.removeAllExceptPreVars(similarTransitions).iterator(ProductAutomaton.allPreVars());
-//		while(iterator.hasNext()) {
-//			BDD temp=(BDD) iterator.next();
-//			int counter=productAutomaton.increaseCounter(productAutomaton.getFirstState(temp));
-//			if(counter>ProductAutomaton.threshold) {
-//				productAutomaton.setLevel(temp,1);
-//			}
-//		}
 		return transition;
 	}
 
@@ -129,21 +125,21 @@ public class DefaultExperiment implements Experiments{
 		BDD h=ProductAutomaton.ithVarSystemPre(0);
 		BDD r1=ProductAutomaton.ithVarSystemPre(1);
 		BDD r2=ProductAutomaton.ithVarSystemPre(2);
-//		BDD r3=ProductAutomaton.ithVarSystemPre(3);
[…]

### 2020-05-05 MotionPlanning/src/printing/PrintAcceptingPath.java

@@ -1,0 +1,42 @@
+package printing;
+
+import java.util.ArrayList;
+
+import net.sf.javabdd.BDD;
+import transitionSystem.ProductAutomaton;
+
+public class PrintAcceptingPath {
+	public PrintAcceptingPath(ArrayList<BDD> path) throws Exception {
+		BDD next;
+		for(int i=0;i<path.size();i++) {
+			next=path.get(i);
+			printAPList(next);
+			
+		}
+	}
+
+	private void printAPList(BDD state) throws Exception {
+		ArrayList<String> apList=findAPList(state);
+		System.out.print("[");
+		for(int j=0;j<apList.size();j++) {
+			if(j<apList.size()-1) {
+				System.out.print(apList.get(j)+",");
+			}
+			else {
+				System.out.print(apList.get(j));
+			}
+		}
+		System.out.println("]");
+	}
+
+	private ArrayList<String> findAPList(BDD state) throws Exception {
+		ArrayList<String> list=new ArrayList<String>();
+		for(int i=0;i<ProductAutomaton.numAPSystem;i++) {
+			if(! state.and(ProductAutomaton.ithVarSystemPre(i)).isZero()) {
+				list.add(ProductAutomaton.apListSystem.get(i));
+			}
+		}
+		list.add(Integer.toString(state.scanVar(ProductAutomaton.propertyDomainPre()).intValue()));
+		return list;
+	}
+}

### 2020-05-05 MotionPlanning/src/transitionSystem/ProductAutomaton.java

@@ -1,11 +1,11 @@
 package transitionSystem;
 
 import net.sf.javabdd.BDD;
-import net.sf.javabdd.BDD.BDDIterator;
 import net.sf.javabdd.BDDDomain;
 import net.sf.javabdd.BDDException;
 import net.sf.javabdd.BDDFactory;
 import net.sf.javabdd.BDDPairing;
+import printing.PrintAcceptingPath;
 import printing.PrintProductAutomaton;
 import transitionSystem.emptinessCheck.EmptinessCheck;
 import transitionSystem.extraExceptions.StateException;
@@ -32,7 +32,7 @@ public class ProductAutomaton{
 	BDD productAutomatonBDD;
 	BDD initStates;
 	static BDD propertyBDD;
-	public BDD sampledTransitions;
+	public BDD sampledTransitions, sampledProductTransitions;
 	public static BDDFactory factory;
 	public static int numVars[];
 	public static int numAPSystem, threshold, levelOfTransitions, varsBeforeSystemVars, numAllVars;
@@ -52,6 +52,7 @@ public class ProductAutomaton{
 		ProductAutomaton.propertyBDD=propertyBDD;
 		this.productAutomatonBDD=factory.zero();
 		this.sampledTransitions=factory.zero();
+		this.sampledProductTransitions=factory.zero();
 		
 		varsBeforeSystemVars=numVars[0]+numVars[1]+numVars[2];
 		numAllVars=varsBeforeSystemVars+2*numAPSystem;
@@ -184,6 +185,10 @@ public class ProductAutomaton{
 	
 	public static int getNumStates() {
 		return numStates;
+	}
+	
+	public BDD getSampledProductTransitions() {
+		return sampledProductTransitions;
 	}
 	
 	/**
@@ -616,11 +621,11 @@ public class ProductAutomaton{
 	 * @throws StateException 
 	 */
 	public BDD preImage(BDD states) throws StateException, Exception {
-		if(! hasOnlyPostVars(states)) {
+		if(! hasOnlyPreVars(states)) {
 			throw new StateException("BDD has extra vars");
 		}
 		BDD temp=factory.zero();
-		temp.orWith(changePreVarsToPostVars(states));
+		temp=temp.or(changePreVarsToPostVars(states));
 		return removeAllExceptPreVars(productAutomatonBDD.and(temp));
 	}
 	
[…]

### 2020-05-05 MotionPlanning/src/transitionSystem/emptinessCheck/EmptinessCheck.java

@@ -24,6 +24,7 @@ public class EmptinessCheck {
 	int max=0;
 	int totalAcceptingSets;
 	int[] rank;
+	BDD statesSCC;
 	public EmptinessCheck(ProductAutomaton productAutomaton){
 		this.productAutomaton=productAutomaton;
 		factory=ProductAutomaton.factory;
@@ -36,7 +37,8 @@ public class EmptinessCheck {
 		}
 	}
 	
-	public boolean findAcceptingPath() throws Exception{
+//	public boolean findAcceptingPath() throws Exception{
+	public ArrayList<BDD> findAcceptingPath() throws Exception{
 		push(null, productAutomaton.getInitStates());
 		
 		while(! todo.isEmpty()) {
@@ -56,86 +58,99 @@ public class EmptinessCheck {
 				}
 				else if(getRank(toState)>0) {
 					if(containAllAcc(merge(accSet,getRank(toState)))) {
-						return true;
-//						return findPath();
+						return findPath();
 					}
 				}
 			}
 		}
-		return false;
-//		return null;		
+		return null;		
 	}
 	
 	private ArrayList<BDD> findPath() throws Exception {
-		BDD statesSCC=factory.zero();
+		BDD statesSCC=initializeStatesSCC();
 		ArrayList<BDD> path=new ArrayList<BDD>();
-		int topRank=SCC.getTop().getRoot();
-		int[] visitedAcceptingSets=new int[totalAcceptingSets];
+		ArrayList<BDD> milestones=new ArrayList<BDD>();
+		int[] visitedAcceptingSets=new int[totalAcceptingSets+1];
+		
+		milestones.add(statesSCC.satOne(ProductAutomaton.allPreVars(),false));
+		bfs(productAutomaton.getInitStates(),milestones.get(0),path);
 		for(int i=0;i<totalAcceptingSets;i++) {
-			visitedAcceptingSets[i]=0;
+ 			milestones.add(bfs(milestones.get(i),visitedAcceptingSets,path));
 		}
+		bfs(milestones.get(totalAcceptingSets),milestones.get(0),path);
+		return path;
+	}
+
+	private BDD initializeStatesSCC() throws Exception {
+		statesSCC=factory.zero();
+		int topRank=SCC.getTop().getRoot();
 		for(int i=0;i<ProductAutomaton.getNumStates();i++){
 			if(rank[i]>=topRank) {
 				statesSCC=statesSCC.or(productAutomaton.getStateFromID(i));
 			}
 		}
[…]

### 2020-05-06 learning Logic/draft.tex
a

@@
 \subsection{A sample run of the algorithm}
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{Fig/example} \caption{A smaller example to illustrate the algorithm}
+	\label{fig:example}
+\end{figure}
 
+Look at the system in Figure \ref{fig:example}.
+The property we want to satisfy for this system is $G F (r_1 \wedge b) \wedge G F (r_2\wedge b)$ and the deterministic B\"uchi automaton for this LTL formula is also given in Figure \ref{fig:example}.
+The algorithm stores the product automaton as a BDD, intially it is $\bot$.
+$currentStates$ stores the set of states we have seen till now which is initialized to $initState$.
+For this Example $initState=\{H,0\}$.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{Fig/product2} \caption{Product automaton after 2 iterations}
+	\label{fig:product2}
+\end{figure}
+
+After sampling transitions $\{H\} \rightarrow \{r_1\}$ and $\{r_1\}\rightarrow \{r_1,b\}$ initially, the product automaton looks like figure \ref{fig:product2}.
+Now, in line 4, $ASK$ procedure returns a list of BDDs such that $i$th indexed element is a set of states from which there exist a path in the product automaton of length $i$ whose last transition is an accepting one.
+In lines 7-13, it tries to sample a transition.
+Initially it will try to sample a transition from states which are at distance 1 from the accepting transitions and then it will increase the length with each iteration.
+The procedure $SAMPLE$ on line 10 will sample transitions until it finds a transition from $fromStates$ to $toStates$ or it exceeds a given threshold time/number.
+If it is still not able to find any "good" transition then just sample randomly from any one of the state in $currentStates$ (line 14-15).
+Suppose it samples $\{r_1\}\rightarrow \{H\}$ next.
+Line 16 adds post of the sampled transition i.e $\{H\}$ into $currentSates$.
+Line 17 adds the sampled transition of level 3 in the product automaton and a few other "similar" transitions of level 2.
+For our example the product automaton after 3rd iteration look like figure \ref{fig:product3}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{Fig/product3} \caption{Product automaton after 3 iterations}
+	\label{fig:product3}
+\end{figure}
 \end{document}

### 2020-05-16 learning Logic/draft.tex
a

@@
 \newcommand{\jana}[1]{{\color{magenta} #1}}
+\newcommand{\kush}[1]{{\color{blue} #1}}
 
@@
 
+For the example given in Figure \ref{fig:office}, without considering $wall$ atomic proposition, it was taking approx.
+15 seconds to do everything.
+There is some room for improvement and it might get reduced to approx.
+10 seconds.
+
 \begin{figure}
@@
 \end{figure}
+
+\vspace{20pt}
+\textbf{Points from the discussion on 13.05.2020:}
+	\begin{itemize}
+		\item We assume that the robot has the labelling of the whole system beforehand for now. We can consider the other case later.
+		\item Use RRG instead of RRT so that we can go back also.
+		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (a formula over the APs), SAMPLE should return a newly sampled transition from a state where 'fromState' is \textit{true} to a state where 'toState' is \textit{true}.
+It maybe the case that such a transition doesn't exist or probability of it being sampled is very small, in that case, return \emph{null} after some given threshold time/samples.
+		\item \textbf{Jan's idea:} We can keep track of the sampled transitions till now and get their ranking (finding ranking for each sampled transition will require some time because it might have to pass through the whole list of 'reachableStates').
+Sample 50 (some threshold) transitions and add the one with the best rank.
+\kush{What will happen to the other transitions?
+Would we keep all of them in the memory (maintain a min-heap maybe)?}
+	\end{itemize}
+
 \end{document}

### 2020-05-16 MotionPlanning/src/Planning.java

@@ -40,8 +40,6 @@ public class Planning{
         productAutomaton.setInitState(productAutomaton.getInitStates().and(initStateSystem));
         BDD currentStates=initStateSystem.id();
         
-//        productAutomaton.getBDD().andWith(exper.addFilters());
-        
         double timeForSampling=0, preTimeSampling=0, postTimeSampling=0, pathTime=0;
         int iterationNumber=0;
         while(true) { //until the property is satisfied
@@ -83,22 +81,22 @@ public class Planning{
         		System.out.println("Yay!!!!!");
         		break;
         	}
-        	initialize.getProductAutomaton().createDot(iterationNumber);
+//        	initialize.getProductAutomaton().createDot(iterationNumber);
         	iterationNumber++;
         }
         
         factory.done();
 		
-		
-		
-		double endTime = System.nanoTime();
+        double endTime = System.nanoTime();
+        
+		System.out.println("\nTotal sampled transitions = "+iterationNumber);
 		System.out.print("\n\nTotal time taken (in ms):");
         System.out.println((endTime-startTime)/1000000);
         System.out.print("Time taken sampling (in ms):");
         System.out.println(timeForSampling/1000000);
         System.out.print("Time taken other than sampling (in ms):");
         System.out.println((endTime-startTime-timeForSampling)/1000000);
-        System.out.print("Path Time (in ms):");
+        System.out.print("Path checking time (in ms):");
         System.out.println(pathTime/1000000);
 
 	}

### 2020-05-16 MotionPlanning/src/modules/DefaultExperiment.java

@@ -19,22 +19,31 @@ public class DefaultExperiment implements Experiments{
 	ProductAutomaton productAutomaton;
 	BDDFactory factory;
 	BDD productAutomatonBDD;
-	public DefaultExperiment(BDD propertyBDD, 
-			int[] numVars, 
-			ArrayList<String> apListProperty,
-			ArrayList<String> apListSystem) throws Exception
-	{
-		this.productAutomaton=new ProductAutomaton(propertyBDD);
-		this.factory=productAutomaton.getBDD().getFactory();
-		this.productAutomatonBDD=productAutomaton.getBDD();
-	}
 	
+
+//	public DefaultExperiment(BDD propertyBDD, 
+//			int[] numVars, 
+//			ArrayList<String> apListProperty,
+//			ArrayList<String> apListSystem) throws Exception
+//	{
+//		this.productAutomaton=new ProductAutomaton(propertyBDD);
+//		this.factory=productAutomaton.getBDD().getFactory();
+//		this.productAutomatonBDD=productAutomaton.getBDD();
+//	}
+	
+	/**
+	 * constructor
+	 * @param productAutomaton
+	 */
 	public DefaultExperiment(ProductAutomaton productAutomaton) {
 		this.productAutomaton=productAutomaton;
 		this.factory=productAutomaton.getBDD().getFactory();
 		this.productAutomatonBDD=productAutomaton.getBDD();
 	}
 
+	/**
+	 * LEARN procedure
+	 */
 	@Override
 	public BDD learn(BDD fromState, BDD toState) throws Exception {
 		BDD transition=fromState.and(productAutomaton.changePreSystemVarsToPostSystemVars(toState));
@@ -49,7 +58,7 @@ public class DefaultExperiment implements Experiments{
 		}
 		
 		learnSimilarTransitions(fromState, toState);
-		
+		  
 		
 		int counter=productAutomaton.increaseCounter(productAutomaton.getFirstState(fromState));
 		
@@ -62,6 +71,13 @@ public class DefaultExperiment implements Experiments{
 		return transition;
 	}
 
+	/**
+	 * procedure to add maybe (level 2) transitions give a transition from 'fromState' to 'toState'
+	 * @param fromState
+	 * @param toState
+	 * @return BDD representing the set of similar transitions
+	 * @throws Exception
+	 */
 	private BDD learnSimilarTransitions(BDD fromState, BDD toState) throws Exception{
[…]

### 2020-05-16 MotionPlanning/src/printing/PrintProductAutomaton.java

@@ -17,7 +17,7 @@ public class PrintProductAutomaton
 		writer = new BufferedWriter(new FileWriter("/home/kush/Projects/robotmotionplanning/MotionPlanning/temp/productAutomaton"+num+".dot"));
 		
 		this.productAutomaton=productAutomaton;
-		BDD productAutomatonBDD=productAutomaton.getBDD().and(ProductAutomaton.transitionLevelDomain().ithVar(3));
+		BDD productAutomatonBDD=productAutomaton.getBDD();
 		
 		int[] numVars=ProductAutomaton.numVars;
 		int totalVars=0;
@@ -63,15 +63,15 @@ public class PrintProductAutomaton
 		if(level==3) {
 			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, color=green, label=\""+ labelString+" ("+acceptingSet+")\"];\n");
 		}
-//		else if(level==2) {
-//			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, color=blue, label=\""+ labelString+" ("+acceptingSet+")\"];\n");
-//		}
-//		else if(level==1) {
-//			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, color=red, label=\""+ labelString+" ("+acceptingSet+")\"];\n");;
-//		}
-//		else if(level==0) {
-//			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, ,color=black, label=\""+ labelString+" ("+acceptingSet+")\"];\n");
-//		}
+		else if(level==2) {
+			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, color=blue, label=\""+ labelString+" ("+acceptingSet+")\"];\n");
+		}
+		else if(level==1) {
+			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, color=red, label=\""+ labelString+" ("+acceptingSet+")\"];\n");;
+		}
+		else if(level==0) {
+			writer.append(fromStateID+" -> "+ toStateID+" [style=filled, ,color=black, label=\""+ labelString+" ("+acceptingSet+")\"];\n");
+		}
 	}
 	
 	private String getTransitionLabelString(BDD transition) throws Exception{

### 2020-05-16 MotionPlanning/src/transitionSystem/ProductAutomaton.java

@@ -164,8 +164,16 @@ public class ProductAutomaton{
 	
 //------------------------------------------------------------------------------------------------------------------	
 /**
- * For getters
+ * For getters and setters
  */
+	
+	/**
+	 * set the initial state
+	 * @param initStates
+	 */
+	public void setInitState(BDD initStates) {
+		this.initStates=initStates;
+	}
 	
 	/**
 	 * 
@@ -175,30 +183,45 @@ public class ProductAutomaton{
 		return productAutomatonBDD;
 	}
 	
+	/**
+	 * 
+	 * @return BDD representing the property automaton
+	 */
 	public static BDD getPropertyBDD() {
 		return propertyBDD;
 	}
 	
+	/**
+	 * 
+	 * @return set of initial states
+	 */
 	public BDD getInitStates() {
 		return initStates;
 	}
 	
+	/**
+	 * 
+	 * @return total number of states in the product automaton
+	 */
 	public static int getNumStates() {
 		return numStates;
-	}
-	
-	public BDD getSampledProductTransitions() {
-		return sampledProductTransitions;
 	}
 	
 	/**
 	 * 
 	 * @return number of states in the product automaton
 	 */
-	public static int numberOfStates(){
+	private static int numberOfStates(){
 		return (int) Math.pow(2, numAPSystem+propertyDomainPre().varNum());
 	}
 	
+	/**
+	 * 
+	 * @return all the sampled transitions in the product automaton
+	 */
+	public BDD getSampledProductTransitions() {
+		return sampledProductTransitions;
+	}
 	
 	/**
 	 * 
@@ -643,6 +666,16 @@ public class ProductAutomaton{
 		return changePostVarsToPreVars(removeAllExceptPostVars(productAutomatonBDD.and(states)));
 	}
 	
+	/**
+	 * 
+	 * @param i
+	 * @return first states of all accepting transitions of ith accepting set
+	 * @throws Exception
+	 */
+	public BDD getAcceptingStates(int i) throws Exception {
+		return removeAllExceptPreVars(productAutomatonBDD.and(acceptingSetDomain().ithVar(i)));
+	}
+	
 	
 //-------------------------------------------------------------------------------------------------------------------
 
@@ -813,20 +846,9 @@ public class ProductAutomaton{
 	public int getLevel(BDD transition) throws TransitionException {
 		if(transition.pathCount()>1) {
[…]

### 2020-05-16 learning Logic/draft.tex
a

@@
 		\item We assume that the robot has the labelling of the whole system beforehand for now. We can consider the other case later.
-		\item Use RRG instead of RRT so that we can go back also.
-		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (a formula over the APs), SAMPLE should return a newly sampled transition from a state where 'fromState' is \textit{true} to a state where 'toState' is \textit{true}.
-It maybe the case that such a transition doesn't exist or probability of it being sampled is very small, in that case, return \emph{null} after some given threshold time/samples.
+		\item Use RRG instead of RRT so that we can go back also. But in RRG, multiple transitions can be added with each sampled state. \kush{Will have to change the implementation a bit.}
+		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (formulas over the APs of the system), SAMPLE should return a newly sampled transition from a state where 'fromState' is \textit{true} to a state where 'toState' is \textit{true}.
+It maybe the case that such a transition doesn't exist or probability of it being sampled is very small, in that case, return \emph{null} after some given threshold time/samples.\newline
+		\textbf{How do we do this with RRG?}
+		\begin{itemize}
+			\item Each node in the graph ($G$) will store the APs associated with it. 
+			\item If the whole tree is stored as a BDD, finding nodes where 'fromStates' is very easy but storing it in a BDD is very difficult. \kush{Find a way to do both the things efficiently.}
+			\item 
+		\end{itemize}
+
 		\item \textbf{Jan's idea:} We can keep track of the sampled transitions till now and get their ranking (finding ranking for each sampled transition will require some time because it might have to pass through the whole list of 'reachableStates').

### 2020-05-16 learning Logic/draft.tex
a

@@
 		\item We assume that the robot has the labelling of the whole system beforehand for now. We can consider the other case later.
-		\item Use RRG instead of RRT so that we can go back also. But in RRG, multiple transitions can be added with each sampled state. \kush{Will have to change the implementation a bit.}
-		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (formulas over the APs of the system), SAMPLE should return a newly sampled transition from a state where 'fromState' is \textit{true} to a state where 'toState' is \textit{true}.
+
+		\item Use RRG instead of RRT so that we can go back also.
+
+		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (formulas over the APs of the system), SAMPLE should return a newly sampled transition from a state where `fromState' is \textit{true} to a state where `toState' is \textit{true}.
 It maybe the case that such a transition doesn't exist or probability of it being sampled is very small, in that case, return \emph{null} after some given threshold time/samples.\newline
@@
 			\item Each node in the graph ($G$) will store the APs associated with it. 
-			\item If the whole tree is stored as a BDD, finding nodes where 'fromStates' is very easy but storing it in a BDD is very difficult. \kush{Find a way to do both the things efficiently.}
-			\item 
+			\item If the whole tree is stored as a BDD, finding nodes where `fromStates' is very easy but storing it in a BDD is very difficult. \kush{Find a way to do both the things efficiently.}
 		\end{itemize}
 
-		\item \textbf{Jan's idea:} We can keep track of the sampled transitions till now and get their ranking (finding ranking for each sampled transition will require some time because it might have to pass through the whole list of 'reachableStates').
+		\item \textbf{Jan's idea:} We can keep track of the sampled transitions till now and get their ranking (finding ranking for each sampled transition will require some time because it might have to pass through the whole list of `reachableStates').
 Sample 50 (some threshold) transitions and add the one with the best rank.

### 2020-09-03 learning Logic/draft.tex
a

@@
 
+\subsection{LEARN and ASK procedures}
+
+\textit{LEARN:} Learn procedure takes a transition $t=(s_1, s_2)$ (or a set of transitions) and add similar `maybe' transitions to the product automaton.
+For defining these similar transitions we first need a set of atomic propositions called domain of changes, which is defined as $\{a \oplus  b \mid a\in s_1, b\in _2\}$.
+A transition similar to $t$ would be $(s_1', s_2')$ where $s_1'\ (s_2')$ is a state which agrees with $s_1\ (s_2)$ on domain of changes.
+For e.g, if $t = (\{r_1\}, \{r_1, b\})$ then, the set $DOC=\{b_1\}$ and one of the learned transition would be $(\{r_2\},\{r_2, b\})$.
+If we have a different atomic proposition for all the bins and $t = (\{r_1\}, \{r_1, b_1\})$ then the learned transition will be $(\{r_2\}, \{r_2,b_1\})$ [not $(\{r_2\}, \{r_2,b_2\})$].
+
+\textit{ASK:} Ask procedure takes the product automaton and returns a list of set of states called $reachableStates$ whose $i$th element is a set of states which can see an accepting transition in the product automaton in $\leq i$ steps.
+Look at the example in figure \ref{fig:ask}, transition with label $(*)$ is an accepting transition.
+$reachableStates[0]=\{s_6\}$, [1]=$\{s_5, s_6\}$, [2]=$\{s_3,s_5,s_6\}$ and [3]=$\{s_0,s_1,s_2,s_3,s_5,s_6\}$
+
+\begin{figure}
+	\begin{tikzpicture}[
+	roundnode/.style={circle, draw=black, fill=black!5, very thick, minimum size=3mm}]
+
+	\node[roundnode] (A) {$s_0$};
+	\node[roundnode] (B) [below=of A] {$s_1$};
+	\node[roundnode] (C) [below=of B] {$s_2$};
+	\node[roundnode] (D) [right=of A] {$s_3$};
+	\node[roundnode] (E) [below=of D] {$s_4$};
+	\node[roundnode] (F) [right=of D] {$s_5$};
+	\node[roundnode] (G) [right=of F] {$s_6$};
+
+	\draw[->] (A) -- (D);
+	\draw[->] (B) -- (D);
+	\draw[->] (C) -- (E);
+	\draw[->] (C) -- (D);
+	\draw[->] (D) -- (F);
+	\draw[->] (F) -- (G) node[anchor = south, pos = 0.5] {$(*)$};
+	\end{tikzpicture}
+	\caption{A part of the product automaton}
+	\label{fig:ask}
+\end{figure}
+
 \subsection{A mini case study}
@@
 
+\section{Fernando's Ideas}
+Suppose the system knows the environment, it's free space and obstacles.
+I.e.
+$X_{\text{obs}}, X_{\text{free}}$ are known.
+Let's also say that the set of $k$ atomic propositions $Ap$, denoted as $\Sigma = \{Ap_1, Ap_2, \dots, Ap_k\}$, is known.
+However, let's suppose the labeling function $L: X \to 2^\Sigma$ is partially known, and constructed as the robot navigates the environment.
+For example, it can attribute atomic propositions to regions of the workspace when within $\epsilon$-distance to it.
+
+I think approaching it this way would make our contribution richer.
+If we assume to know $L$, we actually don't need to know anything else about the workspace besides the positions of $Ap$s in question, for example the bins.
+No spatial relation is necessary.
+Now, if $L$ is not fully known, the problem becomes more interesting.
+Our system will then try to learn and predict characteristics of the workspace as it moves through it.
+
+Also, what comes to my mind is regarding the mission formula and how much information we feed to it.
+In $G F (r_1 \wedge b) \wedge G F (r_2\wedge b)$ we are already giving out that there exists a bin in each room.
+If one of the rooms does not have a bin, the formula is not satisfiable anymore.
+If the goal is to collect all bins, the formula should somehow contain only bins.
+The relation of bins being inside rooms should be learned by the model.
+
+Now, how to make sure the robot does not visit the same bin infinitely often? Should we know beforehand the number of bins and write a formula like $\bigwedge_{i \in [1,n]} GF b_i$?
+
+This raises another question: is there any grouping of $Ap$s?
+How does the algorithm knows that $r_1$ and $r_2$ are both rooms, and not a room and a lake?
+Or the inference that a transition from $r_1$ to $b_1$ can mean that there might exist a transition from $r_2$ to $b_2$, but not from $h$ to $b_3$?
+
 \end{document}

### 2020-09-20 MotionPlanning/src/Planning.java
v 1.0

@@ -8,6 +8,7 @@ import modules.learnAskExperiments.DefaultExperiment;
 import java.io.OutputStream;
 import java.io.PrintStream;
 import java.util.ArrayList;
+import java.util.Arrays;
 import java.util.List;
 
 import org.jgrapht.graph.DefaultEdge;
@@ -59,20 +60,24 @@ public class Planning
         
         // loop of the main algo starts here
         double samplingTime = 0, preTimeSampling = 0, pathStartTime = 0, pathTime = 0, learnStartTime = 0, learnTime = 0;
+        int[] advice = new int[50];
+        for(int k=0;k<50;k++) {
+        	advice[k]=0;
+        }
         int iterationNumber					= 0;
         boolean computePath 				= false;
         List<DefaultEdge> finalPath = new ArrayList<DefaultEdge>();
-        PrintStream out = System.out;		// Don't output random things from libraries
+//        PrintStream out = System.out;		// Don't output random things from libraries
         
         while(true)  //until the property is satisfied
         {
 
         	// Don't output random things
-        	System.setOut(new PrintStream(OutputStream.nullOutputStream()));
+//        	System.setOut(new PrintStream(OutputStream.nullOutputStream()));
         	
         	
         	// Sample transitions
-        	ArrayList<BDD> reachableStates	= exper.ask(currentStates);
+        	ArrayList<BDD> reachableStates	= exper.advice(currentStates);
         	int currentPathLength			= 1;
         	
         	BDD transition = null, fromStates = null, toStates = null;
@@ -88,21 +93,18 @@ public class Planning
         		samplingTime				+= System.nanoTime() - preTimeSampling;
         		if(transition == null) {
         			currentPathLength++;
+        			continue;
+        		}
+        		else {
+        			advice[currentPathLength-1]++;
         			break;
         		}
-        	}	
-        	
-        	if(transition == null) // sample from the set of current states
-        	{
-        		preTimeSampling 			= System.nanoTime();
-        		transition					= rrg.sample(currentStates, productAutomaton);
-        		samplingTime				+= System.nanoTime() - preTimeSampling;
         	}
         	
         	if(transition == null) // sample anywhere
         	{
         		preTimeSampling 			= System.nanoTime();
-        		transition					= rrg.sample(productAutomaton);
+        		transition					= rrg.sampleRandomly(productAutomaton);
         		samplingTime				+= System.nanoTime() - preTimeSampling;
         	}
         	
[…]

### 2020-09-20 MotionPlanning/src/modules/RRG.java

@@ -92,14 +92,8 @@ public class RRG
 	 * @param productAutomaton
 	 * @return
 	 */
-	public BDD buildGraph(BDD fromStates, BDD toStates, Point2D xRand2D, ProductAutomaton productAutomaton) 
+	public BDD buildGraph(BDD fromStates, BDD toStates, Point2D xRand2D, ProductAutomaton productAutomaton, int type) 
 	{
-//		if(numPoints == 0) {
-//			setStartingPoint(xRand2D);
-//			numPoints++;
-//		}
-		
-		
 		
 		BDD transitions 			= ProductAutomaton.factory.zero();
 		
@@ -174,8 +168,12 @@ public class RRG
 					
 					BDD transition;
 					try {
-						transition = productAutomaton.changePreSystemVarsToPostSystemVars(Environment.getLabelling().getLabel(xNew2D));
+						transition		= productAutomaton.changePreSystemVarsToPostSystemVars(Environment.getLabelling().getLabel(xNew2D));
 						transition 		= transition.and(Environment.getLabelling().getLabel(xNearest2D));
+						if(type == 1 && transition.and(productAutomaton.getBDD()).and(fromStates).and(productAutomaton.changePreVarsToPostVars(toStates)).and(ProductAutomaton.getPropertyBDD()).and(ProductAutomaton.getLabelEquivalence()).isZero()) {
+							System.out.println("ajhdkjajaskfaj");
+							return false;
+						}
 						transitions.orWith(transition);
 					} catch (Exception e1) {
 						e1.printStackTrace();
@@ -286,7 +284,7 @@ public class RRG
 		} else 
 		{
 			Point2D temp = new Point2D.Float((float) (source.getX() + ((eta-0.00001)*(dest.getX()-source.getX())/d)), (float) (source.getY()+((eta-0.00001)*(dest.getY()-source.getY())/d)));
-			if((float) source.distance(temp)>0.051f) {
+			if((float) source.distance(temp)>1.0f) {
 				System.out.println("I am fucked");
 			}
 			return temp;
@@ -313,7 +311,7 @@ public class RRG
 		{
 			p 			= env.sample();
 			totalSampledPoints++;
-			transition 	= buildGraph(fromStates, toStates, p, productAutomaton);
+			transition 	= buildGraph(fromStates, toStates, p, productAutomaton, 1);
 			if(! transition.isZero())
 			{
 				return transition;
@@ -329,39 +327,38 @@ public class RRG
[…]

### 2020-09-20 MotionPlanning/src/modules/learnAskExperiments/DefaultExperiment.java

@@ -150,26 +150,26 @@ public class DefaultExperiment implements Experiments
 	 * ASK procedure
 	 */
 	@Override
-	public ArrayList<BDD> ask(BDD currentStates) throws Exception
+	public ArrayList<BDD> advice(BDD currentStates) throws Exception
 	{
 		ArrayList<BDD> reachableStates	= new ArrayList<BDD>();
 //		
-//		reachableStates.add(productAutomaton.finalStates());
-//		reachableStates.add(productAutomaton.preImageOfFinalStates());
+		reachableStates.add(productAutomaton.finalStates());
+		reachableStates.add(productAutomaton.preImageOfFinalStates());
+		
+		if(reachableStates.get(0).isZero() || reachableStates.get(1).isZero()) {
+			return reachableStates;
+		}
+		
+		int i	= 1;
+		BDD backwardReachableStates		= productAutomaton.preImageOfFinalStates();
+		while(!productAutomaton.preImage(backwardReachableStates).and(backwardReachableStates.not()).isZero()) {
+			reachableStates.add(productAutomaton.preImage(reachableStates.get(i)));
+			i++;
+			backwardReachableStates		= backwardReachableStates.or(reachableStates.get(i));
+		}
 //		
-//		if(reachableStates.get(0).isZero() || reachableStates.get(1).isZero()) {
-//			return reachableStates;
-//		}
-//		
-//		int i	= 1;
-//		BDD backwardReachableStates		= productAutomaton.preImageOfFinalStates();
-//		while(!productAutomaton.preImage(backwardReachableStates).and(backwardReachableStates.not()).isZero()) {
-//			reachableStates.add(productAutomaton.preImage(reachableStates.get(i)));
-//			i++;
-//			backwardReachableStates		= backwardReachableStates.or(reachableStates.get(i));
-//		}
-//		
-		reachableStates.add(ProductAutomaton.factory.one());
+//		reachableStates.add(ProductAutomaton.factory.one());
 		return reachableStates;
 		
 	}

### 2020-09-20 MotionPlanning/src/modules/learnAskExperiments/Experiments.java

@@ -11,7 +11,7 @@ public interface Experiments
 
 	public BDD learn(BDD fromState, BDD toState) throws Exception;
 	
-	public ArrayList<BDD> ask(BDD startingStates) throws Exception;
+	public ArrayList<BDD> advice(BDD startingStates) throws Exception;
 	
 	public ProductAutomaton getProductAutomaton();
 }

### 2020-09-21 Presentation/document.tex
Upload New File

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usetheme{Copenhagen}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}

\newcommand{\xmark}{\ding{55}}%
\newcommand{\cmark}{\ding{51}}

\title[Robot Motion Planning]{Robot Motion Planning}
\author{F. Barbosa, K. Grover, J. K\v{r}et\'{i}nsk\'{y}, J. Tumova}
%\author{Presented by Kush}
\date{}
\begin{document}
	\frame{\titlepage}

	\begin{frame}
		\frametitle{Motion Planning Problem}
		\begin{center}
			\begin{tikzpicture}
				\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
				\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
				\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
				\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
				\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
				\filldraw[black] (0.2, 3) circle (0.4mm);
			\end{tikzpicture}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{RRG}
		\begin{center}
			\begin{tikzpicture}
			\onslide<0->{
				\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
				\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
				\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
				\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
				\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
				\filldraw[black] (0.2, 3) circle (0.4mm);
			}
			\onslide<1-2>{
				\filldraw[red] (1.8, 3.8) circle (0.4mm);
			}
			\onslide<2->{
				\filldraw[black] (0.6, 3.2) circle (0.4mm);
				\draw (0.2,3) -- (0.6, 3.2);
			}
			\onslide<3->{
				\filldraw[black] (0.45, 2.6) circle (0.4mm);
				\draw (0.2,3) -- (0.45, 2.6);
				\filldraw[black] (0.8, 3.5) circle (0.4mm);
				\draw (0.6,3.2) -- (0.8, 3.5);
				\filldraw[black] (0.9, 2.9) circle (0.4mm);
				\draw (0.6,3.2) -- (0.9, 2.9);
			}
			\onslide<4>{
				\filldraw[red] (0.8, 2.5) circle (0.4mm);
			}
			\onslide<5->{
				\filldraw[black] (0.8, 2.5) circle (0.4mm);
				\draw (0.45, 2.6) -- (0.8, 2.5);
				\draw (0.9, 2.9) -- (0.8, 2.5);
			}
			\end{tikzpicture}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Jan's vision}
		\begin{itemize}
		\item Want a robot that can move around on its own and figure out the stuff it has to do. \pause \textcolor{red}{\xmark}\\
		Too big a goal right now from both theoretical and practical perspectives.\\
		\pause \vspace{10pt}
		What can we do?\\
		Reachability. \textcolor{green}{\cmark} \pause\\
		LTL or a subclass of LTL. \\
		Given a labeling of the environment. Find a path that satisfies a given LTL formula.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{LTL Motion Planning}
		\begin{center}
			\begin{tikzpicture}
			\filldraw[dashed, fill = gray!10!white] (0.25,0.25) -- (1.55,0.25) -- (1.55,1.15) -- (0.25,1.15) -- (0.25, 0.25);
			\filldraw[dashed, fill = green!10!white] (0,0) -- (0.35,0) -- (0.35,0.35) -- (0,0.35) -- (0,0);
			\filldraw[dashed, fill = gray!10!white] (1.85,4.45) -- (2.75,4.45) -- (2.75,5.75) -- (1.85,5.75) -- (1.85, 4.45);
			\filldraw[dashed, fill = green!10!white] (2.65,5.65) -- (3,5.65) -- (3,6) -- (2.65,6) -- (2.65,5.65);

			\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
			\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
			\filldraw[fill = green!40!white, draw = black] (2.8, 5.8) rectangle (3, 6);
			\filldraw[fill = black!10!white, draw = black] (2, 4.6) rectangle (2.6, 5.6);

			\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
			\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
			\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
			\filldraw[black] (0.2, 3) circle (0.4mm);
			\end{tikzpicture}

			\emph{Specification:} $G F (r_1 \wedge b) \wedge G F (r_2 \wedge b)$
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Solution}
		\begin{itemize}
			\item Build an abstraction of the system.
			\item Construct the product automaton with the property automaton (ldba).
			\item Find a path in the abstraction.
			\item Lift this path to the original system.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Better Solution}
		Can we do better?\\ \pause \vspace{10pt}
		\textbf{Jan's idea:} Get help in sampling from current abstraction. \\
		Depending on previous experiences. In other words, predict which samples would be better from previous samples. \\ \pause \vspace{10pt}
		For e.g: The bin was close to the table in first room so try to look close to the table in the other room to find the bin.
	\end{frame}

	\begin{frame}
		\frametitle{Better Solution}
		How do we formalize this? \\ \pause
		\begin{itemize}
			\item Have `maybe' and `maybe not' transitions in the abstraction.\\
			\item Whenever you see a transition, add similar `maybe' transitions in the abstraction as well.
			\item What does similar mean here?\\
			$(r_1, t) \rightarrow (r_1, b) \implies (r_2, t) \rightarrow (r_2,b)$. \\
			\begin{itemize}
				\item Compute domain of changes: Set of APs changing in a transition.\\
				$DOC = \{t,b\}$, $(s_1 \oplus s_2)$
				\item Add transitions $s_1' \rightarrow s_2'$ where $s_1'$ and $s_2'$ are states which agree with $s_1$ and $s_2$ on $DOC$ respectively.
			\end{itemize}
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Algorithm}
		\begin{center}
			\begin{tikzpicture}
			\filldraw[dashed, fill = gray!10!white] (0.25,0.25) -- (1.55,0.25) -- (1.55,1.15) -- (0.25,1.15) -- (0.25, 0.25);
			\filldraw[dashed, fill = green!10!white] (0,0) -- (0.35,0) -- (0.35,0.35) -- (0,0.35) -- (0,0);
			\filldraw[dashed, fill = gray!10!white] (1.85,4.45) -- (2.75,4.45) -- (2.75,5.75) -- (1.85,5.75) -- (1.85, 4.45);
			\filldraw[dashed, fill = green!10!white] (2.65,5.65) -- (3,5.65) -- (3,6) -- (2.65,6) -- (2.65,5.65);

			\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
			\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
			\filldraw[fill = green!40!white, draw = black] (2.8, 5.8) rectangle (3, 6);
			\filldraw[fill = black!10!white, draw = black] (2, 4.6) rectangle (2.6, 5.6);

			\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
			\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
			\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
			\filldraw[black] (0.2, 3) circle (0.4mm);

			\onslide<2->{
				\node[state,initial,initial text=,scale=0.8] (q1) at (5,5) {$0$};
				\node[state,scale=0.8] (q2) at (8,5) {$1$};

				\draw (q1) edge[loop above] node{$r_1 \wedge r_2 \wedge b\ \ast$} (q1);
				\draw (q1) edge[loop below] node{$(r_1 \wedge \neg b) \vee \neg r_1$} (q1);
				\draw (q1) edge[bend left, above, ->] node{$r_1 \wedge \neg r_2 \wedge b$} (q2);
				\draw (q2) edge[loop above] node{$(r_2 \wedge b) \vee \neg r_1$} (q2);
				\draw (q2) edge[bend left, below, ->] node{$r_2 \wedge b\ \ast$} (q1);
			}

			\onslide<2->{
				\filldraw[black] (0.6, 3.2) circle (0.4mm);
				\draw (0.2,3) -- (0.6, 3.2);
				\filldraw[black] (0.45, 2.6) circle (0.4mm);
				\draw (0.2,3) -- (0.45, 2.6);
				\filldraw[black] (0.8, 3.5) circle (0.4mm);
				\draw (0.6,3.2) -- (0.8, 3.5);
				\filldraw[black] (0.9, 2.9) circle (0.4mm);
				\draw (0.6,3.2) -- (0.9, 2.9);
				\filldraw[black] (0.8, 2.5) circle (0.4mm);
				\draw (0.45, 2.6) -- (0.8, 2.5);
				\draw (0.9, 2.9) -- (0.8, 2.5);
				\filldraw[black] (1.1, 3.8) circle (0.4mm);
				\draw (0.8, 3.5) -- (1.1, 3.8);
				\filldraw[black] (1.4, 3.9) circle (0.4mm);
				\draw (1.1, 3.8) -- (1.4, 3.9);
			}

			\onslide<3->{
				\filldraw[black] (1.3, 4.3) circle (0.4mm);
				\draw (1.4, 3.9) -- (1.3, 4.3);

				\node[state,initial,initial text=] (0h) at (5,1) {$0,h$};
			}
			\onslide<4->{
				\node[state] (0r1) at (8,1) {$0,r_1$};
				\draw (0h) edge[color = green, bend left, above, ->] node{$(r_1 \wedge \neg b) \vee \neg r_1$} (0r1);
			}
			\end{tikzpicture}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Learn}
		\begin{itemize}
			\item Sample a transition $s_1$ to $s_2$. 	\pause
			\item Add similar transitions.	\pause
			\item If $s_1$ has been visited 50 times, all the `maybe' transitions gets converted to `maybe not' transitions.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Ask}
			\begin{itemize}
				\item Maintain a set of states in the abstraction which we have visited till now, \emph{visitedStates}. \pause
				\item Look at the accepting `maybe' transitions and get a list of sets of states $reach[n]$, such that $reach[i]$ will have all the states in \emph{visitedStates} that can reach an accepting transition in $i$ steps. First try to sample transitions from $reach[1] \rightarrow reach[0]$, then from $reach[2] \rightarrow reach[1]$ and so on.
			\end{itemize}

	\end{frame}

	\begin{frame}
		\frametitle{What's next?}
		\begin{itemize}
			\item `Definitely not' transitions. \pause
			\item Can we do full LTL? or LTL $\setminus \{X\}$. \pause
			\item What kind of filters can we use to learn useful similar transitions? \pause
			\item Make implementation faster. \pause
			\item Figure out the best thresholds. 	\pause
			\item Still very hard because we have to compete with very efficient algorithm.
		\end{itemize}

	\end{frame}

	\begin{frame}
		\frametitle{After this project}
		\begin{itemize}
			\item A robot that does not have information of the whole environment and can only look at things nearby. \pause
			\item Figure out atomic propositions as it goes on to make learning better (not the ones in the property).
		\end{itemize}
	\end{frame}

\end{document}

### 2020-09-21 MotionPlanning/src/Planning.java
Example with obstacles

@@ -26,8 +26,8 @@ public class Planning
 	public static void main(String[] args) throws Exception
 	{    	
     	
-		// Initialise everything-------------------------------------------
-		double startTime 					= System.nanoTime();
+		// Initialise everything <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+		double startTime					= System.nanoTime();
 		
 		// use default values for everything
 		new PlanningSettings();
@@ -52,60 +52,62 @@ public class Planning
         BDD currentStates					= initStateSystem.id(); //stores the set of states in the abstraction we have seen till now
         
         double initializationTime 			= System.nanoTime() - startTime;
-        //-------------------------------------------------------------------
+//      >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
         
         
-        
-        
-        
-        // loop of the main algo starts here
-        double samplingTime = 0, preTimeSampling = 0, pathStartTime = 0, pathTime = 0, learnStartTime = 0, learnTime = 0;
-        int[] advice = new int[50];
-        for(int k=0;k<50;k++) {
-        	advice[k]=0;
+        // for output of advice
+        int[] advice 						= new int[50];
+        for(int k=0 ;k<50; k++) 
+        {
+        	advice[k]						= 0;
         }
-        int iterationNumber					= 0;
-        boolean computePath 				= false;
-        List<DefaultEdge> finalPath = new ArrayList<DefaultEdge>();
-//        PrintStream out = System.out;		// Don't output random things from libraries
         
+        // for calculating time of different parts
+        double samplingTime = 0, pathTime = 0, learnStartTime = 0, learnTime = 0, processStartTime = 0, adviceTime = 0;
+        
+        int iterationNumber					= 0;
+    	BDD transition = null, fromStates = null, toStates = null;
+        boolean computePath 				= false;
+        List<DefaultEdge> finalPath 		= new ArrayList<DefaultEdge>();
+        boolean needNewAdvice = true;
+        ArrayList<BDD> reachableStates = null;
         while(true)  //until the property is satisfied
         {
 
         	// Don't output random things
 //        	System.setOut(new PrintStream(OutputStream.nullOutputStream()));
         	
[…]

### 2020-09-21 MotionPlanning/src/abstraction/ProductAutomaton.java

@@ -7,11 +7,10 @@ import net.sf.javabdd.BDDFactory;
 import net.sf.javabdd.BDDPairing;
 import planningIO.printing.PrintAcceptingPath;
 import planningIO.printing.PrintProductAutomaton;
+import settings.PlanningException;
 
 import java.util.ArrayList;
 
-import abstraction.exceptions.StateException;
-import abstraction.exceptions.TransitionException;
 import modules.emptinessCheck.EmptinessCheck;
 
 /**
@@ -48,9 +47,9 @@ public class ProductAutomaton
 	/**
 	 * <p>Initializes Product Automaton to an empty automaton</p>
 	 * @param propertyBDD BDD for the property automaton
-	 * @throws Exception 
+	 * @throws PlanningException 
 	 */
-	public ProductAutomaton(BDD propertyBDD) throws Exception
+	public ProductAutomaton(BDD propertyBDD) throws PlanningException
 	{
 		ProductAutomaton.propertyBDD	= propertyBDD;
 		this.productAutomatonBDD		= factory.zero();
@@ -137,9 +136,9 @@ public class ProductAutomaton
 	/**
 	 * <p>Creates BDDPairing for changing pre system vars to label vars</p>
 	 * @return BDDPairing for changing pre system vars to label vars
-	 * @throws Exception 
+	 * @throws PlanningException 
 	 */
-	public static BDDPairing statesToLabelPairing() throws Exception 
+	public static BDDPairing statesToLabelPairing() throws PlanningException 
 	{
 		BDDPairing newPairing 	= factory.makePair();
 		for(int i=0; i<numAPSystem; i++) 
@@ -294,13 +293,13 @@ public class ProductAutomaton
 	 * 
 	 * @param i
 	 * @return ith system var in pre vars
-	 * @throws Exception 
+	 * @throws PlanningException 
 	 */
-	public static BDD ithVarSystemPre(int i) throws Exception 
+	public static BDD ithVarSystemPre(int i) throws PlanningException 
 	{
 		if(i > numAPSystem) 
 		{
-			throw new Exception("Index out of bounds");
+			throw new PlanningException("Index out of bounds");
 		}
 		return factory.ithVar(varsBeforeSystemVars + i);
 	}
@@ -309,13 +308,13 @@ public class ProductAutomaton
 	 * 
 	 * @param i
 	 * @return ith system var in post vars
-	 * @throws Exception 
[…]

### 2020-09-21 MotionPlanning/src/abstraction/exceptions/StateException.java

@@ -1,40 +1,0 @@
-package abstraction.exceptions;
-
-public class StateException extends Exception
-{
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 1L;
-
-	public StateException()
-	{
-		// TODO Auto-generated constructor stub
-	}
-
-	public StateException(String message)
-	{
-		super(message);
-		// TODO Auto-generated constructor stub
-	}
-
-	public StateException(Throwable cause)
-	{
-		super(cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public StateException(String message, Throwable cause)
-	{
-		super(message, cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public StateException(String message, Throwable cause, boolean enableSuppression, boolean writableStackTrace)
-	{
-		super(message, cause, enableSuppression, writableStackTrace);
-		// TODO Auto-generated constructor stub
-	}
-
-}

### 2020-09-21 MotionPlanning/src/abstraction/exceptions/TransitionException.java

@@ -1,40 +1,0 @@
-package abstraction.exceptions;
-
-public class TransitionException extends Exception
-{
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 1L;
-
-	public TransitionException()
-	{
-		// TODO Auto-generated constructor stub
-	}
-
-	public TransitionException(String message)
-	{
-		super(message);
-		// TODO Auto-generated constructor stub
-	}
-
-	public TransitionException(Throwable cause)
-	{
-		super(cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public TransitionException(String message, Throwable cause)
-	{
-		super(message, cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public TransitionException(String message, Throwable cause, boolean enableSuppression, boolean writableStackTrace)
-	{
-		super(message, cause, enableSuppression, writableStackTrace);
-		// TODO Auto-generated constructor stub
-	}
-
-}

### 2020-09-21 MotionPlanning/src/abstraction/exceptions/learningException.java

@@ -1,40 +1,0 @@
-package abstraction.exceptions;
-
-public class learningException extends Exception
-{
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 1L;
-
-	public learningException()
-	{
-		// TODO Auto-generated constructor stub
-	}
-
-	public learningException(String message)
-	{
-		super(message);
-		// TODO Auto-generated constructor stub
-	}
-
-	public learningException(Throwable cause)
-	{
-		super(cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public learningException(String message, Throwable cause)
-	{
-		super(message, cause);
-		// TODO Auto-generated constructor stub
-	}
-
-	public learningException(String message, Throwable cause, boolean enableSuppression, boolean writableStackTrace)
-	{
-		super(message, cause, enableSuppression, writableStackTrace);
-		// TODO Auto-generated constructor stub
-	}
-
-}

### 2020-09-21 MotionPlanning/src/abstraction/exceptions/package-info.java

@@ -1,1 +1,0 @@
-package abstraction.exceptions;

### 2020-09-21 MotionPlanning/src/environment/Label.java

@@ -50,4 +50,9 @@ public class Label
 	{
 		return apListSystem;
 	}
+	
+	public ArrayList<Path2D> getAreas()
+	{
+		return areas;
+	}
 }

### 2020-09-21 MotionPlanning/src/modules/RRG.java

@@ -137,19 +137,6 @@ public class RRG
 				// If the points are according to the advice
 				if(env.collisionFree(xNearest2D, xNew2D))	//check if it is collision free
 				{	
-//					System.out.println("Sampled Transition: " + xNearest2D.toString() + " ---> " + xNew2D.toString());
-//					try {
-//						if(! Environment.getLabelling().getLabel(xNearest2D).equals(Environment.getLabelling().getLabel(xNew2D))) {
-//							System.out.println(Environment.getLabelling().getLabel(xNearest2D).toString() + " ---> " + Environment.getLabelling().getLabel(xNew2D).toString());
-//						}
-//					} catch (Exception e1) {
-//						e1.printStackTrace();
-//					}
-					
-					
-					
-					
-					
 					final float radius;	// radius for which the neighbours will be considered to add further edges
 					if(numPoints > 1) 
 					{
@@ -171,7 +158,6 @@ public class RRG
 						transition		= productAutomaton.changePreSystemVarsToPostSystemVars(Environment.getLabelling().getLabel(xNew2D));
 						transition 		= transition.and(Environment.getLabelling().getLabel(xNearest2D));
 						if(type == 1 && transition.and(productAutomaton.getBDD()).and(fromStates).and(productAutomaton.changePreVarsToPostVars(toStates)).and(ProductAutomaton.getPropertyBDD()).and(ProductAutomaton.getLabelEquivalence()).isZero()) {
-							System.out.println("ajhdkjajaskfaj");
 							return false;
 						}
 						transitions.orWith(transition);
@@ -216,21 +202,24 @@ public class RRG
 					
 					// add point to the Rtree
 					Rectangle rect 			= new Rectangle(xNew.x, xNew.y, xNew.x, xNew.y);
-					PrintStream out = System.out;
-					System.setOut(new PrintStream(OutputStream.nullOutputStream()));
+//					System.setOut(new PrintStream(OutputStream.nullOutputStream()));
 					tree.add(rect, numPoints);
-					System.setOut(out);
+//					System.setOut(System.out);
 					treePoints.add(xNew);
 					numPoints++;
 				}
 		        return true;
 		    }
 		};
[…]

### 2020-09-21 MotionPlanning/src/modules/emptinessCheck/EmptinessCheck.java

@@ -6,10 +6,10 @@ package modules.emptinessCheck;
 import java.util.ArrayList;
 
 import abstraction.ProductAutomaton;
-import abstraction.exceptions.StateException;
 import net.sf.javabdd.BDD;
 import net.sf.javabdd.BDD.BDDIterator;
 import net.sf.javabdd.BDDFactory;
+import settings.PlanningException;
 
 /**
  * @author kush
@@ -42,7 +42,7 @@ public class EmptinessCheck {
 	}
 	
 //	public boolean findAcceptingPath() throws Exception{
-	public ArrayList<BDD> findAcceptingPath() throws Exception
+	public ArrayList<BDD> findAcceptingPath() throws PlanningException
 	{
 		push(null, productAutomaton.getInitStates());
 		while(! todo.isEmpty()) 
@@ -57,7 +57,7 @@ public class EmptinessCheck {
 				BDD toState		= productAutomaton.getSecondState(transition);
 				if(toState.isZero()) 
 				{
-					throw new StateException("State does not exist");
+					throw new PlanningException("State does not exist");
 				}
 				ArrayList<Integer> accSet	= productAutomaton.findAcceptingSets(transition);
 				if(getRank(toState) == -1) 
@@ -76,7 +76,7 @@ public class EmptinessCheck {
 		return null;		
 	}
 	
-	private ArrayList<BDD> findPath() throws Exception 
+	private ArrayList<BDD> findPath() throws PlanningException 
 	{
 		BDD statesSCC				= initializeStatesSCC();
 		ArrayList<BDD> path			= new ArrayList<BDD>();
@@ -93,7 +93,7 @@ public class EmptinessCheck {
 		return path;
 	}
 
-	private BDD initializeStatesSCC() throws Exception 
+	private BDD initializeStatesSCC() throws PlanningException 
 	{
 		statesSCC			= factory.zero();
 		int topRank			= SCC.getTop().getRoot();
@@ -107,7 +107,7 @@ public class EmptinessCheck {
 		return statesSCC;
 	}
 
-	private BDD bfs(BDD fromState, int[] visitedAcceptingSets, ArrayList<BDD> path) throws StateException, Exception 
+	private BDD bfs(BDD fromState, int[] visitedAcceptingSets, ArrayList<BDD> path) throws PlanningException
 	{
 		ArrayList<BDD> setStates	= new ArrayList<BDD>();
 		setStates.add(fromState);
@@ -154,7 +154,7 @@ public class EmptinessCheck {
[…]

### 2020-09-21 MotionPlanning/src/modules/emptinessCheck/SCCNode.java

@@ -14,7 +14,7 @@ public class SCCNode {
 	BDD rem;
 	SCCNode previous=null;
 	
-	public SCCNode(int root, ArrayList<Integer> last, ArrayList<Integer> acc, BDD rem) throws Exception{
+	public SCCNode(int root, ArrayList<Integer> last, ArrayList<Integer> acc, BDD rem) {
 		checkRem(rem);
 		this.root=root;
 		this.last=last;

### 2020-09-21 MotionPlanning/src/modules/emptinessCheck/stackSCC.java

@@ -3,11 +3,12 @@ package modules.emptinessCheck;
 import java.util.ArrayList;
 
 import net.sf.javabdd.BDD;
+import settings.PlanningException;
 
 public class stackSCC {
 	SCCNode top;
 		
-	public void push(int root, ArrayList<Integer> last, ArrayList<Integer> acc, BDD rem) throws Exception{
+	public void push(int root, ArrayList<Integer> last, ArrayList<Integer> acc, BDD rem) throws PlanningException{
 		SCCNode newNode=new SCCNode(root,last,acc,rem);
 		newNode.setPrevious(top);
 		top=newNode;

### 2020-09-21 MotionPlanning/src/modules/emptinessCheck/stackTODO.java

@@ -1,11 +1,12 @@
 package modules.emptinessCheck;
 
 import net.sf.javabdd.BDD;
+import settings.PlanningException;
 
 public class stackTODO {
 	todoNode top;
 	
-	public void push(BDD state, BDD transitions) throws Exception{
+	public void push(BDD state, BDD transitions) throws PlanningException {
 		todoNode newNode=new todoNode(state, transitions); 
 		newNode.setPrevious(top);
 		top=newNode;

### 2020-09-21 MotionPlanning/src/modules/emptinessCheck/todoNode.java

@@ -1,22 +1,22 @@
 package modules.emptinessCheck;
 
 import abstraction.ProductAutomaton;
-import abstraction.exceptions.StateException;
 import net.sf.javabdd.BDD;
+import settings.PlanningException;
 
 public class todoNode {
 	BDD state, transitions;
 	todoNode previous=null;
 	
-	todoNode(BDD state, BDD transitions) throws Exception{
+	todoNode(BDD state, BDD transitions) throws PlanningException{
 		checkValidity(transitions);
 		this.state=state;
 		this.transitions=transitions;
 	}
 
-	private void checkValidity(BDD transitions) throws Exception {
+	private void checkValidity(BDD transitions) throws PlanningException {
 		if(transitions.satCount(ProductAutomaton.allPreVars())>1) {
-			throw new StateException("More than one state");
+			throw new PlanningException("More than one state");
 		}
 	}
 	

### 2020-09-21 MotionPlanning/src/modules/learnAskExperiments/DefaultExperiment.java

@@ -184,27 +184,27 @@ public class DefaultExperiment implements Experiments
 	 */
 	public BDD addFilters() throws Exception 
 	{
-		BDD h	= ProductAutomaton.ithVarSystemPre(0);
-		BDD r1	= ProductAutomaton.ithVarSystemPre(1);
-		BDD r2	= ProductAutomaton.ithVarSystemPre(2);
-		BDD r3	= ProductAutomaton.ithVarSystemPre(3);
-		BDD r4	= ProductAutomaton.ithVarSystemPre(4);
-		BDD r5	= ProductAutomaton.ithVarSystemPre(5);
-		BDD r6	= ProductAutomaton.ithVarSystemPre(6);
-//		BDD c	= ProductAutomaton.ithVarSystemPre(5);
-//		BDD t	= ProductAutomaton.ithVarSystemPre(3);
-		BDD b	= ProductAutomaton.ithVarSystemPre(7);
-		BDD filter1	= h.imp((r1.or(r2).or(r3).or(r4).or(r5).or(r6).or(b)).not());
-		BDD filter2	= r1.imp((h.or(r2).or(r3).or(r4).or(r5).or(r6)).not());
-		BDD filter3	= r2.imp((h.or(r1).or(r3).or(r4).or(r5).or(r6)).not());
-		BDD filter4	= r3.imp((h.or(r1).or(r2).or(r4).or(r5).or(r6)).not());
-		BDD filter5 = r4.imp((h.or(r1).or(r2).or(r3).or(r5).or(r6)).not());
-		BDD filter6	= r5.imp((h.or(r1).or(r2).or(r3).or(r4).or(r6)).not());
-		BDD filter7	= r6.imp((h.or(r1).or(r2).or(r3).or(r4).or(r5)).not());
-		BDD filter	= filter1.and(filter2).and(filter3).and(filter4).and(filter5).and(filter6).and(filter7);
-		BDD filterPrime	= productAutomaton.changePreVarsToPostVars(filter);
-		return filter.and(filterPrime);
-//		return factory.one();
+//		BDD h	= ProductAutomaton.ithVarSystemPre(0);
+//		BDD r1	= ProductAutomaton.ithVarSystemPre(1);
+//		BDD r2	= ProductAutomaton.ithVarSystemPre(2);
+//		BDD r3	= ProductAutomaton.ithVarSystemPre(3);
+//		BDD r4	= ProductAutomaton.ithVarSystemPre(4);
[…]

### 2020-09-21 MotionPlanning/src/planningIO/printing/PrintAcceptingPath.java

@@ -4,10 +4,11 @@ import java.util.ArrayList;
 
 import abstraction.ProductAutomaton;
 import net.sf.javabdd.BDD;
+import settings.PlanningException;
 
 public class PrintAcceptingPath 
 {
-	public PrintAcceptingPath(ArrayList<BDD> path) throws Exception 
+	public PrintAcceptingPath(ArrayList<BDD> path) throws PlanningException 
 	{
 		BDD next;
 		for(int i=0; i<path.size(); i++) 
@@ -19,7 +20,7 @@ public class PrintAcceptingPath
 		System.out.print("\n\n");
 	}
 
-	private void printAPList(BDD state) throws Exception 
+	private void printAPList(BDD state) throws PlanningException 
 	{
 		ArrayList<String> apList	= findAPList(state);
 		System.out.print("[");
@@ -37,7 +38,7 @@ public class PrintAcceptingPath
 		System.out.print("]  ");
 	}
 
-	private ArrayList<String> findAPList(BDD state) throws Exception 
+	private ArrayList<String> findAPList(BDD state) throws PlanningException 
 	{
 		ArrayList<String> list		= new ArrayList<String>();
 		for(int i=0; i<ProductAutomaton.numAPSystem; i++) 

### 2020-09-21 MotionPlanning/src/planningIO/printing/PrintProductAutomaton.java

@@ -4,9 +4,9 @@ import java.io.BufferedWriter;
 import java.io.FileWriter;
 
 import abstraction.ProductAutomaton;
-import abstraction.exceptions.StateException;
 import net.sf.javabdd.BDD;
 import net.sf.javabdd.BDD.BDDIterator;
+import settings.PlanningException;
 
 public class PrintProductAutomaton
 {
@@ -109,7 +109,7 @@ public class PrintProductAutomaton
 		String stateLabel	= Integer.toString(state.scanVar(ProductAutomaton.propertyDomainPre()).intValue());
 		if(state.pathCount() != 1)
 		{
-			throw new StateException("More than one state");
+			throw new PlanningException("More than one state");
 		}
 		for(int i=0; i<ProductAutomaton.numAPSystem; i++) 
 		{

### 2020-09-21 MotionPlanning/src/planningIO/printing/ShowGraph.java

@@ -43,6 +43,33 @@ public class ShowGraph extends JFrame {
 		plot.setLimits(env.getBoundsX()[0], env.getBoundsX()[1], env.getBoundsY()[0], env.getBoundsY()[1]);
 		
 		
+		DefaultEdge nextEdge;
+		Vertex source, target;
+		
+		// Whole Graph
+		plot.setColor("black");
+		Iterator<DefaultEdge> ite = graph.edgeSet().iterator();
+		while(ite.hasNext())
+		{
+			nextEdge = ite.next();
+			source = graph.getEdgeSource(nextEdge);
+			target = graph.getEdgeTarget(nextEdge);
+			plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
+		}
+
+		// Found Path
+		plot.setColor("green");
+		ite = path.iterator();
+		while(ite.hasNext())
+		{
+			nextEdge = ite.next();
+			source = graph.getEdgeSource(nextEdge);
+			target = graph.getEdgeTarget(nextEdge);
+			plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
+		}
+		
+		
+		//Obstacles
 		plot.setColor("blue");
 		Iterator<Path2D> i = env.getObstacles().iterator();
 		while(i.hasNext())
@@ -67,28 +94,31 @@ public class ShowGraph extends JFrame {
 			}
 		}
 		
-		DefaultEdge nextEdge;
-		Vertex source, target;
+		// Label
+		plot.setColor("red");
+		i = Environment.getLabelling().getAreas().iterator();
+		while(i.hasNext())
+		{
+			Path2D rect = i.next();
+			PathIterator it = rect.getPathIterator(null);
+			float[] coords = new float[] {0f, 0f}, from = new float[] {0f, 0f}, to;
+			while(! it.isDone()) {
+				switch(it.currentSegment(coords)) {
+					case(PathIterator.SEG_MOVETO):
+						from = coords.clone();
+						break;
+					case(PathIterator.SEG_LINETO):
+						to = coords.clone();
+						plot.add("line", new double[] {from[0], to[0]}, new double[] {from[1], to[1]});
+						from = to.clone();
+						break;
+					default:
+						break;
+				}
+				it.next();
+			}
+		}
 		
-		plot.setColor("black");
[…]

### 2020-09-21 MotionPlanning/src/settings/PlanningSettings.java

@@ -53,7 +53,7 @@ public class PlanningSettings
 																					"After how many samples, return null if don't find anything from the advised transitions"},									
 			{ INTEGER_TYPE,		TRANSITION_THRESHOLD,					"Threshold for transitions",			"1.0",			new Integer(20),															"",
 																					"Threshold after how many copies of a transition would decrease its level by 1."},
-			{ FLOAT_TYPE,		ETA,									"Maximum radius for RRG",				"1.0",			new Float(1.0),																"",
+			{ FLOAT_TYPE,		ETA,									"Maximum radius for RRG",				"1.0",			new Float(1.5),																"",
 																					"Maximum radius to find the neighbours"}
 		};
 																			

### 2021-02-23 MotionPlanning/src/Planning.java
erro in libbuddy

@@ -3,7 +3,8 @@ import net.sf.javabdd.BDD.BDDIterator;
 import settings.Initialize;
 import settings.PlanningException;
 import settings.PlanningSettings;
-import modules.Discretization;
+import modules.KnownGrid;
+import modules.KnownRRG;
 import modules.RRG;
 import modules.UnknownRRG;
 import modules.learnAskExperiments.DefaultExperiment;
@@ -19,6 +20,7 @@ import org.jgrapht.graph.DefaultEdge;
 import abstraction.ProductAutomaton;
 import environment.Environment;
 import environment.Label;
+
 
 /**
  * @author kush
@@ -66,6 +68,7 @@ public class Planning
         currentStates			= initStateSystem.id(); //stores the set of states in the abstraction we have seen till now
         
         rrg.setStartingPoint(env.getInit());
+
         productAutomaton.setInitState(productAutomaton.getInitStates().and(initStateSystem)); // and for initial state in the product automaton
         
         initTime 				= System.nanoTime() - beginTime;
@@ -78,9 +81,14 @@ public class Planning
 	void printOutput() throws Exception 
 	{
 //    	productAutomaton.createDot(iterationNumber);
-		Pair<Float, Float> pathLength = rrg.plotGraph(finalPath);
+		Pair<Float, Float> length = new Pair<Float, Float>(0f, 0f);
+		if(! (boolean) PlanningSettings.get("firstExplThenPlan")) {
+			UnknownRRG urrg = (UnknownRRG) rrg;
+			length = urrg.plotGraph(finalPath);			
+		}
 		if((boolean) PlanningSettings.get("firstExplThenPlan")) {
-			pathLength.setFirst(moveLength);
+			length.setFirst(moveLength);
+			length.setSecond(pathLength);
 		}
 		double totalTime = System.nanoTime() - beginTime;
         
@@ -90,8 +98,8 @@ public class Planning
 		
 		System.out.println("\n\nTotal sampled points = " + rrg.totalSampledPoints);	
 		System.out.println("\nTotal useful sampled points = " + rrg.totalPoints);
-		System.out.println("Movement Length = " + pathLength.getFirst());
-		System.out.println("Remaining path Length = " + pathLength.getSecond());
+		System.out.println("Movement Length = " + length.getFirst());
+		System.out.println("Remaining path Length = " + length.getSecond());
 		System.out.println("Sampled from advice = " + adviceSamples);
 		System.out.println("\nAdvice samples: " + Arrays.toString(rrg.adviceSampled));
[…]

### 2021-02-23 MotionPlanning/src/environment/Environment.java

@@ -11,7 +11,7 @@ import java.awt.geom.Rectangle2D;
 import java.util.ArrayList;
 import java.util.Iterator;
 
-import modules.Discretization;
+import modules.Grid;
 
 public class Environment {
 	Shape bounds;
@@ -25,11 +25,11 @@ public class Environment {
 	Area allObs;
 	int numOfObs;
 	int numOfWalls;
-	static Point2D.Float init;
+	static Point2D init;
 	static Label labelling;
-	Discretization discretization;
+	Grid discretization;
 	
-	public Environment(float[] boundsX, float[] boundsY, ArrayList<Path2D> obstacles, ArrayList<Path2D> walls, Point2D.Float init, Label labelling) 
+	public Environment(float[] boundsX, float[] boundsY, ArrayList<Path2D> obstacles, ArrayList<Path2D> walls, Point2D init, Label labelling) 
 	{
 		this.bounds				= new Rectangle2D.Float(boundsX[0],boundsY[0],boundsX[1]-boundsX[0],boundsY[1]-boundsY[0]);
 		this.setBoundsX(boundsX);
@@ -43,7 +43,7 @@ public class Environment {
 		Environment.labelling	= labelling;
 	}
 	
-	public void setdiscretization(Discretization discretization) 
+	public void setdiscretization(Grid discretization) 
 	{
 		this.discretization = discretization;
 	}
@@ -103,7 +103,8 @@ public class Environment {
 	
 	public Boolean collisionFreeFromOpaqueObstacles(Point2D start, Point2D end) 
 	{
-		Line2D line		= new Line2D.Float(start,end), boundary;
+		Line2D line		= new Line2D.Float(start, end);
+		Line2D boundary;
 		float[] coords1	= new float[6], coords2=new float[6];
 		int segmentType;
 		

### 2021-02-23 MotionPlanning/src/environment/GenerateRandomEnv.java

@@ -1,0 +1,537 @@
+package environment;
+
+import java.awt.geom.AffineTransform;
+import java.awt.geom.Path2D;
+import java.awt.geom.PathIterator;
+import java.awt.geom.Point2D;
+import java.io.IOException;
+import java.util.ArrayList;
+
+import settings.PlanningSettings;
+
+public class GenerateRandomEnv {
+	public Environment env;
+
+	public GenerateRandomEnv() throws IOException {
+		System.out.println("Generating random environment...");
+		
+		
+		Point2D p1, p3, init;
+		ArrayList<Path2D> seeThroughObstacles = new ArrayList<Path2D>();
+		ArrayList<Path2D> obstacles	= new ArrayList<Path2D>();
+		ArrayList<Path2D> labels = new ArrayList<Path2D>();
+		ArrayList<String> apList = new ArrayList<String>();
+		Path2D tables = new Path2D.Float();
+		Path2D bins = new Path2D.Float();
+
+		
+		p1 = new Point2D.Float(0f,0f);
+		p3 = new Point2D.Float(6f,3f);
+		
+		init = new Point2D.Float(0.1f,1.5f);
+		
+		Path2D rect;
+		
+		rect = new Path2D.Float();
+		rect.moveTo(0f, 0.95f);
+		rect.lineTo(0.8f, 0.95f);
+		rect.lineTo(0.8f, 1.05f);
+		rect.lineTo(0f, 1.05f);
+		rect.lineTo(0f, 0.95f);
+		rect.closePath();
+		obstacles.add(rect);
+		
+		rect = new Path2D.Float();
+		rect.moveTo(1.2f, 0.95f);
+		rect.lineTo(2f, 0.95f);
+		rect.lineTo(2f, 1.05f);
+		rect.lineTo(1.2f, 1.05f);
+		rect.lineTo(1.2f, 0.95f);
+		rect.closePath();
+		obstacles.add(rect);
+		
+		rect = new Path2D.Float();
+		rect.moveTo(0f, 1.95f);
+		rect.lineTo(0.8f, 1.95f);
+		rect.lineTo(0.8f, 2.05f);
+		rect.lineTo(0f, 2.05f);
+		rect.lineTo(0f, 1.95f);
+		rect.closePath();
+		obstacles.add(rect);
[…]

### 2021-02-23 MotionPlanning/src/modules/Discretization.java

@@ -1,682 +1,0 @@
-package modules;
-
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import java.util.Iterator;
-import java.util.Set;
-
-import org.jgrapht.Graph;
-import org.jgrapht.GraphPath;
-import org.jgrapht.alg.shortestpath.DijkstraShortestPath;
-import org.jgrapht.alg.util.Pair;
-import org.jgrapht.graph.DefaultEdge;
-import org.jgrapht.graph.SimpleGraph;
-
-import abstraction.ProductAutomaton;
-import environment.EdgeSupplier;
-import environment.Environment;
-import environment.Vertex;
-import environment.VertexSupplier;
-import net.sf.javabdd.BDD;
-import settings.PlanningSettings;
-
-public class Discretization 
-{ 
-	float x1, x2, y1, y2, size;
-	int numX, numY;
-	int[][] discretization; // 0: don't know    1: free     2: visited     3:Obstacle
-	ArrayList<ArrayList<BDD>> labels;
-	boolean flag[][]; // used to compute frontiers
-	ArrayList<ArrayList<int[]>> frontiers;
-	ArrayList<Pair<Point2D, Float>> adviceFrontiers;
-	Graph<Vertex, DefaultEdge> graph;
-	
-	
-	public Discretization(Environment env, float size) 
-	{
-		this.x1 = env.getBoundsX()[0];
-		this.y1 = env.getBoundsY()[0];
-		this.x2 = env.getBoundsX()[1];
-		this.y2 = env.getBoundsY()[1];
-		this.numX = (int) ((x2 - x1)/size);
-		this.numY = (int) ((y2 - y1)/size);
-		this.size = size;
-		this.flag = new boolean[numX][numY];
-		discretization = new int[numX][numY];
-		this.graph = new SimpleGraph<Vertex, DefaultEdge>(new VertexSupplier(), new EdgeSupplier(), true);
-		labels = new ArrayList<ArrayList<BDD>>();
-		
-		for(int i=0;i<numX;i++) {
-			labels.add(new ArrayList<BDD>());
-			for(int j=0;j<numY;j++) {
-				labels.get(i).add(ProductAutomaton.factory.zero());
-				discretization[i][j] = 0;
-			}
-		}
-		frontiers = new ArrayList<ArrayList<int[]>>();
-		adviceFrontiers = new ArrayList<Pair<Point2D, Float>>();
-	}
-
-	public void initializeGraph() {
[…]

### 2021-02-23 MotionPlanning/src/modules/KnownGrid.java

@@ -1,0 +1,230 @@
+package modules;
+
+import java.awt.geom.Point2D;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.util.Set;
+
+import org.jgrapht.Graph;
+import org.jgrapht.GraphPath;
+import org.jgrapht.alg.shortestpath.DijkstraShortestPath;
+import org.jgrapht.alg.util.Pair;
+import org.jgrapht.graph.DefaultEdge;
+import org.jgrapht.graph.SimpleGraph;
+
+import abstraction.ProductAutomaton;
+import environment.EdgeSupplier;
+import environment.Environment;
+import environment.Vertex;
+import environment.VertexSupplier;
+import net.sf.javabdd.BDD;
+import settings.PlanningSettings;
+
+public class KnownGrid extends Grid {
+
+	Graph<Vertex, DefaultEdge> graph;
+	private ArrayList<DefaultEdge> movement;
+
+	public KnownGrid(Environment env, float size) 
+	{
+		super(env, size);
+		this.graph = new SimpleGraph<Vertex, DefaultEdge>(new VertexSupplier(), new EdgeSupplier(), true);
+		this.movement = new ArrayList<DefaultEdge>();
+	}
+	
+	public void initializeGraph() {
+		for(int i=0; i<numX; i++) {
+			for(int j=0; j<numY; j++) {
+				Vertex v = new Vertex(findCentre(i,j));
+				graph.addVertex(v);
+			}
+		}
+	}
+	
+	private void updateGraph(Point2D centre) {
+		Point2D left   = new Point2D.Float((float) centre.getX()-size, (float) centre.getY());
+		Point2D down   = new Point2D.Float((float) centre.getX(), (float) centre.getY()+size);
+		Vertex leftV   = findTheVertex(left);
+		Vertex downV   = findTheVertex(down);
+		Vertex centreV = findTheVertex(centre);
+		
+		int i = clampX(findCell(left)[0]);
+		int j = clampY(findCell(left)[1]);
+		if(grid[i][j] == 1){
+			DefaultEdge edge = graph.addEdge(leftV, centreV);
+			if(edge != null) {
+				graph.setEdgeWeight(edge, size);
+			}
+		}
+		i = clampX(findCell(down)[0]);
+		j = clampY(findCell(down)[1]);
+		if(grid[i][j] == 1){
+			DefaultEdge edge = graph.addEdge(downV, centreV);
[…]

### 2021-02-23 MotionPlanning/src/modules/KnownRRG.java

@@ -4,40 +4,27 @@ import java.awt.geom.Point2D;
 import java.util.ArrayList;
 import java.util.List;
 
+import org.jgrapht.Graph;
 import org.jgrapht.alg.util.Pair;
 import org.jgrapht.graph.DefaultEdge;
-import org.jgrapht.graph.SimpleGraph;
-
 import com.infomatiq.jsi.Point;
 import com.infomatiq.jsi.Rectangle;
 
 import abstraction.ProductAutomaton;
-import environment.EdgeSupplier;
 import environment.Environment;
 import environment.Vertex;
-import environment.VertexSupplier;
 import gnu.trove.TIntProcedure;
 import net.sf.javabdd.BDD;
-import planningIO.StoreGraph;
-import planningIO.printing.ShowGraph;
+import planningIO.StoreGraphKnown;
+import planningIO.printing.ShowGraphKnown;
 import settings.PlanningSettings;
 
 public class KnownRRG extends RRG {
 	
-	// used for exporting graph
-	boolean flagBin = false;
-	boolean flagRoom = false;
-	boolean flagFirstMove = false;
+	
 	
 	public KnownRRG(Environment env) {
-		this.env 				= env;
-		this.maximumRRGStepSize = (float) PlanningSettings.get("eta");
-		float[] sub 			= new float[] {env.getBoundsX()[1]-env.getBoundsX()[0], env.getBoundsY()[1]-env.getBoundsY()[0]};
-		this.gamma 				= 2.0 * Math.pow(1.5,0.5) * Math.pow(sub[0]*sub[1]/Math.PI,0.5);
-		this.graph 				= new SimpleGraph<Vertex, DefaultEdge>(new VertexSupplier(), new EdgeSupplier(), true);
-		this.discretization 	= new Discretization(env, (float) PlanningSettings.get("discretizationSize"));
-		this.forwardSampledTransitions = ProductAutomaton.factory.zero();
-		this.tree.init(null);
+		super(env);
 	}
 	
 	
@@ -84,11 +71,11 @@ public class KnownRRG extends RRG {
 //					plotting the first time it sees a bin
 					try {
 						if(! flagBin && ! Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars+7)).isZero()) {
-							plotGraph(null);
+							plotGraph(null, null);
 							flagBin = true;
 						}
[…]

### 2021-02-23 MotionPlanning/src/modules/UnknownGrid.java

@@ -1,0 +1,217 @@
+package modules;
+
+import java.awt.geom.Point2D;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.util.Set;
+
+import org.jgrapht.Graph;
+import org.jgrapht.GraphPath;
+import org.jgrapht.alg.shortestpath.DijkstraShortestPath;
+import org.jgrapht.alg.util.Pair;
+import org.jgrapht.graph.DefaultEdge;
+
+import com.infomatiq.jsi.Point;
+import com.infomatiq.jsi.SpatialIndex;
+import environment.Environment;
+import environment.Vertex;
+import gnu.trove.TIntProcedure;
+
+public class UnknownGrid extends Grid {
+
+	ArrayList<Pair<Point2D, Integer>> adviceFrontiers;
+	public int currentRank = -1;
+	Graph<Vertex, DefaultEdge> graph;
+	SpatialIndex tree;
+	ArrayList<Point> treePoints = new ArrayList<Point>();; // list of all the points in the Rtree/graphRRG
+	Environment env;
+	
+	public UnknownGrid(Environment env, float size, Graph<Vertex, DefaultEdge> graph, SpatialIndex tree, ArrayList<Point> treePoints) 
+	{
+		super(env, size);
+		this.adviceFrontiers = new ArrayList<Pair<Point2D, Integer>>();
+		this.graph = graph;
+		this.tree = tree;
+		this.env = env;
+		this.treePoints = treePoints;
+	}
+
+	private Pair<Point2D, Pair<Integer, Float>> findBestAdviceFrontier(Point2D xRobot) {
+		Point2D bestPoint = findFrontierCenter(frontiers.get(0));
+		float currentIG;
+		float bestIG = 0;
+		int bestRank = -1;
+		int bestIndex = 0;
+		for(int i=0; i<adviceFrontiers.size(); i++) {
+			int currentRank = adviceFrontiers.get(i).getSecond();
+			currentIG = findIGInAdviceFrontier(adviceFrontiers.get(i), bestRank, xRobot);
+			if(currentIG > bestIG) {
+				bestPoint = adviceFrontiers.get(i).getFirst();
+				bestIG = currentIG;
+				bestRank = currentRank;
+				bestIndex = i;
+			}
+		}
+		return new Pair<Point2D, Pair<Integer, Float>>(bestPoint, new Pair<Integer, Float>(bestIndex, bestIG));
+	}
+	
+	
+	/*
[…]

### 2021-02-23 MotionPlanning/src/modules/UnknownRRG.java

@@ -5,8 +5,8 @@ import java.awt.geom.Point2D;
 import java.io.IOException;
 import gnu.trove.TIntProcedure;
 import net.sf.javabdd.BDD;
-import planningIO.StoreGraph;
-import planningIO.printing.ShowGraph;
+import planningIO.StoreGraphUnknown;
+import planningIO.printing.ShowGraphUnknown;
 import settings.PlanningSettings;
 
 import java.util.ArrayList;
@@ -16,32 +16,28 @@ import org.jgrapht.GraphPath;
 import org.jgrapht.alg.shortestpath.DijkstraShortestPath;
 import org.jgrapht.alg.util.Pair;
 import org.jgrapht.graph.DefaultEdge;
-import org.jgrapht.graph.SimpleGraph;
-
 import com.infomatiq.jsi.Point;
 import com.infomatiq.jsi.Rectangle;
 import abstraction.ProductAutomaton;
-import environment.EdgeSupplier;
 import environment.Environment;
 import environment.Vertex;
-import environment.VertexSupplier;
 
 public class UnknownRRG extends RRG
 {
-	public Discretization discretization;
-	public ArrayList<Point2D> movement = new ArrayList<Point2D>();
+	public UnknownGrid grid;
 
 	float sensingRadius;
 	public int numOfFrontierUpdates = 0;
 	public float moveTime = 0;
-	Point2D currentRobotPosition;
 
-	// used for exporting graph
+	// used for exporting graphRRG
 	boolean flagBin = false;
 	boolean flagRoom = false;
 	boolean flagFirstMove = false;
 	
 	boolean explorationComplete = false;
+	
+	ArrayList<BDD> movementBDD;
 	
 	
 	/**
@@ -50,17 +46,11 @@ public class UnknownRRG extends RRG
 	 */
 	public UnknownRRG(Environment env) 
 	{
-		this.env 				= env;
-		this.maximumRRGStepSize = (float) PlanningSettings.get("eta");
-		float[] sub 			= new float[] {env.getBoundsX()[1]-env.getBoundsX()[0], env.getBoundsY()[1]-env.getBoundsY()[0]};
+		super(env);
 		this.sensingRadius		= (float) PlanningSettings.get("sensingRadius");
-		this.gamma 				= 2.0 * Math.pow(1.5,0.5) * Math.pow(sub[0]*sub[1]/Math.PI,0.5);
-		this.graph 				= new SimpleGraph<Vertex, DefaultEdge>(new VertexSupplier(), new EdgeSupplier(), true);
[…]

### 2021-02-23 MotionPlanning/src/planningIO/BuchiAutomataParser.java

@@ -49,12 +49,12 @@ public class BuchiAutomataParser
 			{
 				newEdge	= ProductAutomaton.factory.one();
 				newEdge.andWith(bddDomain[0].ithVar(stateId));
-				newEdge.andWith(bddDomain[1].ithVar(i));
+				newEdge = newEdge.and(bddDomain[1].ithVar(i));
 				
 				try 
 				{
 					newEdge.andWith(bddDomain[2].ithVar(accSignature.get(0)+1));
-//					System.out.println("Added edge in buchi from state "+stateId+" ----"+labelExpr.toString()+"----> "+i+" {"+accSignature.get(0)+"}");
+					System.out.println("Added edge in buchi from state "+stateId+" ----"+labelExpr.toString()+"----> "+i+" {"+accSignature.get(0)+"}");
 				} catch(NullPointerException E) 
 				{
 					newEdge.andWith(bddDomain[2].ithVar(0));

### 2021-02-23 MotionPlanning/src/planningIO/EnvironmentReader.java

@@ -21,8 +21,8 @@ public class EnvironmentReader
 	{
 		
 		Point2D.Float p1, p2, p3, p4, init;
-		ArrayList<Path2D> obstacles = new ArrayList<Path2D>();
-		ArrayList<Path2D> walls		= new ArrayList<Path2D>();
+		ArrayList<Path2D> seeThroughObstacles = new ArrayList<Path2D>();
+		ArrayList<Path2D> obstacles	= new ArrayList<Path2D>();
 		
 		
 		BufferedReader reader 		= new BufferedReader(new FileReader(envFile));
@@ -63,7 +63,7 @@ public class EnvironmentReader
 		
 		
 		
-		// For walls
+		// For obstacles
 		p 		= Pattern.compile(rectangle);
 		Path2D rect;
 		int i 	= 0;
@@ -73,9 +73,12 @@ public class EnvironmentReader
         }
 		
 		line = reader.readLine();
+		if(line.equalsIgnoreCase("obstacles")) {
+			line = reader.readLine();
+		}
 		while(line != null)
 		{
-			if(line.equalsIgnoreCase("obstacles")) {
+			if(line.equalsIgnoreCase("see through obstacles")) {
 				break;
 			}
 			System.out.println(line);
@@ -94,7 +97,7 @@ public class EnvironmentReader
 			rect.lineTo(Float.parseFloat(m.group(7)), Float.parseFloat(m.group(8)));
 			rect.lineTo(Float.parseFloat(m.group(1)), Float.parseFloat(m.group(2)));
 			rect.closePath();
-			walls.add(rect);
+			obstacles.add(rect);
 			
 			
 			if((boolean) PlanningSettings.get("verbosity"))
@@ -108,7 +111,7 @@ public class EnvironmentReader
 		
 		
 		
-		// For obstacles
+		// For see through obstacles
 		p 		= Pattern.compile(rectangle);
 		i 	= 0;
 		if((boolean) PlanningSettings.get("verbosity"))
@@ -133,7 +136,7 @@ public class EnvironmentReader
 			rect.lineTo(Float.parseFloat(m.group(7)), Float.parseFloat(m.group(8)));
 			rect.lineTo(Float.parseFloat(m.group(1)), Float.parseFloat(m.group(2)));
 			rect.closePath();
-			obstacles.add(rect);
+			seeThroughObstacles.add(rect);
 			
 			
 			if((boolean) PlanningSettings.get("verbosity"))
[…]

### 2021-02-23 MotionPlanning/src/planningIO/PropertyReader.java

@@ -9,6 +9,7 @@ import java.io.InputStreamReader;
 import java.util.ArrayList;
 
 import jhoafparser.parser.generated.ParseException;
+import settings.PlanningSettings;
 
 public class PropertyReader 
 {
@@ -17,13 +18,18 @@ public class PropertyReader
 	
 	public PropertyReader(ArrayList<String> apListSystem, String propertyFile) throws IOException 
 	{
-		// Read property from the file
-		BufferedReader propertyReader 	= new BufferedReader(new FileReader(propertyFile));
-        String propertyString			= propertyReader.readLine();
-        propertyReader.close();
-        
+		String propertyString;
+		if((boolean) PlanningSettings.get("randomEnv")) {
+			propertyString = "GF (r1 & b) & GF (r2 & b) & GF (r3 & b) & GF (r4 & b) & GF (r5 & b) & GF (r6 & b)";
+		}
+		else {
+			// Read property from the file
+			BufferedReader propertyReader 	= new BufferedReader(new FileReader(propertyFile));
+	        propertyString			= propertyReader.readLine();
+	        propertyReader.close();
+		}
         // Output 
-        System.out.println("\nProperty to satisfty: " + propertyString + "\n\n");
+        System.out.println("\nConstructing automaton for property: " + propertyString + " ...\n\n");
         
     	
         // Using SPOT
@@ -51,7 +57,7 @@ public class PropertyReader
         
         BufferedReader r 		= new BufferedReader(new InputStreamReader(p3.getInputStream()));
         String line;
-        BufferedWriter writer 	= new BufferedWriter(new FileWriter("/home/kush/Projects/robotmotionplanning/MotionPlanning/temp/propertyAutomata.hoa"));
+        BufferedWriter writer 	= new BufferedWriter(new FileWriter("temp/propertyAutomata.hoa"));
         
         while (true) 
         {
@@ -67,6 +73,7 @@ public class PropertyReader
 			propertyParser		= new BuchiAutomataParser( p1.getInputStream(), 
 					p2.getInputStream(), 
 					apListSystem);
+			System.out.println("done!");
 		} catch (ParseException e) 
         {
 			e.printStackTrace();

### 2021-02-23 MotionPlanning/src/planningIO/StoreGraph.java

@@ -1,172 +1,0 @@
-package planningIO;
-
-import java.awt.geom.Path2D;
-import java.awt.geom.PathIterator;
-import java.io.BufferedWriter;
-import java.io.FileWriter;
-import java.io.IOException;
-import java.util.Iterator;
-import java.util.List;
-
-import org.jgrapht.Graph;
-import org.jgrapht.graph.DefaultEdge;
-
-import environment.Environment;
-import environment.Vertex;
-
-public class StoreGraph 
-{
-	public float movementLength, remainingPathLength;
-
-	public StoreGraph(Environment env, Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> finalPath, List<DefaultEdge> movement, String fileName) {
-		try {
-			outputGraph(graph, fileName);
-			outputObstacles(env);
-			outputMovement(graph, movement, fileName);
-			outputFinalPath(graph, finalPath, fileName);
-		} catch (IOException e) {
-			e.printStackTrace();
-		}
-	}
-
-	public StoreGraph(Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> movement, String fileName) {
-		try {
-			outputGraph(graph, fileName);
-			outputMovement(graph, movement, fileName);
-		} catch (IOException e) {
-			e.printStackTrace();
-		}
-	}
-	
-	
-	public void outputGraph(Graph<Vertex, DefaultEdge> graph, String fileName) throws IOException {
-		BufferedWriter writer 		= new BufferedWriter(new FileWriter("temp/"+fileName+".csv"));
-		
-		writer.write("x1,y1,x2,y2\n");
-		
-		DefaultEdge nextEdge;
-		Vertex source, target;
-		
-		// Whole Graph
-		Iterator<DefaultEdge> ite = graph.edgeSet().iterator();
-		while(ite.hasNext())
-		{
-			nextEdge = ite.next();
-			source = graph.getEdgeSource(nextEdge);
-			target = graph.getEdgeTarget(nextEdge);
-			writer.write(source.getPoint().getX() + "," +source.getPoint().getY() + "," + target.getPoint().getX() + "," + target.getPoint().getY() + "\n");
-		}
-		
-		writer.close();
-	}
-	
-	public void outputFinalPath(Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> finalPath, String fileName) throws IOException {
[…]

### 2021-02-23 MotionPlanning/src/planningIO/StoreGraphUnknown.java

@@ -1,0 +1,172 @@
+package planningIO;
+
+import java.awt.geom.Path2D;
+import java.awt.geom.PathIterator;
+import java.io.BufferedWriter;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.util.Iterator;
+import java.util.List;
+
+import org.jgrapht.Graph;
+import org.jgrapht.graph.DefaultEdge;
+
+import environment.Environment;
+import environment.Vertex;
+
+public class StoreGraphUnknown
+{
+	public float movementLength, remainingPathLength;
+
+	public StoreGraphUnknown(Environment env, Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> finalPath, List<DefaultEdge> movement, String fileName) {
+		try {
+			outputGraph(graph, fileName);
+			outputObstacles(env);
+			outputMovement(graph, movement, fileName);
+			outputFinalPath(graph, finalPath, fileName);
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+	}
+
+	public StoreGraphUnknown(Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> movement, String fileName) {
+		try {
+			outputGraph(graph, fileName);
+			outputMovement(graph, movement, fileName);
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+	}
+	
+	
+	public void outputGraph(Graph<Vertex, DefaultEdge> graph, String fileName) throws IOException {
+		BufferedWriter writer = new BufferedWriter(new FileWriter("temp/"+fileName+".csv"));
+		
+		writer.write("x1,y1,x2,y2\n");
+		
+		DefaultEdge nextEdge;
+		Vertex source, target;
+		
+		// Whole Graph
+		Iterator<DefaultEdge> ite = graph.edgeSet().iterator();
+		while(ite.hasNext())
+		{
+			nextEdge = ite.next();
+			source = graph.getEdgeSource(nextEdge);
+			target = graph.getEdgeTarget(nextEdge);
+			writer.write(source.getPoint().getX() + "," +source.getPoint().getY() + "," + target.getPoint().getX() + "," + target.getPoint().getY() + "\n");
+		}
+		
+		writer.close();
+	}
+	
+	public void outputFinalPath(Graph<Vertex, DefaultEdge> graph, List<DefaultEdge> finalPath, String fileName) throws IOException {
[…]

### 2021-02-23 MotionPlanning/src/planningIO/printing/ShowGraph.java

@@ -1,168 +1,0 @@
-package planningIO.printing;
-
-import java.awt.Graphics;
-import java.awt.geom.Path2D;
-import java.awt.geom.PathIterator;
-import java.util.Iterator;
-import java.util.List;
-
-import javax.swing.JFrame;
-
-import org.jgrapht.Graph;
-import org.jgrapht.graph.DefaultEdge;
-
-import environment.Environment;
-import environment.Vertex;
-import ij.gui.Plot;
-
-public class ShowGraph extends JFrame {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 1L;
-
-	
-	Graph<Vertex, DefaultEdge> graph;
-	Environment env;
-	List<DefaultEdge> movement;
-	List<DefaultEdge> finalPath;
-	
-	
-	public ShowGraph(Graph<Vertex, DefaultEdge> graph, Environment env, List<DefaultEdge> movement, List<DefaultEdge> finalPath)
-	{
-		this.graph = graph;
-		this.env = env;
-		this.movement = movement;
-		this.finalPath = finalPath;
-	}
-	
-	
-	@Override
-	public void paint(Graphics l) 
-	{
-		Plot plot = new Plot("plot", "", "");
-		plot.setLimits(env.getBoundsX()[0], env.getBoundsX()[1], env.getBoundsY()[0], env.getBoundsY()[1]);
-		
-		
-		DefaultEdge nextEdge;
-		Vertex source, target;
-		
-		// Whole Graph
-		plot.setColor("black");
-		Iterator<DefaultEdge> ite = graph.edgeSet().iterator();
-		while(ite.hasNext())
-		{
-			nextEdge = ite.next();
-			source = graph.getEdgeSource(nextEdge);
-			target = graph.getEdgeTarget(nextEdge);
-			plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
-		}
-
-		// Robot Movement
-		plot.setColor("green");
-		if(movement!=null) {
-			ite = movement.iterator();
-			while(ite.hasNext())
-			{
-				nextEdge = ite.next();
-				source = graph.getEdgeSource(nextEdge);
-				target = graph.getEdgeTarget(nextEdge);
-				plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
-			}
-		}
-		
-		// Final Path
[…]

### 2021-02-23 MotionPlanning/src/planningIO/printing/ShowGraphKnown.java

@@ -1,0 +1,171 @@
+package planningIO.printing;
+
+import java.awt.Graphics;
+import java.awt.geom.Path2D;
+import java.awt.geom.PathIterator;
+import java.util.Iterator;
+import java.util.List;
+
+import javax.swing.JFrame;
+
+import org.jgrapht.Graph;
+import org.jgrapht.graph.DefaultEdge;
+
+import environment.Environment;
+import environment.Vertex;
+import ij.gui.Plot;
+
+public class ShowGraphKnown extends JFrame {
+	
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 1L;
+	
+	Graph<Vertex, DefaultEdge> graphRRG, graphMovement;
+	Environment env;
+	List<DefaultEdge> movement;
+	List<DefaultEdge> finalPath;
+	
+	
+	public ShowGraphKnown(Graph<Vertex, DefaultEdge> graphRRG, Graph<Vertex, DefaultEdge> graphMovement, Environment env, List<DefaultEdge> movement, List<DefaultEdge> finalPath)
+	{
+		this.graphRRG = graphRRG;
+		this.graphMovement = graphMovement;
+		this.env = env;
+		this.movement = movement;
+		this.finalPath = finalPath;
+	}
+	
+	
+	@Override
+	public void paint(Graphics l) 
+	{
+		Plot plot = new Plot("plot", "", "");
+		plot.setLimits(env.getBoundsX()[0], env.getBoundsX()[1], env.getBoundsY()[0], env.getBoundsY()[1]);
+		
+		
+		DefaultEdge nextEdge;
+		Vertex source, target;
+		
+		// Whole Graph
+		plot.setColor("black");
+		Iterator<DefaultEdge> ite = graphRRG.edgeSet().iterator();
+		while(ite.hasNext())
+		{
+			nextEdge = ite.next();
+			source = graphRRG.getEdgeSource(nextEdge);
+			target = graphRRG.getEdgeTarget(nextEdge);
+			plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
+		}
+
+		// Robot Movement
+		plot.setColor("green");
+		if(movement != null) {
+			ite = movement.iterator();
+			while(ite.hasNext())
+			{
+				nextEdge = ite.next();
+				source = graphMovement.getEdgeSource(nextEdge);
+				target = graphMovement.getEdgeTarget(nextEdge);
[…]

### 2021-02-23 MotionPlanning/src/planningIO/printing/ShowGraphUnknown.java

@@ -1,0 +1,168 @@
+package planningIO.printing;
+
+import java.awt.Graphics;
+import java.awt.geom.Path2D;
+import java.awt.geom.PathIterator;
+import java.util.Iterator;
+import java.util.List;
+
+import javax.swing.JFrame;
+
+import org.jgrapht.Graph;
+import org.jgrapht.graph.DefaultEdge;
+
+import environment.Environment;
+import environment.Vertex;
+import ij.gui.Plot;
+
+public class ShowGraphUnknown extends JFrame {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 1L;
+
+	
+	Graph<Vertex, DefaultEdge> graph;
+	Environment env;
+	List<DefaultEdge> movement;
+	List<DefaultEdge> finalPath;
+	
+	
+	public ShowGraphUnknown(Graph<Vertex, DefaultEdge> graph, Environment env, List<DefaultEdge> movement, List<DefaultEdge> finalPath)
+	{
+		this.graph = graph;
+		this.env = env;
+		this.movement = movement;
+		this.finalPath = finalPath;
+	}
+	
+	
+	@Override
+	public void paint(Graphics l) 
+	{
+		Plot plot = new Plot("plot", "", "");
+		plot.setLimits(env.getBoundsX()[0], env.getBoundsX()[1], env.getBoundsY()[0], env.getBoundsY()[1]);
+		
+		
+		DefaultEdge nextEdge;
+		Vertex source, target;
+		
+		// Whole Graph
+		plot.setColor("black");
+		Iterator<DefaultEdge> ite = graph.edgeSet().iterator();
+		while(ite.hasNext())
+		{
+			nextEdge = ite.next();
+			source = graph.getEdgeSource(nextEdge);
+			target = graph.getEdgeTarget(nextEdge);
+			plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
+		}
+
+		// Robot Movement
+		plot.setColor("green");
+		if(movement != null) {
+			ite = movement.iterator();
+			while(ite.hasNext())
+			{
+				nextEdge = ite.next();
+				source = graph.getEdgeSource(nextEdge);
+				target = graph.getEdgeTarget(nextEdge);
+				plot.add("line", new double[] {source.getPoint().getX(), target.getPoint().getX()}, new double[] {source.getPoint().getY(), target.getPoint().getY()});
+			}
+		}
+		
[…]

### 2021-02-23 MotionPlanning/src/settings/PlanningSettings.java

@@ -17,7 +17,7 @@ public class PlanningSettings
 	public static final String BDD_FACTORY_CACHE_SIZE		=	"bddFactoryCacheSize";
 	public static final String USE_SPOT						= 	"useSpot";
 	public static final String MAX_LEVEL_TRANSITION			=	"maxLevelTransition";
-	public static final String SAMPLING_THRESHOLD			=	"samplingThreshold";
+	public static final String RANDOM_ENV					=	"randomEnv";
 	public static final String TRANSITION_THRESHOLD			= 	"transitionThreshold";
 	public static final String ETA							=	"eta";
 	public static final String SENSING_RADIUS				=	"sensingRadius";
@@ -26,6 +26,8 @@ public class PlanningSettings
 	public static final String USE_ADVICE					=	"useAdvice";
 	public static final String FIRST_EXPL_THEN_PLAN			=	"firstExplThenPlan";
 	public static final String DEBUG						=	"debug";
+	public static final String INPUT_FILE					=	"inputFile"; 
+	public static final String ONLY_OPAQUE_OBSTACLES		=	"onlyOpaqueObstacles"; 
 	
 	
 	public static final Logger RTREELOGGER = Logger.getLogger(Logger.GLOBAL_LOGGER_NAME);
@@ -49,8 +51,8 @@ public class PlanningSettings
 			{ INTEGER_TYPE,		MAX_LEVEL_TRANSITION,					"Levels of transitions",				"1.0",			new Integer(4),																"",
 																					"Possible levels of transitions possible"},
 			
-			{ INTEGER_TYPE,		SAMPLING_THRESHOLD,						"Threshold for sampling",				"1.0",			new Integer(50),															"",
-																					"After how many samples, return null if don't find anything from the advised transitions"},	
+			{ BOOLEAN_TYPE,		RANDOM_ENV,								"Generate a random env",				"1.0",			new Boolean(false),															"",
+																					"Generate a random environment and run the algorithm on that"},	
 			
 			{ INTEGER_TYPE,		TRANSITION_THRESHOLD,					"Threshold for transitions",			"1.0",			new Integer(20),															"",
 																					"Threshold after how many copies of a transition would decrease its level by 1."},
[…]

### 2021-02-24 Presentation/Jana's group/scheme.tex
a

\begin{figure}
	\centering
    \begin{tikzpicture}[->,>=stealth', minimum width = 32mm,text width = 30mm]
        \node[state=initial, color = blue, anchor=center, fill = blue!10!white,thin] (A) {System};
        \node[state, color = blue, below of = A, fill = blue!10!white, node distance =1.5cm,thin] (B) {RRG graph};
        \node[state, color = blue, below of = B, fill = blue!10!white, node distance =1.5cm,ultra thick] (C) {Semantic abstraction};
        \node[state, color = red, below of = C, fill = red!10!white, xshift = -2cm, node distance =1.5cm,thin] (D) {Product};
        \node[state, color = green!50!black, left of = A, fill = green!10!white, node distance =4cm,thin] (E) {LTL property};
        \node[state, color = green!50!black, below of = E, fill = green!10!white, node distance =2.25cm,thin] (F) {Automaton};

        \path (A) edge[] (B)
        (B) edge[ultra thick] node[right,pos=0.5]{\texttt{Learn}} (C)
        (C) edge (D)
        (E) edge (F)
        (F) edge (D)
        (D.east) edge[bend right=70,looseness=1.5, ultra thick] node[right,pos=0.9]{\texttt{Bias}} (B.east)
        %guidance $\rightarrow$ better paths and learning w.r.t. property
        ;
    \end{tikzpicture}
    \caption{Scheme of our model-checking-inspired approach with novel elements drawn thickly. }
    \label{fig:guide}
\end{figure}

### 2021-02-24 MotionPlanning/lib/owl/javadoc/member-search-index.js

@@ -1,0 +1,1 @@
+memberSearchIndex = [{"p":"owl.grammar","c":"LTLLexer","l":"_ATN"},{"p":"owl.grammar","c":"LTLParser","l":"_ATN"},{"p":"owl.grammar","c":"LTLLexer","l":"_decisionToDFA"},{"p":"owl.grammar","c":"LTLParser","l":"_decisionToDFA"},{"p":"owl.grammar","c":"LTLLexer","l":"_serializedATN"},{"p":"owl.grammar","c":"LTLParser","l":"_serializedATN"},{"p":"owl.grammar","c":"LTLLexer","l":"_sharedContextCache"},{"p":"owl.grammar","c":"LTLParser","l":"_sharedContextCache"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"AbstractCachedStatesAutomaton()","url":"%3Cinit%3E()"},{"p":"owl.automaton","c":"AbstractImplicitAutomaton","l":"AbstractImplicitAutomaton(ValuationSetFactory, Collection<S>, A)","url":"%3Cinit%3E(owl.factories.ValuationSetFactory,java.util.Collection,A)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeMapVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeMapVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeMapVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeMapVisitor)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeTreeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeTreeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeTreeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeTreeVisitor)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(
[…]

### 2021-02-24 MotionPlanning/lib/owl/javadoc/package-search-index.js

@@ -1,0 +1,1 @@
+packageSearchIndex = [{"l":"All Packages","url":"allpackages-index.html"},{"l":"owl.automaton"},{"l":"owl.automaton.acceptance"},{"l":"owl.automaton.acceptance.optimizations"},{"l":"owl.automaton.algorithms"},{"l":"owl.automaton.edge"},{"l":"owl.automaton.output"},{"l":"owl.automaton.transformations"},{"l":"owl.automaton.util"},{"l":"owl.cinterface"},{"l":"owl.collections"},{"l":"owl.factories"},{"l":"owl.factories.jbdd"},{"l":"owl.game"},{"l":"owl.game.algorithms"},{"l":"owl.game.output"},{"l":"owl.grammar"},{"l":"owl.ltl"},{"l":"owl.ltl.algorithms"},{"l":"owl.ltl.ltlf"},{"l":"owl.ltl.parser"},{"l":"owl.ltl.rewriter"},{"l":"owl.ltl.robust"},{"l":"owl.ltl.util"},{"l":"owl.ltl.visitors"},{"l":"owl.run"},{"l":"owl.run.modules"},{"l":"owl.run.parser"},{"l":"owl.translations"},{"l":"owl.translations.canonical"},{"l":"owl.translations.delag"},{"l":"owl.translations.dpa2safety"},{"l":"owl.translations.dra2dpa"},{"l":"owl.translations.ltl2dpa"},{"l":"owl.translations.ltl2dra"},{"l":"owl.translations.ltl2ldba"},{"l":"owl.translations.ltl2nba"},{"l":"owl.translations.mastertheorem"},{"l":"owl.translations.modules"},{"l":"owl.translations.nba2dpa"},{"l":"owl.translations.nba2ldba"},{"l":"owl.translations.rabinizer"},{"l":"owl.util"},{"l":"owl.util.annotation"}]

### 2021-02-24 MotionPlanning/lib/owl/javadoc/type-search-index.js

@@ -1,0 +1,1 @@
+typeSearchIndex = [{"p":"owl.automaton","l":"AbstractCachedStatesAutomaton"},{"p":"owl.automaton","l":"AbstractImplicitAutomaton"},{"p":"owl.automaton.acceptance.optimizations","l":"AcceptanceOptimizations"},{"p":"owl.automaton.acceptance.optimizations","l":"AcceptanceOptimizations.AcceptanceOptimizationTransformer"},{"p":"owl.game.output","l":"Aig"},{"p":"owl.game.output","l":"AigConsumer"},{"p":"owl.game.output","l":"AigerPrinter"},{"p":"owl.game.output","l":"AigFactory"},{"p":"owl.game.output","l":"AigPrintable"},{"p":"owl.automaton.acceptance","l":"AllAcceptance"},{"l":"All Classes","url":"allclasses-index.html"},{"p":"owl.grammar","l":"LTLParser.AndExpressionContext"},{"p":"owl.translations.ltl2ldba","l":"AnnotatedLDBA"},{"p":"owl.automaton.util","l":"AnnotatedState"},{"p":"owl.automaton.util","l":"AnnotatedStateOptimisation"},{"p":"owl.translations.mastertheorem","l":"AsymmetricEvaluatedFixpoints"},{"p":"owl.translations.ltl2ldba","l":"AsymmetricLDBAConstruction"},{"p":"owl.translations.ltl2ldba","l":"AsymmetricProductState"},{"p":"owl.translations.ltl2dpa","l":"AsymmetricRankingState"},{"p":"owl.grammar","l":"LTLParser.AtomExpressionContext"},{"p":"owl.automaton","l":"Automaton"},{"p":"owl.automaton","l":"AutomatonFactory"},{"p":"owl.automaton","l":"AutomatonOperations"},{"p":"owl.automaton","l":"AutomatonReader"},{"p":"owl.run.modules","l":"OutputWriters.AutomatonStats"},{"p":"owl.automaton","l":"AutomatonUtil"},{"p":"owl.automaton","l":"Views.AutomatonView"},{"p":"owl.ltl","l":"Biconditional"},{"p":"owl.grammar","l":"LTLParser.BinaryExpressionContext"},{"p":"owl.grammar","l":"LTLParser.BinaryOpContext"},{"p":"owl.grammar","l":"LTLParser.BinaryOperationContext"},{"p":"owl.ltl","l":"Formula.BinaryTempor
[…]

### 2021-02-24 MotionPlanning/planner.sh

@@ -1,3 +1,15 @@
 #!/bin/bash
 
-java -jar planning.jar "$@"
+rm temp/output/*
+echo "iterations,sampled,rrgsize,movement,remaining,time," > temp/output/output.csv
+
+arg1="$1"
+shift
+
+for (( i=1; i <= $arg1; i++ ))
+do
+   echo "Current run: $i"
+   java -jar planning.jar "$@" > temp/output/output$i.txt
+done
+
+python3 temp/avg.py 

### 2021-02-24 MotionPlanning/src/Planning.java

@@ -1,5 +1,6 @@
 import net.sf.javabdd.BDD;
 import net.sf.javabdd.BDD.BDDIterator;
+import net.sf.javabdd.BDDFactory;
 import settings.Initialize;
 import settings.PlanningException;
 import settings.PlanningSettings;
@@ -44,6 +45,7 @@ public class Planning
 	double learnTime 		= 0;
 	double explTime 		= 0;
 	double adviceTime 		= 0;
+	double totalTime		= 0;
     boolean needNewAdvice 	= true;
     ArrayList<BDD> advice 	= null;
     float moveLength, pathLength;
@@ -52,12 +54,13 @@ public class Planning
     /**
      * initialize everything
      * @throws Exception
+	 * @param factory
      */
-	public Planning() throws Exception 
+	public Planning(BDDFactory factory) throws Exception
 	{
 		beginTime				= System.nanoTime();
 		
-		Initialize initialize	= new Initialize();// read the files and initialise everything
+		Initialize initialize	= new Initialize(factory);// read the files and initialise everything
         rrg 					= initialize.getRRG();
         productAutomaton		= initialize.getProductAutomaton();
         exper					= new DefaultExperiment(productAutomaton); // learn and advce procedures
@@ -80,44 +83,57 @@ public class Planning
 	 */
 	void printOutput() throws Exception 
 	{
-//    	productAutomaton.createDot(iterationNumber);
-		Pair<Float, Float> length = new Pair<Float, Float>(0f, 0f);
-		if(! (boolean) PlanningSettings.get("firstExplThenPlan")) {
-			UnknownRRG urrg = (UnknownRRG) rrg;
-			length = urrg.plotGraph(finalPath);			
+		if((boolean) PlanningSettings.get("debug")){
+			System.out.print("Exporting product automaton ... ");
+	    	productAutomaton.createDot(iterationNumber);
+			System.out.println("done");
 		}
+		Pair<Float, Float> length = new Pair<Float, Float>(0f, 0f);
 		if((boolean) PlanningSettings.get("firstExplThenPlan")) {
 			length.setFirst(moveLength);
 			length.setSecond(pathLength);
 		}
-		double totalTime = System.nanoTime() - beginTime;
+		else{
+			UnknownRRG urrg = (UnknownRRG) rrg;
+			System.out.print("Storing output ... ");
+			length = urrg.plotGraph(finalPath);
+			System.out.println("done");
+			moveLength = length.getFirst();
[…]

### 2021-02-24 MotionPlanning/src/environment/GenerateRandomEnv.java

@@ -13,7 +13,7 @@ public class GenerateRandomEnv {
 	public Environment env;
 
 	public GenerateRandomEnv() throws IOException {
-		System.out.println("Generating random environment...");
+		System.out.print("Generating random environment ... ");
 		
 		
 		Point2D p1, p3, init;

### 2021-02-24 MotionPlanning/src/mainCL.java

@@ -1,16 +1,64 @@
+import net.sf.javabdd.BDDFactory;
+import net.sf.javabdd.BuDDyFactory;
 import settings.PlanningSettings;
+
+import java.io.BufferedWriter;
+import java.io.FileWriter;
 
 public class mainCL {
 
 	public static void main(String[] args) throws Exception {
-		new PlanningSettings(args); // use default values for all variables
-		Planning plan = new Planning();
-		
-		if((boolean) PlanningSettings.get("firstExplThenPlan")) {
-			plan.firstExplThenPlan();
-		}
-		else {
-			plan.explAndPlanTogether();
+		new PlanningSettings(args); // set parameters
+		if((boolean) PlanningSettings.get("debug"))
+			PlanningSettings.outputParameters();
+
+		int numOfRuns = (Integer) PlanningSettings.get("numberOfRuns");
+		Object[][] output = new Object[numOfRuns][6];
+		Object[] data = new Object[6];
+		Object[] sum = new Object[]{
+				new Integer(0), // num of iterations
+				new Integer(0), // Total sampled points
+				new Integer(0), // RRG size
+				new Float(0), // movement length
+				new Float(0), // path length
+				new Double(0) // total time
+		};
+
+
+
+		for(int i=0; i < numOfRuns; i++) {
+			BDDFactory factory = BuDDyFactory.init(200000, (int) PlanningSettings.get("bddFactoryCacheSize"));
+			System.out.println("Starting initialization ... ");
+			Planning plan = new Planning(factory);
+			System.out.println("Initialization complete\n\n");
+			if ((boolean) PlanningSettings.get("firstExplThenPlan")) {
+				data = plan.firstExplThenPlan();
+			} else {
+				data = plan.explAndPlanTogether();
+			}
+			System.out.println("\n\n");
+			factory.done();
+
+			BufferedWriter writer = new BufferedWriter(new FileWriter("temp/output/output.csv", true));
+//			writer.write("iterations,sampled,rrgsize,movement,remaining,time,");
+			writer.newLine();
+			for(int j=0; j<6; j++){
+				output[i][j] = data[j];
+				if(j<3) {
+					writer.write(((Integer) data[j]) + ",");
+					sum[j] = new Integer((Integer) sum[j] + (Integer) data[j]);
+				}
+				else if(j<5) {
+					writer.write((Float) data[j] + ",");
[…]

### 2021-02-24 MotionPlanning/src/modules/KnownGrid.java

@@ -115,7 +115,7 @@ public class KnownGrid extends Grid {
 		Vertex temp;
 		while(it.hasNext()) {
 			temp 				= it.next();
-			if(Math.abs(temp.getPoint().getX() - p.getX()) < 0.0001  &&  Math.abs(temp.getPoint().getY() - p.getY()) < 0.0001) {
+			if(Math.abs(temp.getPoint().getX() - p.getX()) < 0.000001  &&  Math.abs(temp.getPoint().getY() - p.getY()) < 0.000001) {
 				return temp;
 			}
 		}
@@ -183,8 +183,9 @@ public class KnownGrid extends Grid {
 			return null;
 		}
 		Pair<Point2D, Pair<Integer, Float>> bestFrontier = findBestFrontier(frontiers, xRobot); // point, index, IG
-		
-		System.out.println("Used frontier: " + bestFrontier.getFirst() + " with IG = " + bestFrontier.getSecond().getSecond());
+
+		if((boolean) PlanningSettings.get("debug"))
+			System.out.println("Used frontier: " + bestFrontier.getFirst() + " with IG = " + bestFrontier.getSecond().getSecond());
 		return new Pair<Point2D, Integer>(bestFrontier.getFirst(), -1);
 	}
 	

### 2021-02-24 MotionPlanning/src/modules/KnownRRG.java

@@ -128,7 +128,8 @@ public class KnownRRG extends RRG {
 
 	public Pair<Float, Float> plotGraph(List<DefaultEdge> finalPath, Graph<Vertex, DefaultEdge> graphMovement) {
 		if(finalPath != null) {
-			new ShowGraphKnown(graph, graphMovement, env, movement, finalPath).setVisible(true);
+			if((boolean) PlanningSettings.get("generatePlot"))
+				new ShowGraphKnown(graph, graphMovement, env, movement, finalPath).setVisible(true);
 			StoreGraphKnown temp = new StoreGraphKnown(env, graph, graphMovement, finalPath, null, "end");
 			return new Pair<Float, Float>(temp.movementLength, temp.remainingPathLength);
 		} else if(! flagBin) {

### 2021-02-24 MotionPlanning/src/modules/RRG.java

@@ -198,8 +198,9 @@ public abstract class RRG {
 	 */
 	protected float getProb(int rank) 
 	{
+
 		if(rank == -1) {
-			return 0.5f;
+			return (float) PlanningSettings.get("biasProb");
 		}
 		else {
 			return 1f;
@@ -230,7 +231,7 @@ public abstract class RRG {
 	
 	/**
 	 * Find a vertex in the graphRRG where labelling is
-	 * @param nextState
+	 * @param label
 	 * @return
 	 */
 	private Vertex findTheVertex(BDD label) 
@@ -260,7 +261,7 @@ public abstract class RRG {
 		while(it.hasNext()) 
 		{
 			temp 				= it.next();
-			if(Math.abs(temp.getPoint().getX() - p.getX()) < 0.0001  &&  Math.abs(temp.getPoint().getY() - p.getY()) < 0.0001) 
+			if(temp.getPoint().getX() == p.getX()  &&  temp.getPoint().getY() == p.getY())
 			{
 				return temp;
 			}
@@ -269,7 +270,7 @@ public abstract class RRG {
 	}
 	
 	/**
-	 * add a vertex to the {@link #Map map}
+	 * add a vertex to the map
 	 * @param v
 	 */
 	protected void addToMap(Vertex v) {

### 2021-02-24 MotionPlanning/src/modules/UnknownGrid.java

@@ -16,6 +16,7 @@ import com.infomatiq.jsi.SpatialIndex;
 import environment.Environment;
 import environment.Vertex;
 import gnu.trove.TIntProcedure;
+import settings.PlanningSettings;
 
 public class UnknownGrid extends Grid {
 
@@ -74,18 +75,21 @@ public class UnknownGrid extends Grid {
 		if(bestAdviceFrontier.getSecond().getSecond() > bestFrontier.getSecond().getSecond()) {
 			int index = bestAdviceFrontier.getSecond().getFirst();
 			adviceFrontiers.remove(index);
-			System.out.println("Used Advice frontier: " + bestAdviceFrontier.getFirst() + " with IG = " + bestAdviceFrontier.getSecond().getSecond());
+			if((boolean) PlanningSettings.get("debug"))
+				System.out.println("Used Advice frontier: " + bestAdviceFrontier.getFirst() + " with IG = " + bestAdviceFrontier.getSecond().getSecond());
 			return new Pair<Point2D, Integer>(bestAdviceFrontier.getFirst(), bestAdviceFrontier.getSecond().getFirst());
 		}
 		else {
-			System.out.println("Used frontier: " + bestFrontier.getFirst() + " with IG = " + bestFrontier.getSecond().getSecond());
+			if((boolean) PlanningSettings.get("debug"))
+				System.out.println("Used frontier: " + bestFrontier.getFirst() + " with IG = " + bestFrontier.getSecond().getSecond());
 			return new Pair<Point2D, Integer>(bestFrontier.getFirst(), -1);
 		}
 	}
 
 	public void addAdviceFrontier(Point2D p, Point2D currentPosition, int rank) {
 		Pair<Point2D, Integer> f = new Pair<Point2D, Integer>(p, rank);
-		System.out.println("Added advice frontier " + p); 
+		if((boolean) PlanningSettings.get("debug"))
+			System.out.println("Added advice frontier " + p);
 		adviceFrontiers.add(f);
 	}
 	

### 2021-02-24 MotionPlanning/src/modules/UnknownRRG.java

@@ -48,7 +48,7 @@ public class UnknownRRG extends RRG
 	{
 		super(env);
 		this.sensingRadius		= (float) PlanningSettings.get("sensingRadius");
-		this.grid 				= new UnknownGrid(env, (float) PlanningSettings.get("discretizationSize"), graph, tree, treePoints);
+		this.grid 				= new UnknownGrid(env, (float) PlanningSettings.get("gridSize"), graph, tree, treePoints);
 		this.movementBDD 		= new ArrayList<BDD>();
 	}
 	
@@ -375,7 +375,8 @@ public class UnknownRRG extends RRG
 	public Pair<Float, Float> plotGraph(List<DefaultEdge> finalPath)  
 	{
 		if(finalPath != null) {
-			new ShowGraphUnknown(graph, env, movement, finalPath).setVisible(true);
+			if((boolean) PlanningSettings.get("generatePlot"))
+				new ShowGraphUnknown(graph, env, movement, finalPath).setVisible(true);
 			StoreGraphUnknown temp = new StoreGraphUnknown(env, graph, finalPath, movement, "end");
 			return new Pair<Float, Float>(temp.movementLength, temp.remainingPathLength);
 		} else if(! flagFirstMove) {

### 2021-02-24 MotionPlanning/src/planningIO/BuchiAutomataParser.java

@@ -54,7 +54,7 @@ public class BuchiAutomataParser
 				try 
 				{
 					newEdge.andWith(bddDomain[2].ithVar(accSignature.get(0)+1));
-					System.out.println("Added edge in buchi from state "+stateId+" ----"+labelExpr.toString()+"----> "+i+" {"+accSignature.get(0)+"}");
+//					System.out.println("Added edge in buchi from state "+stateId+" ----"+labelExpr.toString()+"----> "+i+" {"+accSignature.get(0)+"}");
 				} catch(NullPointerException E) 
 				{
 					newEdge.andWith(bddDomain[2].ithVar(0));

### 2021-02-24 MotionPlanning/src/planningIO/EnvironmentReader.java

@@ -42,11 +42,9 @@ public class EnvironmentReader
 		p3 	= new Point2D.Float(Float.parseFloat(m.group(5)), Float.parseFloat(m.group(6)));
 		p4 	= new Point2D.Float(Float.parseFloat(m.group(7)), Float.parseFloat(m.group(8)));
 		
-		if((boolean) PlanningSettings.get("verbosity"))
-		{
-			System.out.println("Environment boundaries: " + p1.toString() + ", " + p2.toString() + ", " + p3.toString() + ", " + p4.toString());
-		}
-		
+		if((boolean) PlanningSettings.get("debug"))
+			System.out.println("\nEnvironment boundaries: " + p1.toString() + ", " + p2.toString() + ", " + p3.toString() + ", " + p4.toString());
+
 		
 		
 		// For initial point
@@ -56,22 +54,18 @@ public class EnvironmentReader
         m.find();
 		init 	= new Point2D.Float(Float.parseFloat(m.group(1)), Float.parseFloat(m.group(2)));
 		
-		if ((boolean) PlanningSettings.get("verbosity"))
-		{
-			System.out.println("\n\nInitial Point: " + init.toString());
-		}
-		
+		if ((boolean) PlanningSettings.get("debug"))
+			System.out.println("Initial Point: " + init.toString());
+
 		
 		
 		// For obstacles
-		p 		= Pattern.compile(rectangle);
+		p = Pattern.compile(rectangle);
 		Path2D rect;
-		int i 	= 0;
-		if((boolean) PlanningSettings.get("verbosity"))
-        {
-        	System.out.println("\n\nWalls: ");
-        }
-		
+		int i = 0;
+		if((boolean) PlanningSettings.get("debug"))
+        	System.out.println("\nObstacles: ");
+
 		line = reader.readLine();
 		if(line.equalsIgnoreCase("obstacles")) {
 			line = reader.readLine();
@@ -81,7 +75,8 @@ public class EnvironmentReader
 			if(line.equalsIgnoreCase("see through obstacles")) {
 				break;
 			}
-			System.out.println(line);
+			if((boolean) PlanningSettings.get("debug"))
+				System.out.println(line);
 			try
 			{
 				m = p.matcher(line);
@@ -98,13 +93,7 @@ public class EnvironmentReader
 			rect.lineTo(Float.parseFloat(m.group(1)), Float.parseFloat(m.group(2)));
 			rect.closePath();
 			obstacles.add(rect);
[…]

### 2021-02-24 MotionPlanning/src/planningIO/PropertyReader.java

@@ -28,8 +28,9 @@ public class PropertyReader
 	        propertyString			= propertyReader.readLine();
 	        propertyReader.close();
 		}
-        // Output 
-        System.out.println("\nConstructing automaton for property: " + propertyString + " ...\n\n");
+        // Output
+        if((boolean) PlanningSettings.get("debug"))
+            System.out.print("\nConstructing automaton for property: " + propertyString + " ... ");
         
     	
         // Using SPOT
@@ -39,7 +40,7 @@ public class PropertyReader
 //      ProcessBuilder builder3 = new ProcessBuilder(command,"--deterministic",propertyString);
         
         // Using OWL
-        String command					= "/home/kush/Projects/robotmotionplanning/owl/build/install/owl/bin/ltl2ldba";
+        String command					= "lib/owl/bin/ltl2ldba";
 //        String command 					= "/home/kush/Projects/robotmotionplanning/MotionPlanning/lib/ltl2ldba";
         ProcessBuilder builder1 		= new ProcessBuilder(command, propertyString);
         ProcessBuilder builder2 		= new ProcessBuilder(command, propertyString);
@@ -73,7 +74,8 @@ public class PropertyReader
 			propertyParser		= new BuchiAutomataParser( p1.getInputStream(), 
 					p2.getInputStream(), 
 					apListSystem);
-			System.out.println("done!");
+            if((boolean) PlanningSettings.get("debug"))
+                System.out.println("done!");
 		} catch (ParseException e) 
         {
 			e.printStackTrace();

### 2021-02-24 MotionPlanning/src/planningIO/printing/PrintProductAutomaton.java

@@ -16,7 +16,7 @@ public class PrintProductAutomaton
 	public PrintProductAutomaton(ProductAutomaton productAutomaton, int num) throws Exception
 	{
 		visited 				= new boolean[ProductAutomaton.getNumStates()];
-		writer 					= new BufferedWriter(new FileWriter("foundPoint/productAutomaton"+num+".dot"));
+		writer 					= new BufferedWriter(new FileWriter("temp/productAutomaton"+num+".dot"));
 		
 		this.productAutomaton	= productAutomaton;
 		BDD productAutomatonBDD = productAutomaton.getBDD();

### 2021-02-24 MotionPlanning/src/settings/Initialize.java

@@ -26,11 +26,11 @@ public class Initialize
 	
 	
 	@SuppressWarnings("static-access")
-	public Initialize() throws Exception
+	public Initialize(BDDFactory factory) throws Exception
 	{
-		
-		factory = BuDDyFactory.init(20, (int) PlanningSettings.get("bddFactoryCacheSize"));
-		ProductAutomaton.factory = factory;
+		this.factory = factory;
+//		factory = BuDDyFactory.init(20, (int) PlanningSettings.get("bddFactoryCacheSize"));
+		ProductAutomaton.factory = Initialize.factory;
         ProductAutomaton.threshold = (int) PlanningSettings.get("transitionThreshold");
 		
 //		Read environment

### 2021-02-24 MotionPlanning/src/settings/PlanningSettings.java

@@ -13,7 +13,7 @@ public class PlanningSettings
 	public static final String COLOUR_TYPE  = "c";
 	
 	
-	public static final String VERBOSITY					=	"verbosity";
+	public static final String BIAS_PROB					=	"biasProb";
 	public static final String BDD_FACTORY_CACHE_SIZE		=	"bddFactoryCacheSize";
 	public static final String USE_SPOT						= 	"useSpot";
 	public static final String MAX_LEVEL_TRANSITION			=	"maxLevelTransition";
@@ -22,14 +22,34 @@ public class PlanningSettings
 	public static final String ETA							=	"eta";
 	public static final String SENSING_RADIUS				=	"sensingRadius";
 	public static final String BATCH_SIZE					=  	"batchSize";
-	public static final String DISCRETIZATION_SIZE			=  	"discretizationSize";
+	public static final String GRID_SIZE 					=  	"gridSize";
 	public static final String USE_ADVICE					=	"useAdvice";
 	public static final String FIRST_EXPL_THEN_PLAN			=	"firstExplThenPlan";
 	public static final String DEBUG						=	"debug";
-	public static final String INPUT_FILE					=	"inputFile"; 
-	public static final String ONLY_OPAQUE_OBSTACLES		=	"onlyOpaqueObstacles"; 
-	
-	
+	public static final String INPUT_FILE					=	"inputFile";
+	public static final String ONLY_OPAQUE_OBSTACLES		=	"onlyOpaqueObstacles";
+	public static final String NUMBER_OF_RUNS				=	"numberOfRuns";
+	public static final String GENERATE_PLOT				=	"generatePlot";
+
+	public static void outputParameters() {
+		System.out.println("Parameters values: ");
+
+		if((boolean) get("randomEnv"))
+			System.out.println("Environment: Random");
+		else
+			System.out.println("Environment: " + (String) get("inputFile"));
+
+		if((boolean) get("useAdvice"))
+			System.out.println("Use Advice: Yes");
+		else
+			System.out.println("Use Advice: No");
+
+		System.out.println("Sensing radius: " + (Float) get("sensingRadius"));
+		System.out.println("Batch size: " + (Integer) get("batchSize"));
+		System.out.println("Discretization cell size: " + (Float) get("gridSize"));
[…]

### 2021-02-24 MotionPlanning/temp/avg.py

@@ -1,0 +1,27 @@
+import pandas as pd
+
+filename = "temp/output/output.csv"
+csv_reader = pd.read_csv(filename, delimiter=',')
+
+iterations = csv_reader['iterations']
+sampled = csv_reader['sampled']
+rrgsize = csv_reader['rrgsize']
+movement = csv_reader['movement']
+remaining = csv_reader['remaining']
+time = csv_reader['time']
+
+iterations_avg = iterations.sum() / iterations.size
+sampled_avg = sampled.sum() / sampled.size
+rrgsize_avg = rrgsize.sum() / rrgsize.size
+movement_avg = movement.sum() / movement.size
+remaining_avg = remaining.sum() / remaining.size
+time_avg = time.sum() / time.size
+
+print("\n======= Averages computed over " + str(time.size) + " run(s) =======")
+print("Number of iterations | " + str(iterations_avg))
+print("Total sampled points | " + str(sampled_avg))
+print("RRG size             | " + str(rrgsize_avg))
+print("Movement length      | " + str(movement_avg))
+print("Remaining length     | " + str(remaining_avg))
+print("Time taken (in ms)   | " + str(time_avg))
+print("===============================================")

### 2021-02-24 MotionPlanning/temp/plotGraph.py

@@ -1,0 +1,87 @@
+import numpy as np
+import networkx as nx
+
+import matplotlib
+# matplotlib.use('TkAgg')
+from matplotlib import pyplot as plt
+import csv
+
+plt.figure()
+plt.gca().set_aspect('equal', 'datalim')
+# axs.set_xlim((0, 6))
+# axs.set_ylim((0, 3))
+plt.xlim(-1,7)
+plt.ylim(-1,4)
+
+import pandas as pd
+
+def plotfile(filename, col, dashed, thick):
+    csv_reader = pd.read_csv(filename, delimiter=',')
+    
+    x1 = csv_reader['x1']
+    x2 = csv_reader['x2']
+    y1 = csv_reader['y1']
+    y2 = csv_reader['y2']
+    # plt.plot((x1, x2), (y1, y2), color = col)
+        
+
+    # for i in range(len(x1)):
+    if(dashed):
+        plt.plot((x1, x2), (y1, y2), color = col, linestyle = 'dashed')
+    elif(thick):
+        plt.plot((x1, x2), (y1, y2), color = col, linewidth = 3)
+    else: 
+        plt.plot((x1, x2), (y1, y2), color = col)
+
+
+
+    # with open(filename) as csv_file:
+        # line_count = 0
+        # for row in csv_reader:
+        #     if line_count == 0:
+        #         line_count += 1
+    # with open(filename) as csv_file:
+        #     else:
+        #         # x = [float(row[0]), float(row[2])]
+        #         # y = [float(row[1]), float(row[3])]
+        #         if(dashed):
+        #             plt.plot(x, y, color = col, linestyle = 'dashed')
+        #         elif(thick):
+        #             plt.plot(x, y, color = col, linewidth = 2)
+        #         else: 
+        #             plt.plot(x, y, color = col)
+        #         line_count += 1
+
+plotfile("firstMove.csv", "#20168a", False, False)
+plotfile("obstacles.csv", "red", False, True)
+plotfile("labels.csv", "black", True, False)
+circle = plt.Circle((0.1,1.5), 1.0, color = "orange", fill = False, linewidth = 2)
+plt.plot([0.1], [1.5], marker='o', markersize=7, color="#33e026", )
+plt.gca().add_artist(circle)
+plt.show()
+
+plotfile("bin.csv", "#20168a", False, False)
+plotfile("bin-movement.csv", "#33e026", False, True)
+plotfile("obstacles.csv", "red", False, True)
[…]

### 2021-02-24 MotionPlanning/lib/owl/javadoc/member-search-index.js
a

@@ -1,1 +1,0 @@
-memberSearchIndex = [{"p":"owl.grammar","c":"LTLLexer","l":"_ATN"},{"p":"owl.grammar","c":"LTLParser","l":"_ATN"},{"p":"owl.grammar","c":"LTLLexer","l":"_decisionToDFA"},{"p":"owl.grammar","c":"LTLParser","l":"_decisionToDFA"},{"p":"owl.grammar","c":"LTLLexer","l":"_serializedATN"},{"p":"owl.grammar","c":"LTLParser","l":"_serializedATN"},{"p":"owl.grammar","c":"LTLLexer","l":"_sharedContextCache"},{"p":"owl.grammar","c":"LTLParser","l":"_sharedContextCache"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"AbstractCachedStatesAutomaton()","url":"%3Cinit%3E()"},{"p":"owl.automaton","c":"AbstractImplicitAutomaton","l":"AbstractImplicitAutomaton(ValuationSetFactory, Collection<S>, A)","url":"%3Cinit%3E(owl.factories.ValuationSetFactory,java.util.Collection,A)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeMapVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeMapVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeMapVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeMapVisitor)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeTreeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeTreeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeTreeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeTreeVisitor)"},{"p":"owl.automaton","c":"AbstractCachedStatesAutomaton","l":"accept(Automaton.EdgeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(Automaton.EdgeVisitor<S>)","url":"accept(owl.automaton.Automaton.EdgeVisitor)"},{"p":"owl.automaton","c":"Automaton","l":"accept(
[…]

### 2021-02-24 MotionPlanning/lib/owl/javadoc/package-search-index.js

@@ -1,1 +1,0 @@
-packageSearchIndex = [{"l":"All Packages","url":"allpackages-index.html"},{"l":"owl.automaton"},{"l":"owl.automaton.acceptance"},{"l":"owl.automaton.acceptance.optimizations"},{"l":"owl.automaton.algorithms"},{"l":"owl.automaton.edge"},{"l":"owl.automaton.output"},{"l":"owl.automaton.transformations"},{"l":"owl.automaton.util"},{"l":"owl.cinterface"},{"l":"owl.collections"},{"l":"owl.factories"},{"l":"owl.factories.jbdd"},{"l":"owl.game"},{"l":"owl.game.algorithms"},{"l":"owl.game.output"},{"l":"owl.grammar"},{"l":"owl.ltl"},{"l":"owl.ltl.algorithms"},{"l":"owl.ltl.ltlf"},{"l":"owl.ltl.parser"},{"l":"owl.ltl.rewriter"},{"l":"owl.ltl.robust"},{"l":"owl.ltl.util"},{"l":"owl.ltl.visitors"},{"l":"owl.run"},{"l":"owl.run.modules"},{"l":"owl.run.parser"},{"l":"owl.translations"},{"l":"owl.translations.canonical"},{"l":"owl.translations.delag"},{"l":"owl.translations.dpa2safety"},{"l":"owl.translations.dra2dpa"},{"l":"owl.translations.ltl2dpa"},{"l":"owl.translations.ltl2dra"},{"l":"owl.translations.ltl2ldba"},{"l":"owl.translations.ltl2nba"},{"l":"owl.translations.mastertheorem"},{"l":"owl.translations.modules"},{"l":"owl.translations.nba2dpa"},{"l":"owl.translations.nba2ldba"},{"l":"owl.translations.rabinizer"},{"l":"owl.util"},{"l":"owl.util.annotation"}]

### 2021-02-24 MotionPlanning/lib/owl/javadoc/type-search-index.js

@@ -1,1 +1,0 @@
-typeSearchIndex = [{"p":"owl.automaton","l":"AbstractCachedStatesAutomaton"},{"p":"owl.automaton","l":"AbstractImplicitAutomaton"},{"p":"owl.automaton.acceptance.optimizations","l":"AcceptanceOptimizations"},{"p":"owl.automaton.acceptance.optimizations","l":"AcceptanceOptimizations.AcceptanceOptimizationTransformer"},{"p":"owl.game.output","l":"Aig"},{"p":"owl.game.output","l":"AigConsumer"},{"p":"owl.game.output","l":"AigerPrinter"},{"p":"owl.game.output","l":"AigFactory"},{"p":"owl.game.output","l":"AigPrintable"},{"p":"owl.automaton.acceptance","l":"AllAcceptance"},{"l":"All Classes","url":"allclasses-index.html"},{"p":"owl.grammar","l":"LTLParser.AndExpressionContext"},{"p":"owl.translations.ltl2ldba","l":"AnnotatedLDBA"},{"p":"owl.automaton.util","l":"AnnotatedState"},{"p":"owl.automaton.util","l":"AnnotatedStateOptimisation"},{"p":"owl.translations.mastertheorem","l":"AsymmetricEvaluatedFixpoints"},{"p":"owl.translations.ltl2ldba","l":"AsymmetricLDBAConstruction"},{"p":"owl.translations.ltl2ldba","l":"AsymmetricProductState"},{"p":"owl.translations.ltl2dpa","l":"AsymmetricRankingState"},{"p":"owl.grammar","l":"LTLParser.AtomExpressionContext"},{"p":"owl.automaton","l":"Automaton"},{"p":"owl.automaton","l":"AutomatonFactory"},{"p":"owl.automaton","l":"AutomatonOperations"},{"p":"owl.automaton","l":"AutomatonReader"},{"p":"owl.run.modules","l":"OutputWriters.AutomatonStats"},{"p":"owl.automaton","l":"AutomatonUtil"},{"p":"owl.automaton","l":"Views.AutomatonView"},{"p":"owl.ltl","l":"Biconditional"},{"p":"owl.grammar","l":"LTLParser.BinaryExpressionContext"},{"p":"owl.grammar","l":"LTLParser.BinaryOpContext"},{"p":"owl.grammar","l":"LTLParser.BinaryOperationContext"},{"p":"owl.ltl","l":"Formula.BinaryTempor
[…]

### 2021-03-03 MotionPlanning/experiments.sh
Stable

@@ -1,0 +1,94 @@
+#!/bin/bash
+
+arg1="$1"
+shift
+arg2="$1"
+shift
+
+
+mkdir -p results
+rm -rf results/*
+
+mkdir -p results/environments
+mkdir -p results/opaque
+mkdir -p results/see_through
+
+mkdir -p results/opaque/separate_with_advice
+mkdir -p results/opaque/separate_without_advice
+mkdir -p results/opaque/together_with_advice
+mkdir -p results/opaque/together_without_advice
+
+mkdir -p results/see_through/separate_with_advice
+mkdir -p results/see_through/separate_without_advice
+mkdir -p results/see_through/together_with_advice
+mkdir -p results/see_through/together_without_advice
+
+
+mkdir -p results/opaque/separate_with_advice/output
+mkdir -p results/opaque/separate_without_advice/output
+mkdir -p results/opaque/together_with_advice/output
+mkdir -p results/opaque/together_without_advice/output
+
+mkdir -p results/see_through/separate_with_advice/output
+mkdir -p results/see_through/separate_without_advice/output
+mkdir -p results/see_through/together_with_advice/output
+mkdir -p results/see_through/together_without_advice/output
+
+
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/see_through/separate_with_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/see_through/separate_without_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/see_through/together_with_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/see_through/together_without_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/opaque/separate_with_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/opaque/separate_without_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/opaque/together_with_advice/output.csv
+echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/opaque/together_without_advice/output.csv
+
+
+echo "Starting simulations..."
+
+for (( i=1; i <= $arg1; i++ ))
+do
+   dir=results/environments/env$i
+   mkdir -p $dir
+   dir=results/environments/env$i/temp
[…]

### 2021-03-03 MotionPlanning/planner.sh

@@ -1,145 +1,0 @@
-#!/bin/bash
-
-arg1="$1"
-shift
-
-mkdir -p results
-rm -rf results/*
-
-# mkdir -p results/separate_with_advice
-# mkdir -p results/separate_without_advice
-# mkdir -p results/together_with_advice
-# mkdir -p results/together_without_advice
-
-# mkdir -p results/separate_with_advice/output
-# mkdir -p results/separate_without_advice/output
-# mkdir -p results/together_with_advice/output
-# mkdir -p results/together_without_advice/output
-
-# touch results/separate_with_advice/output.csv
-# touch results/separate_without_advice/output.csv
-# touch results/together_with_advice/output.csv
-# touch results/together_without_advice/output.csv
-
-# echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/separate_with_advice/output.csv
-# echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/separate_without_advice/output.csv
-# echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/together_with_advice/output.csv
-# echo "iterations,sampled,rrgsize,movement,remaining,total,time," > results/together_without_advice/output.csv
-
-
-# echo "Separate with advice:"
-# for (( i=1; i <= $arg1; i++ ))
-# do
-#    echo "Current run: $i"
-#    java -jar planning.jar Examples/6rooms/example --property-file myAut.hoa --only-opaque-obstacles --first-expl-then-plan --output-directory results/separate_with_advice/ > results/separate_with_advice/output/output$i.txt
-# done
-# echo "Output written to results/separate_with_advice/final_result.txt"
-# echo " "
-# python3 avg.py results/separate_with_advice/output.csv > results/separate_with_advice/final_result.txt
-
-
-
-# echo "Separate without advice:"
-# for (( i=1; i <= $arg1; i++ ))
-# do
-#    echo "Current run: $i"
-#    java -jar planning.jar Examples/6rooms/example --property-file myAut.hoa --only-opaque-obstacles --first-expl-then-plan --no-advice --output-directory results/separate_without_advice/ > results/separate_without_advice/output/output$i.txt
-# done
-# echo "Output written to results/separate_without_advice/final_result.txt"
-# echo " "
-# python3 avg.py results/separate_without_advice/output.csv > results/separate_without_advice/final_result.txt
-
-
-
-# echo "Together with advice:"
-# for (( i=1; i <= $arg1; i++ ))
-# do
-#    echo "Current run: $i"
[…]

### 2021-03-03 MotionPlanning/plotGraph.py

@@ -50,7 +50,7 @@ dir_together = sys.argv[1] + "together/"
 
 set_axes()
 plot_file(dir_together + "firstMove.csv", "#20168a", False, 0)
-plot_file(dir_together + "obstacles.csv", "red", False, 1)
+plot_file(dir_together + "obstacles.csv", "black", False, 1)
 plot_file(dir_together + "labels.csv", "black", True, 0)
 circle = plt.Circle((0.1,1.5), 1.0, color = "orange", fill = False, linewidth = 5)
 plt.gca().add_artist(circle)
@@ -61,7 +61,7 @@ plt.savefig(dir_together + "tog_firstMove.png", format = 'png', dpi = 300, bbox_
 set_axes()
 plot_file(dir_together + "bin.csv", "#20168a", False, 0)
 plot_file(dir_together + "bin-movement.csv", "#33e026", False, 2)
-plot_file(dir_together + "obstacles.csv", "red", False, 1)
+plot_file(dir_together + "obstacles.csv", "black", False, 1)
 plot_file(dir_together + "labels.csv", "black", True, 0)
 plot_last_point(dir_together + "bin-movement.csv", "#33e026")
 plt.savefig(dir_together + "tog_bin.png", format = 'png', dpi = 300, bbox_inches='tight')
@@ -70,7 +70,7 @@ plt.savefig(dir_together + "tog_bin.png", format = 'png', dpi = 300, bbox_inches
 set_axes()
 plot_file(dir_together + "room.csv", "#20168a", False, 0)
 plot_file(dir_together + "room-movement.csv", "#33e026", False, 2)
-plot_file(dir_together + "obstacles.csv", "red", False, 1)
+plot_file(dir_together + "obstacles.csv", "black", False, 1)
 plot_file(dir_together + "labels.csv", "black", True, 0)
 plot_last_point(dir_together + "room-movement.csv", "#33e026")
 plt.savefig(dir_together + "tog_room.png", format = 'png', dpi = 300, bbox_inches='tight')
@@ -79,11 +79,11 @@ plt.savefig(dir_together + "tog_room.png", format = 'png', dpi = 300, bbox_inche
 set_axes()
 plot_file(dir_together + "end.csv", "#20168a", False, 0)
 plot_file(dir_together + "end-movement.csv", "#33e026", False, 2)
-plot_file(dir_together + "end-finalpath.csv", "orange", False, 2)
[…]

### 2021-03-03 MotionPlanning/ran_env_gen.py

@@ -36,8 +36,11 @@ f_env.write("[(3.95,0), (4.05,0), (4.05,0.95), (3.95,0.95)]\n")
 f_env.write("[(3.95,2.05), (4.05,2.05), (4.05,3), (3.95,3)]\n")
 
 # Tables
-if(sys.argv[2] == "--see-through"):
-    f_env.write("see through obstacles\n")
+try:
+    if(sys.argv[2] == "--see-through"):
+        f_env.write("see through obstacles\n")
+except IndexError:
+    pass
 
 l_bin = 0.2
 w_bin = 0.2

### 2021-03-03 MotionPlanning/src/Planning.java

@@ -84,11 +84,8 @@ public class Planning
 	 */
 	void printOutput() throws Exception 
 	{
-		if((boolean) PlanningSettings.get("debug")){
-			System.out.print("Exporting product automaton ... ");
-	    	productAutomaton.createDot(iterationNumber);
-			System.out.println("done");
-		}
+
+
 		Pair<Float, Float> length = new Pair<Float, Float>(0f, 0f);
 		if((boolean) PlanningSettings.get("firstExplThenPlan")) {
 			length.setFirst(moveLength);
@@ -103,8 +100,14 @@ public class Planning
 			pathLength = length.getSecond();
 		}
 		totalTime = System.nanoTime() - beginTime;
-        
-        int adviceSamples = 0;
+
+		if((boolean) PlanningSettings.get("debug")){
+			System.out.print("Exporting product automaton ... ");
+			productAutomaton.createDot(iterationNumber);
+			System.out.println("done");
+		}
+
+		int adviceSamples = 0;
 		for( int k=0; k<10; k++ )
 			adviceSamples += rrg.adviceSampled[k];
 
@@ -210,7 +213,7 @@ public class Planning
         while( true )
         {
         	if( debug ) {
-	        	if( iterationNumber == 1 ) {
+	        	if( iterationNumber == 100 ) {
 //	        		urrg.grid.printDiscretization();
 //	        		urrg.grid.printFrontiers();
 	        		break;
@@ -300,7 +303,10 @@ public class Planning
         boolean computePath = true;
 
 		System.out.print("Starting planning ... ");
-		while( true ){
+		double timeout = (double) PlanningSettings.get("timeout");
+		double currentTime = System.nanoTime()/1000000;
+
+		while( currentTime - (beginTime/1000000) < timeout ){
 			iterationNumber++;
 
 			if((boolean) PlanningSettings.get("debug"))
@@ -316,20 +322,37 @@ public class Planning
         		computePath = true;
         	learn(transitions);
         	
-        	startTime = System.nanoTime();
-        	if( computePath ) {
-        		ArrayList<BDD> path	= productAutomaton.findAcceptingPath(movementBDD);
-        		if( path != null ) {
-        			System.out.println("Path found :D");
-					if((boolean) PlanningSettings.get("debug"))
-						productAutomaton.printPath(path);
-                	finalPath = rrg.liftPath(path);
-            		break;
-            	}
-        	}
[…]

### 2021-03-03 MotionPlanning/src/mainCL.java

@@ -26,7 +26,7 @@ public class mainCL {
 		};
 
 		for(int i=0; i < numOfRuns; i++) {
-			BDDFactory factory = BuDDyFactory.init(600000, (int) PlanningSettings.get("bddFactoryCacheSize"));
+			BDDFactory factory = BuDDyFactory.init(1000000, (int) PlanningSettings.get("bddFactoryCacheSize"));
 			System.out.println("Starting initialization ... ");
 			Planning plan = new Planning(factory);
 			System.out.println("Initialization complete\n\n");
@@ -51,6 +51,9 @@ public class mainCL {
 					sum[j] = new Float((Float) sum[j] + (Float) data[j]);
 				}
 				else if(j<7) {
+					BufferedWriter w = new BufferedWriter(new FileWriter("temp/timeout.txt"));
+					w.write((int) (((Double) data[j])/1000000) + "");
+					w.close();
 					writer.write(((Double) data[j])/1000000 + ",");
 					sum[j] = new Double((Double) sum[j] + (Double) data[j]);
 				}

### 2021-03-03 MotionPlanning/src/modules/KnownRRG.java

@@ -71,18 +71,20 @@ public class KnownRRG extends RRG {
 					updateRrgRadius(); 
 					addSymbolicTransitions(xNearest2D, xNew2D);
 					addGraphEdge(xNearest2D, xNew2D);
-					
-					try {
-						if(! flagBin && ! Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars+7)).isZero()) {
-							StoreGraphKnown temp = new StoreGraphKnown(graph, grid.getGraph(), grid.getMovement(), (String) PlanningSettings.get("outputDirectory") + "bin");
-							flagBin = true;
+
+					if((boolean) PlanningSettings.get("exportPlotData")) {
+						try {
+							if (!flagBin && !Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars + 7)).isZero()) {
+								StoreGraphKnown temp = new StoreGraphKnown(graph, grid.getGraph(), grid.getMovement(), (String) PlanningSettings.get("outputDirectory") + "bin");
+								flagBin = true;
+							}
+							if (!flagRoom && !Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars + 4)).isZero()) {
+								StoreGraphKnown temp = new StoreGraphKnown(graph, grid.getGraph(), grid.getMovement(), (String) PlanningSettings.get("outputDirectory") + "room");
+								flagRoom = true;
+							}
+						} catch (Exception e1) {
+							e1.printStackTrace();
 						}
-						if(! flagRoom && ! Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars+4)).isZero()) {
-							StoreGraphKnown temp = new StoreGraphKnown(graph, grid.getGraph(), grid.getMovement(), (String) PlanningSettings.get("outputDirectory") + "room");
-							flagRoom = true;
-						}
-					} catch (Exception e1) {
-						e1.printStackTrace();
 					}
 				
 					tree.nearestN(xNew, 

### 2021-03-03 MotionPlanning/src/modules/UnknownGrid.java

@@ -127,10 +127,10 @@ public class UnknownGrid extends Grid {
 		int bestIndex = 0;
 		for(int i=0;i<frontiers.size();i++) {
 			tempPoint = findFrontierBestPoint(frontiers.get(i));
-//			if(tempPoint.equals(xRobot)) {
+			if(tempPoint.equals(xRobot)) {
 //				printDiscretization();
-//				continue;
-//			}
+				continue;
+			}
 			currentIG = (float) frontiers.get(i).size() / computeDistance(graph, xRobot, tempPoint);
 //			if(distance(centers.get(i), xRobot) < distance(closest, xRobot)) {
 			if(currentIG > bestIG) {

### 2021-03-03 MotionPlanning/src/modules/UnknownRRG.java

@@ -136,20 +136,20 @@ public class UnknownRRG extends RRG
 					addSymbolicTransitions(xNearest2D, xNew2D);
 					addGraphEdge(xNearest2D, xNew2D);
 										
-//					plotting the first time it sees a bin
-					try {
-						if(! flagBin && ! Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars+7)).isZero()) {
-							StoreGraphUnknown temp = new StoreGraphUnknown(graph, movement, (String) PlanningSettings.get("outputDirectory") + "bin");
-							flagBin = true;
+					if((boolean) PlanningSettings.get("exportPlotData")) {
+						try {
+							if (!flagBin && !Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars + 7)).isZero()) {
+								StoreGraphUnknown temp = new StoreGraphUnknown(graph, movement, (String) PlanningSettings.get("outputDirectory") + "bin");
+								flagBin = true;
+							}
+							if (!flagRoom && !Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars + 4)).isZero()) {
+								StoreGraphUnknown temp = new StoreGraphUnknown(graph, movement, (String) PlanningSettings.get("outputDirectory") + "room");
+								flagRoom = true;
+							}
+						} catch (Exception e1) {
+							e1.printStackTrace();
 						}
-						if(! flagRoom && ! Environment.getLabelling().getLabel(xNew2D).and(ProductAutomaton.factory.ithVar(ProductAutomaton.varsBeforeSystemVars+4)).isZero()) {
-							StoreGraphUnknown temp = new StoreGraphUnknown(graph, movement, (String) PlanningSettings.get("outputDirectory") + "room");
-							flagRoom = true;
-						}
-					} catch (Exception e1) {
-						e1.printStackTrace();
 					}
-					
 					tree.nearestN(xNew, 
 							new TIntProcedure() // For each neighbour of 'xNew' in the given radius, execute this method
 							{
@@ -267,10 +267,12 @@ public class UnknownRRG extends RRG
 		if(rank != -1 && productAutomaton.sampledTransitions.and(transition).isZero()) // if sampled from advice and have not sampled it before
[…]

### 2021-03-03 MotionPlanning/src/planningIO/EnvironmentReader.java

@@ -103,7 +103,7 @@ public class EnvironmentReader
 		// For see through obstacles
 		p = Pattern.compile(rectangle);
 		i = 0;
-		if((boolean) PlanningSettings.get("debug"))
+		if((boolean) PlanningSettings.get("debug") && ! (boolean) PlanningSettings.get("onlyOpaqueObstacles"))
         	System.out.println("\nSee through Obstacles: ");
 
 		while(line != null)
@@ -124,8 +124,12 @@ public class EnvironmentReader
 			rect.lineTo(Float.parseFloat(m.group(7)), Float.parseFloat(m.group(8)));
 			rect.lineTo(Float.parseFloat(m.group(1)), Float.parseFloat(m.group(2)));
 			rect.closePath();
-			seeThroughObstacles.add(rect);
-			
+			if((boolean) PlanningSettings.get("onlyOpaqueObstacles")) {
+				obstacles.add(rect);
+			}
+			else {
+				seeThroughObstacles.add(rect);
+			}
 			if((boolean) PlanningSettings.get("debug"))
             	System.out.println(line);
 

### 2021-03-03 MotionPlanning/temp/avg.py

@@ -1,27 +1,0 @@
-import pandas as pd
-
-filename = "temp/output/output.csv"
-csv_reader = pd.read_csv(filename, delimiter=',')
-
-iterations = csv_reader['iterations']
-sampled = csv_reader['sampled']
-rrgsize = csv_reader['rrgsize']
-movement = csv_reader['movement']
-remaining = csv_reader['remaining']
-time = csv_reader['time']
-
-iterations_avg = iterations.sum() / iterations.size
-sampled_avg = sampled.sum() / sampled.size
-rrgsize_avg = rrgsize.sum() / rrgsize.size
-movement_avg = movement.sum() / movement.size
-remaining_avg = remaining.sum() / remaining.size
-time_avg = time.sum() / time.size
-
-print("\n======= Averages computed over " + str(time.size) + " run(s) =======")
-print("Number of iterations | " + str(iterations_avg))
-print("Total sampled points | " + str(sampled_avg))
-print("RRG size             | " + str(rrgsize_avg))
-print("Movement length      | " + str(movement_avg))
-print("Remaining length     | " + str(remaining_avg))
-print("Time taken (in ms)   | " + str(time_avg))
-print("===============================================")

### 2021-03-03 MotionPlanning/temp/plotGraph.py

@@ -1,87 +1,0 @@
-import numpy as np
-import networkx as nx
-
-import matplotlib
-# matplotlib.use('TkAgg')
-from matplotlib import pyplot as plt
-import csv
-
-plt.figure()
-plt.gca().set_aspect('equal', 'datalim')
-# axs.set_xlim((0, 6))
-# axs.set_ylim((0, 3))
-plt.xlim(-1,7)
-plt.ylim(-1,4)
-
-import pandas as pd
-
-def plotfile(filename, col, dashed, thick):
-    csv_reader = pd.read_csv(filename, delimiter=',')
-    
-    x1 = csv_reader['x1']
-    x2 = csv_reader['x2']
-    y1 = csv_reader['y1']
-    y2 = csv_reader['y2']
-    # plt.plot((x1, x2), (y1, y2), color = col)
-        
-
-    # for i in range(len(x1)):
-    if(dashed):
-        plt.plot((x1, x2), (y1, y2), color = col, linestyle = 'dashed')
-    elif(thick):
-        plt.plot((x1, x2), (y1, y2), color = col, linewidth = 3)
-    else: 
-        plt.plot((x1, x2), (y1, y2), color = col)
-
-
-
-    # with open(filename) as csv_file:
-        # line_count = 0
-        # for row in csv_reader:
-        #     if line_count == 0:
-        #         line_count += 1
-    # with open(filename) as csv_file:
-        #     else:
-        #         # x = [float(row[0]), float(row[2])]
-        #         # y = [float(row[1]), float(row[3])]
-        #         if(dashed):
-        #             plt.plot(x, y, color = col, linestyle = 'dashed')
-        #         elif(thick):
-        #             plt.plot(x, y, color = col, linewidth = 2)
-        #         else: 
-        #             plt.plot(x, y, color = col)
-        #         line_count += 1
-
-plotfile("firstMove.csv", "#20168a", False, False)
-plotfile("obstacles.csv", "red", False, True)
-plotfile("labels.csv", "black", True, False)
-circle = plt.Circle((0.1,1.5), 1.0, color = "orange", fill = False, linewidth = 2)
-plt.plot([0.1], [1.5], marker='o', markersize=7, color="#33e026", )
-plt.gca().add_artist(circle)
-plt.show()
-
-plotfile("bin.csv", "#20168a", False, False)
-plotfile("bin-movement.csv", "#33e026", False, True)
-plotfile("obstacles.csv", "red", False, True)
[…]

### 2021-03-03 Paper/1_Introduction.tex
After RSS submission

\section{Introduction}

%Sampling-based motion planning is a well-developed area with plenty of applications \todo{cite some survey} and techniques available, for instance, RRT, RRG \todo{list and cite relevant stuff}.

Motion planning with Linear Temporal Logic (LTL) mission specifications aims for consideration of richer objectives than the traditional A-to-B motion planning. Examples of such objectives include periodic surveillance, request-response, or sequencing. Successful approaches to the problem range from using various cell decomposition techniques, to creating roadmaps abstracting the environment and to sampling-based motion planning. Motion planning with LTL missions is, however, much more challenging in \emph{a priori} unknown environments: efficient treatment of LTL specifications may require exploiting semantic and spatio-temporal dependencies between features of the environment, which are typically unknown beforehand. As an example, consider that we would like a robot to check all waste bins in all offices in an office environment. When finding the first bin, the robot may realize it was next to a desk. While looking for the bin in the next office, it is most natural that the robot starts exploring again next to the desk. At the same time, due to the potential complexity of the environment, it is not desirable to stick fully to all of the observed semantic and spatio-temporal correlations as not all of them are relevant for the specification satisfaction.

%However, techniques tailored or even just applicable to the setting with rich specifications, such as formulae of Linear Temporal Logic (LTL) are scarce \todo{I'm not sure about it (for LTL motion planning there is probably plenty, but sampling-based? if still plenty then we'd restrict already here to partially known labellings); cites}.
%The difficulties arising due to rich logical specifications are twofold.
%Firstly, it is hard to find a satisfying path by a simple random exploration of the state space.
%Indeed, a very particular (and thus rare order) of events may be required, followed by an infinite repetition of its (possibly non-trivial) modifications as the plan does not depend on the current state only but also on the history.
%Secondly, efficient treatment of specifications in temporal logic may require exploiting semantic and spatio-temporal dependencies between atomic propositions, which are typically unknown.
%Moreover, even if there is some domain knowledge or learnt experience, it is hard to reflect.

%\begin{example*}

% As an example, consider an LTL goal specification $\F(\mathit{living\_room}\wedge\F \mathit{kitchen}) \wedge \F\mathit{kitchen}$ expressing that

% encoding a sequence of, e.g., collecting A and then B (B's before A do not count), and a property $\psi=\F c$. % $\psi=\F(d\wedge \F e)$ for switching off the collector and then switching off entirely.
% For the specification $\varphi\wedge\psi$, it might seem sensible to take $c$ if the opportunity occurs, so that we make progress w.r.t.~$\psi$.
% However, we might experimentally observe that we see no $a$'s and $b$'s occur after $c$ (for instance since $c$ denotes switching off the collector), while otherwise $b$'s are frequent if no $c$ is taken.
% As a consequence of this learnt conjecture, we might rather counter-intuitively postpone any progress with $c$, wait for an $a$, then take the abundant $b$ and only afterwards take $c$.
% While such a plan might be still discovered by random trials, if the sequences are longer then we need this kind of learning and reasoning to get reasonably high chance thereof.
% %\end{example}

In this paper, we focus on sampling-based motion planning with missions specified with the syntactically co-safe fragment of LTL (scLTL), and with the robot deployed in \emph{a priori} unknown environments.
%we address both difficulties in the setting of sampling-based motion planning with LTL specifications where, moreover, the atomic-proposition labelling may not be completely known for the whole state space upfront, but only for some sensing radius.
The key idea of our approach is, on the conceptual level, to make the sampling \emph{guided} by a semantic abstraction of the system and by the specification. 
%moreover, \emph{in connection to the system}.
%In our example, the property might guide us to take $c$ since we are waiting for it, but the achieved knowledge of the system w.r.t.\ this property dissuades us from doing so.
The overview of our algorithm is depicted in Fig.~\ref{fig:guide}. We extend the RRG algorithm with learning and biasing; we iteratively learn a semantic abstraction of the system from the gradually growing RRG graph and compose it with an automaton representation of the specification into a so-called product. The product is used to bias sampling in RRG, i.e. to exploit the semantic and spatio-temporal dependencies of features in the environment as well as their relation to satisfying the desired specification.

%On the technical level, we \texttt{learn} a %\emph{semantic abstraction of the system}  and use its \emph{product with the property automaton} to \texttt{bias} further sampling.
%The states of the semantic abstraction reflect (only) the logical validity of \todo{atomic propositions}; its transitions are learnt from the samples, which are anyway generated in the RRG-based planning.

% \begin{figure}
%     \centering
%     \includegraphics[scale=0.6]{scheme}
%     \caption{Guidance scheme \todo{think of a good way to draw this}}
%     \label{fig:scheme}
% \end{figure}
\input{figures/guidance_scheme}

%The specification is thus exploited twice to improve the guidance: (1) We reflect the property itself. (2) We can build the product and reflect the particularities of the (unknown) system \emph{behaviour with respect to the property}.

Compared to the naive two-step approach, which first explores the environment and then plans a trajectory that satisfies the mission, our approach (i) performs both tasks at once and, moreover, (ii) allows mutual exchange of information between the two tasks. 
We show that these two improvements shorten the length of the executed path significantly.
We achieve this while maintaining similar computation time, which will, in reality, be negligible as the robot can execute the algorithm in real-time while navigating in the environment.
Our contribution can be summarized as follows:
\begin{itemize}
    \item We propose a method to learn a semantic abstraction of the system, suitable for planning with scLTL missions. 
    \item We exploit the learned semantic abstraction and, together with consideration of the specification, we bias the growth of the RRG graph towards promising regions (in terms of making progress towards the specification satisfaction).
    \item We experimentally show that the loop between sampling and learning leads to better planning in terms of shorter trajectories when compared to the naive two-step approach. The results indicate more than $50\%$ savings.
    %, and (2) more goal-oriented and thus more explainable plans.
\end{itemize}

The paper is organized as follows. Sec.~\ref{sec:related} introduces relevant related work, and Sec.~\ref{sec:prel} describes preliminary tools needed for the remainder of the paper. The problem is formally defined in Sec.~\ref{sec:prob}, which is followed by the proposed solution and analysis in Sec.~\ref{sec:sol}. Lastly, a case study is presented in Sec.~\ref{sec:cases}, with conclusions and future work in Sec.~\ref{sec:conclusion}.

%\todo{Mention somewhere that partitioning according APs is a very simple semantic abstraction -- tells us only where interesting properties are and how they neighbor with each other}

\subsection{Related Work} \label{sec:related}
One of the first works to propose the use of a sampling-based motion planning algorithm to find a trajectory that satisfies a temporal logic specification is \cite{karaman2009sampling}. In that work, the authors propose the Rapidly-exploring Random Graph (RRG) as an alternative to the Rapidly-exploring Random Tree (RRT) to finding cyclic trajectories that satisfy a deterministic $\mu$-calculus specification. Another approach is presented in \cite{bhatia2010sampling}, but this time for the syntactically co-safe fragment of Linear Temporal Logic (scLTL). Following these, \citet{vasile2013sampling} propose improvements to \cite{karaman2009sampling}, more specifically for dealing with full LTL and for improving scalability. None of these works, however, deals with partially-known environments, nor do they attempt to speed up the search by learning characteristics of the environment.

More recently, \citet{kantaros2018sampling} described an approach for multi-robot systems under global temporal tasks. Instead of using an RRG, the authors propose a two-step approach using RRT$^\star$. The first step constructs a tree until an accepting state of the automaton capturing the evolution of the LTL formula is reached. The second step then grows another tree rooted at this accepting state, and attempts to find a cyclic (infinite) path that satisfies the LTL specification. The same authors then introduce in \cite{kantaros2020stylus} sampling bias guided by the automaton capturing the LTL, something that \cite{luo2019abstraction} also proposes in a similar fashion. Lastly, besides proposing a heuristic to guide the search, \cite{tajvar2020safe} integrates feedback control laws to guarantee feasibility of plans by robots with complex, possibly non-holonomic, dynamics. Although these works propose ways of improving the time taken to find a plan, they all rely on having details of the environment a priori.

To the best of our knowledge, the two papers that are mostly related to ours are \cite{kantaros2020reactive} and \cite{ayala2013temporal}. The former proposes a reactive sampling-based algorithm for path planning in unknown environments under scLTL specifications. However, differently from what we propose, only the obstacle space is initially unknown to \cite{kantaros2020reactive}, i.e. the locations of the regions of interest, therefore the labeling function, are {known} a priori. On the other hand, \citet{ayala2013temporal} considers completely unknown environments, including the labeling function. However, the authors propose an approach over a discretized partitioning of the environment, performing frontier exploration \cite{yamauchi1997frontier} until a path that satisfies the scLTL specification is found. We merge benefits of both approaches by proposing a sampling-based approach on completely unknown environments; furthermore, we propose a way of learning relations between labels, together with exploiting them for guiding the path search.

When it comes to robotic deployment in unknown environments, a crucial initial step might be to efficiently create a map in an exploratory manner. A seminal work on exploration is by \citet{yamauchi1997frontier}, in which the author proposes the method coined \emph{frontier exploration}. Since then, several other approaches have been proposed. Among them is the Receding Horizon Next-Best-View Planner \cite{bircher2016receding} and the Autonomous Exploration Planner \cite{selin2019efficient}, both building upon RRT$^\star$. These works, however, do not focus on capturing various dependencies and relations in the environment. In  contrast, in a probabilistic approach proposed by \citet{aydemir2013active}, a robot uses common-sense knowledge about the relation between objects and semantic room categories. Here, the focus is however on search for objects and not satisfaction of complex LTL goals. 
%These approaches can be used as fallback cases in our algorithm; i.e. when our approach still does not have enough information to bias the growth of the RRG graph, it need to take a step towards exploring the space in a promising direction.

### 2021-03-03 Paper/2_Prelims.tex
After RSS submission

\section{Preliminaries}\label{sec:prel}

Let $\reals$ denote the set of real numbers and $\reals^n$ the $n$-dimensional Euclidean space. 
We use $\ap$  for the finite set of atomic propositions.
For a set $X$, $2^X$ denotes its power set.
%and $X^\omega$ denotes its countably infinite product $(X\times X\times X\dots)$, also called the set of infinite words over $X$. 
A word over an alphabet $Y$ is a sequence of elements of $Y$.
The exclusive-or operation is denoted by $\oplus$, and the disjoint union of sets by $\uplus$.

Consider a robot deployed in an \emph{environment} $\environment \subset \reals^n$ and let $\initpoint \in \environment$ be its initial state. Let $\{\obstacle_1, \obstacle_2, \dots \obstacle_k\}$ be the set of obstacles such that $\obstacle_i\subset \environment$ for all $i \in [1,k]$, and $\freespace = \environment \setminus \bigcup_{i=1}^{k} \obstacle_i$ denotes the obstacle-free space.
A trajectory in the environment $\environment$ is defined by a continuous function $\traj:~[0,1] \to \environment$. 
A trajectory is collision-free if $\traj(t) \in \freespace$, $\forall t \in [0,1]$. 
Regions of the environment $\environment$ are labelled with atomic propositions $\ap$ according to a labelling function $\labelling : \environment \rightarrow 2^{\ap}$, which maps each state in the state-space to a set of atomic propositions that hold true there.

A map of the environment $\environment$ is a partitioning into a finite number of cells of equal size with a predefined precision, which can be labeled as \emph{free}, meaning that the cell lies in $\freespace$, \emph{occupied}, if any point within the cell lies inside the obstacle space (corresponding to an over-approximation of the obstacle set), or \emph{unmapped}, that highlights the cell has not been seen by the robot so far. Every cell is initialised as \emph{unmapped}, and is updated whenever it lies in the line-of-sight of the robot. A cell is called a \emph{frontier} cell if it is marked as \emph{free} and has a neighbouring cell marked as \emph{unmapped}. A \emph{map frontier} is a connected group of frontier cells, and its size is its cardinality.
This is a common approach among the 3D-exploration community, so we refer to papers such as \cite{yamauchi1997frontier,bircher2016receding} for more details.

\subsection{RRG} 
The Rapidly-exploring Random Graph \cite{karaman2011sampling} is an anytime\footnote{An anytime algorithm returns a valid solution even if it is interrupted before termination; moreover, the longer it runs, the more its solution is improved.} sampling-based motion planning algorithm that builds a connected roadmap. 
It incrementally builds a graph $G = (V,E)$ such that $v \in \freespace$, $\forall v \in V$, and an edge $e \in E$ connects two nodes $v_a, v_b \in V$ if there exists a collision-free trajectory $\traj_{v_a}^{v_b}$ between them, with $\traj_{v_a}^{v_b}(0) = v_a$ and $\traj_{v_a}^{v_b}(1) = v_b$.
A path over $G$ is a sequence of nodes $p = v_0, v_1, v_2, \dots$ such that $v_i \in V$ and $(v_i, v_{i+1}) \in E$, for all $i \ge 0$.

\subsection{Syntactically co-safe LTL and DFA}
\begin{definition}[(Syntactically co-safe) Linear Temporal Logic \cite{ltl,KV99}]
    A formula of \emph{LTL} is given by the syntax:
    \begin{equation*}
        \varphi := a \mid \neg a \mid \varphi_1\wedge\varphi_2 \mid \varphi_1\vee\varphi_2 \mid \X\varphi \mid \varphi_1 \U \varphi_2 \mid \G\varphi
    \end{equation*}
    % 
    where, $a\in \ap$ is an atomic proposition, $\neg, \wedge,\vee$ are the Boolean operators `negation', `conjunction', and `disjunction', respectively. $\X$, $\U$, $\G$ denote the LTL operators `next', `until', and `globally' respectively.
    The \emph{syntactically co-safe fragment of LTL (scLTL)} is given by the same syntax, but prohibiting the operator $\G$.
\end{definition}

The semantics of LTL formulas is defined on words over $\alphabet$. 
The Boolean operators have usual semantics. Intuitively, $\X\varphi$ means that $\varphi$ is true in the next time step and $\varphi_1 \U \varphi_2$ asserts that $\varphi_1$ will be true until $\varphi_2$ becomes true.
$\F$ is known as the `finally' or `eventually' operator whose semantics asserts that the property $\varphi$ becomes true at some point in the future. 
As such, it can be defined in terms of $\U$ as $\F\ \varphi \equiv \true\ \U\ \varphi$
$\G$ is known as the `globally' or `always' operator with the semantics that $\varphi$ is always satisfied.
Since the robot moves in continuous time and $\X$ operator is usually defined for discrete time steps, we consider for simplicity properties without $\X$.
However, our approach is applicable for the whole of LTL.

Let $\lang(\varphi)$ denote the set of words that satisfies the LTL formula $\varphi$.

\begin{definition}[Deterministic Finite Automaton]\label{def:buchi}
    A \emph{deterministic finite automaton (DFA)} is a tuple $(\alphabet,Q,q_0,\delta,F)$ where $\alphabet$ is the alphabet, $Q$ is a finite set of states, $q_0$ is an initial state, $\delta : Q \times \alphabet \rightarrow Q$ is a transition function and $F\subseteq Q$ is the set of accepting states.
\end{definition}

A run over a word $w_1,\ldots,w_n$ is a sequence of states $q_0,q_1,\ldots,q_n$ such that $q_{i} = \delta(q_{i-1},w_i)$ for all $i$. 
A word is accepted by the automaton if the run over the word end in $F$. 
We define the language accepted by an DFA $\ba$ as $\lang(\ba)~=~\{w\in (\alphabet)^\omega \mid\ w$ is accepted by $\ba\}$.
It is a standard result that for every scLTL formula $\varphi$, there exists a DFA $\ba$ such that $\lang(\varphi)~=~\lang(\ba)$ and it is effectively constructible.
Consequently, DFA can be used as a precise representation of an scLTL property.
(Our approach can be extended in a straightforward way to so-called B\"uchi automata, which can express the whole of LTL.)

\begin{definition}[State-labelled transition system]
A (state-labelled) transition system (TS) is a tuple $(S,s_0,\Delta,L)$ where $S$ is a finite set of states, $s_0$ is an initial state, $\Delta: S \to 2^S$ is a transition relation, and $L:S \to \alphabet$ is the labelling function.
\label{def:ts}
\end{definition}

A transition system (representing the real system or its abstraction) can be combined with an automaton (representing the property) into a \emph{product}, see Fig.~\ref{fig:guide}.
Runs of the product are thus runs of the transition system monitored by the automaton.
The automaton always reads the atomic propositions true in the current state and, on the whole, determines whether the run satisfies the property or not.
This standard construction is often used in model checking \cite{BK08} and we use it to improve RRG by  \emph{mutual} exchange of information between the two parts.

\begin{definition}[Product]\label{def:product}
    Given a TS $\lts=(S,s_0,\Delta,L)$ and DFA $\ba=(\alphabet,Q,q_0,\delta,F)$, the product $\lts\times\ba$ is the tuple $(S\times Q,\hat{s_0},\hat{\Delta},\hat F)$ where
    \begin{itemize}
        \item $\hat{s_0}=(s_0,\delta(q_0,s_0))$,
        \item $\hat{\Delta}((s,q))=\{(s',\delta(q,s'))\mid s'\in\Delta(s)\}$,
        \item $\hat F=\{(s,q)\mid q\in F\}$ .
    \end{itemize}
\end{definition}

%\todo{J:this seems reverted to the inconsistent state with BA with transition-acceptance (really needed?) but here both BA and product state-acceptance \\ K: Yeah, we agreed on having state-acceptance in the paper since it is easy to understand and in the implementation section, say that we use transition acceptance and they are equivalent}

% \begin{remark}
% As opposed to non-deterministic B\"uchi automata, they can be used also for model checking probabilistic systems\footnote{Not all LDBA are apt for LTL model checking of Markov decision processes, but those produced by the translation of \cite{ltlandldba} can.} but without complications of other types of acceptance conditions\footnote{For instance, Rabin automata and other types of deterministic automata capable of representing the whole LTL while being usable for model checking probabilistic systems all require to avoid some states eventually.}, constraining the accepting paths only by repeated reachability.
% Consequently, our method is applicable also to systems with stochastic noise and featuring probabilistic behaviour. 
% \end{remark}

### 2021-03-03 Paper/3_ProblemFormulation.tex
After RSS submission

\section{Problem Formulation}\label{sec:prob}
% {\color{blue} In this first paragraph we should say that: i) the obstacle space is not known to the robot a priori; ii) regions of the workspace are labeled with atomic propositions, but that the labeling function is also not known a priori; iii) robot has a sensing radius within which it can identify obstacles and labels.}

% Consider a robot which moves in an unknown \emph{workspace} $\environment \subset \reals^n$ and let $\initpoint \in \environment$ be its starting point. Let $\{\obstacle_1, \obstacle_2, \dots \obstacle_k\}$ be the set of obstacles such that $\obstacle_i\subseteq \environment$ for all $i \in [1,k]$. and $\freespace = \environment \setminus \bigcup_{i=1}^{k} \obstacle_i$, denotes the obstacle-free space.

%Regions of the workspace $\environment$ are labelled with atomic propositions $\ap$ according to a labelling function $\labelling : \environment \rightarrow 2^{\ap}$, which maps each state in the state-space to a set of atomic propositions that hold true there.
Consider a robot  deployed in an \emph{a priori} unknown environment. We assume that the set of atomic propositions $\Sigma$ (semantic labels, such as $\mathit{living\_room}$ or $\mathit{wastebin}$) is known beforehand, but not where they hold. In other words, the $\labelling$ function, as well as the obstacle-space, are unknown. Furthermore, we also assume that the robot is equipped with adequate sensors and perception modules that can identify labels and obstacles within a sensing radius $\sensingradius$ around its current position.
% the labelling function $\labelling_x : \environment_x \rightarrow 2^{\ap}$ and obstacles are known to the robot. Here, $\environment_x$ denotes an $n$-sphere around $x$.

\begin{problem}
\label{problem_main}
Given an initial state $\initpoint\in \environment$ {in an a priori unknown environment $\environment$,} and an scLTL specification $\spec$ over the set of atomic propositions $\ap$, find a collision-free trajectory $\traj$ in $\environment_\mathit{free}$ which satisfies $\spec$.
\end{problem}

Since neither obstacles nor the labeling function are known a priori, one cannot use traditional offline approaches described in Sec.~\ref{sec:related} to solve Problem~\ref{problem_main}. The solution must be an online algorithm that learns the obstacle space and the labeling function as it moves in the environment. A straightforward way to solve this problem would be to explore the whole environment and assign labels to features in the environment first, and then use planning approaches. We propose to integrate exploration and planning. As a result, the robot attempts to make progress towards satisfying the specification while exploring, resulting in a possibly shorter travelled distance.

### 2021-03-03 Paper/4_Solution.tex
After RSS submission

\section{Solution}\label{sec:sol}
Our solution to Problem \ref{problem_main} is an algorithm that learns interesting semantic dependencies and relations in the form of a \emph{semantic abstraction} and utilizes this knowledge to bias the growth of a motion graph towards faster satisfaction of the desired LTL specification.

\subsection{Semantic abstraction-guided RRG}

The overall Semantic abstraction-guided RRG (SAG-RRG) procedure is overviewed in Alg.~\ref{alg:SAGRRG}.
Similarly to RRG, the procedure builds a graph $G=(V,E)$ whose vertices $v \in V$ lay within the obstacle-free space $\freespace$, and edges $e \in E$ connect two vertices if a collision-free trajectory exists. 
An iteration of the algorithm starts by updating the map of the environment with information of what is within the sensing radius $r_s$ of the robot (line \ref{sagrrg:storemap}). 
After that, it computes the guidance according to the \emph{Bias} function (line \ref{sagrrg:getbias}), which is described in more detail in Alg.~\ref{alg:bias} and Sec. \ref{sec:bias}.
Then, each iteration of the internal \emph{while} loop (lines \ref{sagrrg:while_begin}-\ref{sagrrg:while_end}) attempts to add one new vertex to $G$, in a similar way to the RRG algorithm. It samples a point from the known-space of the environment and finds its closest neighbour in the current graph (line \ref{sagrrg:sample}).
If the path connecting these two points is collision-free, the sample is considered for being added to the graph.
If the symbolic counterpart of the sampled transition is in bias, the sampled point is stored as a ``bias frontier";
%\todo{J:this whole sentence is cryptic, needs a bit more explanation; it's our contribution, the rest around is standard RRG; K: Changed a bit but not sure if helpful}
otherwise it rejects this sample with some probability $p$ (lines 11-14). This probability depends on how much you want to bias the sampling.
The algorithm then follows the usual RRG procedure: it adds the new vertex and edge to the graph (line \ref{sagrrg:addedge1}) and attempts to connect such vertex to its closest neighbours (lines \ref{sagrrg:for_begin}-\ref{sagrrg:while_end}), with slight modifications  for checking for bias frontiers, and for keeping track of the symbolic transitions $t_\mathrm{symb}$ (lines 16 and 22) and states seen $\mathrm{seen\_st}$ (line 17).
After sampling a batch, it updates the semantic abstraction through the \emph{Learn} procedure (line \ref{sagrrg:learn}), which is detailed in Alg.~\ref{alg:learn} and in Sec. \ref{sec:learn}.
The algorithm then calls the \emph{Move} procedure (Alg.~\ref{alg:move} and Sec.~\ref{sec:move}), which finds the best frontier to move to, and moves the robot to the point in $G$ closest to it.
Finally, the procedure checks if a plan that satisfies the LTL formula has been found.

\begin{remark}
    In Alg.~\ref{alg:SAGRRG} an edge $e \in E$ is defined in a way to ensure the labels along it change only once. Formally, given an edge $e = (v_a, v_b) \in E$, there exists a state $x \in \sigma_{v_a}^{v_b}$ such that i) $\labelling(x') = \labelling(v_a)$, $\forall x' \in \sigma_{v_a}^{x}$, and ii) $\labelling(x'') = \labelling(v_b)$, $\forall x'' \in \sigma_{x+\epsilon}^{v_b}$, where $x+\epsilon$ represents a state in the neighbourhood of $x$.
    %the labels of $v_a$, i.e. $\labelling(v_a)$, holds true for every state between $v_a$ and $x$ (inclusive), and that $\labelling(v_b)$ hold true from $x$ (exclusive) until $v_b$.
\end{remark}

\begin{algorithm}[t]
    \small
    \DontPrintSemicolon
    \SetKwComment{Comment}{//}{}
    \SetKwFunction{Learn}{Learn}
    \SetKwFunction{Bias}{Bias}
    \SetKwFunction{Move}{Move}
    \SetKwFunction{Sample}{Sample}
    \SetKwFunction{Nearest}{Nearest}
    \SetKwFunction{Near}{Near}
    \SetKwFunction{SampleAndExtend}{SampleAndExtend}
    \SetKwFunction{CollisionFree}{CollisionFree}
    \SetKwFunction{UpdateMap}{UpdateMap}
    \SetKwFunction{AcceptingPath}{AcceptingPath}
    \KwIn{$\environment, \initpoint, \spec$}
    \KwOut{A collision free trajectory in $\environment$ which satisfies $\spec$}
    Initialize semantic abstraction\;
    $\; V \gets \initpoint$; $\; E \gets \emptyset$\;
    $\mathrm{curr\_pos} \gets \initpoint$; $\; \mathrm{seen\_st} \gets s(\initpoint)$\;
	\While{$\neg$\AcceptingPath{}}{
	    \UpdateMap{$\mathrm{curr\_pos}, r_s$} \label{sagrrg:storemap}\;
		$\bias \gets$ \Bias{$\mathrm{seen\_st}$} \label{sagrrg:getbias}\;
		$t_\mathrm{symb} \gets \emptyset$; $\; i\gets 0$\;
		\While{$i < \mathrm{batch\_size}$}{\label{sagrrg:while_begin}
		    $[\sampledpoint, \neighbour] \gets$ \SampleAndExtend{$\freespace, V$} \label{sagrrg:sample}\;
		    \If{\CollisionFree{$\neighbour, \sampledpoint$}}{
		        \If{$(s(\neighbour), s(\sampledpoint)) \in \bias$}{
		            add $\sampledpoint$ to bias frontiers \label{sagrrg:biasfront1}\;
		        }
		        \Else{
		            \textbf{continue} to next iteration with prob $p$ \label{sagrrg:continue}\;
		        }
		        $E\gets E \cup (\neighbour, \sampledpoint)$; $V \gets V \cup \sampledpoint$ \label{sagrrg:addedge1}\;
		        $t_\mathrm{symb} \gets t_\mathrm{symb} \cup (s(\neighbour), s(\sampledpoint))$\;
		        $\mathrm{seen\_st} \gets \mathrm{seen\_st} \cup s(x_s)$\;
		        $i\gets i+1$\;
		    }
		    \For{$x \in$ \Near{$\sampledpoint$}}{\label{sagrrg:for_begin}
		        \If{\CollisionFree{$x, \sampledpoint$}}{
		            $E\gets E \cup (x, \sampledpoint)$; $V \gets V \cup x$ \label{sarrg:addedge}\;
		            $t_\mathrm{symb} \gets t_\mathrm{symb} \cup (s(x), s(\sampledpoint))$\;
    		        \If{$(s(x), s(\sampledpoint)) \in \bias$}{
    		            add $\sampledpoint$ to bias frontiers \label{sagrrg:biasfront2}\;
		            }
		        }
		    }
		}\label{sagrrg:while_end}
		\Learn{$t_\mathrm{symb}$} \label{sagrrg:learn}\;
		$\mathrm{curr\_pos} \gets $ \Move{} \label{sagrrg:move}\;
	}
	\Return{accepting\_path}
    \caption{SAG-RRG}
    \label{alg:SAGRRG}
\end{algorithm}

\subsection{Learn}
\label{sec:learn}
This section describes in detail the proposed approach to learning the semantic abstraction of the environment. Intuitively, we try to find transitions that are similar to the sampled ones and add them as special, potential transitions in the abstraction.
Next part describes how we can accommodate these special transitions in the abstraction.

\subsubsection{Semantic Abstraction}

To formalize the semantic abstraction, we propose extending the state-labelled TS to a ``multi-modal'' transition system, our extension of modal transition systems \cite{LT88}:

\begin{definition}[Multi-Modal Transition System]
A tuple $(S,s_0,\Delta,L,\modes,\modemap)$ is called a multi-modal transition system (MM-TS), where $(S,s_0,\Delta, L)$ is a state-labelled transition system (Def. \ref{def:ts}), $\modes$ is a finite ordered set of modes, and $\modemap: \Delta \rightarrow \modes$ is a modal marking. 
\end{definition}

A semantic abstraction of an RRG graph is an MM-TS, where a discrete state $s \in S$ represents a set of points $x \in \mathcal X$ with the same labeling. 
With a slight abuse of notation, we use $x \in s$ to say that $L(x) = L(s)$ and $s(x)$ to denote $s \in S$, such that $x \in s$.

Intuitively, $\modemap$ assigns to each transition in the abstraction a ``degree'' of confidence that a corresponding transitions is present in the corresponding concrete points.
We use two modes\footnote{Although we choose to use two modes in this paper for the sake of simplicity of the exposition, the approach presented throughout the paper is generic enough to use any number of modes. Besides \emph{must} and \emph{may}, one could also use \emph{may not} and \emph{must not}, for instance.} in our MM-TS: \emph{must} and \emph{may}. The former is used for transitions that are known to exist based on samples taken from the environment while the graph is constructed; the latter is an extrapolation to which transitions might exist based on the \emph{must} transitions.
When a new edge $(x, x_{\mathit{new}})$ is added to the SAG-RRG graph $G$, a transition $(s(x),s(x_{\mathit{new}}))$ is added to the MM-TS as a \emph{must} transition, and similar transitions (see Def.~\ref{def:similar}) are added as \emph{may} transitions.
%Loosely speaking, it is added as \emph{may} if it is similar to some \emph{must} transition that is already in the MM-TS.
% and it is marked as \emph{might} if it has been marked as \emph{may} for some time without turning into \emph{must}. Finally, all other transitions are marked as \emph{unknown}.
%\todo{write that we can add more modes as well, like `may not' and work with them}
%
Let us now formalize when we deem two transitions of a MM-TS \emph{similar}.

\begin{definition}[Domain of Change] 
The domain of change for a transition $(s, s') \in \Delta$ is $\DOC(s,s') = L(s) \oplus L(s')$.
\end{definition}

The \emph{domain of change} is essentially the set of all atomic propositions which changed their valuation during the corresponding transition in the MM-TS. For example, given a transition $(s, s')$ where $L(s)=\{a,b\}$ and $L(s')=\{b,c\}$, its $\DOC(s, s')$ is $\{a,c\}$.

\begin{definition}[Similar Transitions] \label{def:similar}
 Two transitions $(s, s'), (\bar s, \bar s') \in \Delta$ are \emph{similar} if and only if $\DOC(s,s')=\DOC(\bar s, \bar s')$, and $\forall a\in \DOC(s, s')$, $a \in L(s) \iff a \in L(\bar s)$ and $a \in L(s') \iff a \in L(\bar s')$.
\end{definition}

Intuitively, similar transitions behave the same on their \emph{domain of changes}. For example, a transition $(s,s')$, where $L(s)=\{a,b\}$ and $L(s')=\{b,c\}$, is similar to $(\bar s, \bar s')$ where $L(\bar s)=\{a,d\}$ and $L(\bar s')=\{d,c\}$.
The idea is that after experiencing the transition $(s,s')$ which leaves $b$ untouched, we may hypothesize $b$ is irrelevant and that the same behaviour is present also in the situation when $b$ does not hold and when some other irrelevant proposition, e.g.~$d$, holds.
However, $b$ still may be a precondition for the transition, hence we introduce the new transition $(\bar s, \bar s')$ only with a low ``confidence''.

\medskip

The formal definition of similarity allows us to clearly identify when transitions in the MM-TS, i.e. the semantic abstraction of an RRG graph, should be marked as \emph{may}.

\medskip

\subsubsection{Multi-modal product}

The semantic abstraction captures existing and possible dependencies and relationships between labels in the environment regardless of the desired specification. We extend the definition of product (Def. \ref{def:product}) to incorporate the knowledge of the specification and thus enable biasing of SAG-RRG sampling to achieve faster specification satisfaction. In short, a multi-modal product (MM-P) is a product as in Def.~\ref{def:product} but with a MM-TS instead of a TS.

\begin{definition}[Multi-modal Product]
    Given a MM-TS $(S,s_0,\Delta,L,\modes,\modemap)$ and a DFA $\ba=(\alphabet,Q,q_0,\delta,F)$, their product (MM-P) is a tuple $(S\times Q,\hat{s_0},\hat{\Delta},\hat F, \modes, \hat \modemap)$, where the first four components are defined as in Def. \ref{def:product} and the remaining two are the modes $\modes$ and a model marking $\hat \modemap: \hat \Delta \rightarrow \hat \modes$, such that $\hat \modemap((s,q),(s',q')) = \modemap(s, s')$.
\end{definition}

Similarly to the multi-modal transition system, the product can be constructed iteratively, along with the construction of the SAG-RRG graph.

\subsubsection{Learn procedure}
\label{sec:learnproc}

\begin{algorithm}[t]
    \small
    % \KwIn{set of symbolic transitions $t_{symb}$}
    %\KwOut{updated MM-TS and MM-P}
    \DontPrintSemicolon
    \SetKwComment{Comment}{//}{}
    \SetKwProg{Fn}{Function}{:}{}
    \SetKwFunction{Learn}{Learn}
    \SetKwFunction{Add}{AddToProduct}
    \SetKwFunction{Find}{FindSimilar}
    \Fn{\Learn{$t_\mathrm{symb}$}}
    {
        \Add{$t_\mathrm{symb},$ \emph{must}} \label{learn:addmust}\;
        $\similar \gets$ \Find{$t_\mathrm{symb}$} \label{learn:similar}\;
        \Add{$t_{sim},$ \emph{may}}\;
    }
            \label{alg:learn}
            \caption{Learn}
\end{algorithm}

The procedure Learn is summarized in Alg. \ref{alg:learn}.
Given a set of transitions $t_{symb}$, this procedure adds them to the MM-TS as \emph{must} transitions, since we know that these transitions are already there. After that, for each $t\in t_{symb}$, it computes the transitions similar to $t$ and add them as \emph{may} transitions in the MM-TS.

% Given a new edge $e = (x,x_\mathit{new})$ in the SAG-RRG graph $G$, we add to our MM-TS a \emph{must} transition $(s(x),s(x_\mathit{new}))$ (line~\ref{learn:add}). We correspondingly add to our MM-P all states $(s(x_\mathit{new}),q')$ with the property that $q'$ is computed from $\delta(q,L(x_\mathit{new}))$, where $(s(x),q)$ is already a state of the MM-P, and we add the corresponding transitions, too. These transitions are also marked as \emph{must} (line~\ref{learn: must}). The procedure then proceeds to find the set of \emph{similar} transitions to all newly added ones according to Def.~\ref{def:similar} (line~{learn: similar}). These are added to MM-P as \emph{may} transitions (line~\ref{learn:may}). 
% \todo{Lastly, the procedure deals with changing some \emph{may} transitions into \emph{might} (lines~\ref{learn: counter}-\ref{learn: addTran1}). To this end, for each state $s$, we keep track of how many samples have been drawn from $x$, $s=s(x)$; if a transition to $s'$ has been marked as \emph{may}, but we have not witnessed a transition from $x \in s$ to a transition $x' \in s'$ for more than ... steps, we change the marking to \emph{might}, indicating that our marking might have been too optimistic.

% The algorithm keeps track of the number of times a state $(s,q)$ has been added the source of a \emph{must} transitions after a certain threshold $\learnthreshold$, it moves all Level 0 transitions sourced at $s_1$ to the set of Level 1 transitions.}
%Line 10 increases the counter for the source of the transition, which stores how many times has this state has been the source of a sampled transition. If the source has been visited more than the threshold $\learnthreshold$, it converts all possible outgoing transitions from that state which have not been sampled till now to level 1 transitions (lines 11-13).

%Instead of dealing with the abstraction and taking its product with the property automaton every time, we do all the computations on the product automaton itself.

%Our algorithm \emph{learns} possible new transitions (level 2) in the abstraction based on previously sampled transitions (level 3), and uses them to generate \emph{bias} on where new samples should be taken from. Given a sampled transition $t$, the \emph{learn} function adds transitions similar to $t$ (according to Def.~\ref{def:similar}) into the product automaton.

\subsection{Bias}
\label{sec:bias}

\begin{algorithm}[t]
    \small
    \SetKwFunction{Bias}{Bias}
    \DontPrintSemicolon
    \SetKwProg{Fn}{Function}{:}{}
    \SetKwFunction{PreImg}{PreImg}
    \SetKwFunction{PostImg}{PostImg}
    \Fn{\Bias{seen\_st}}{
        $\bias[0] \gets$ transitions ending in accepting states $\mathrm{acc\_st}$\;
        $\mathrm{reached}[0] \gets \mathrm{acc\_st}$\;
        $\mathrm{reached}[1] \gets$ \PreImg{$\mathrm{acc\_st}$}\;
        $\mathrm{all\_reached} \gets \mathrm{reached}[0] \cup \mathrm{reached}[1]$\;
        $i\gets 1$ \;
        \While{\PreImg{$\mathrm{reached}[i]$}$ \nsubseteq \mathrm{all\_reached}$}{
            $\mathrm{useful\_pre} \gets$ \PreImg{$\mathrm{reached}[i]$}$ \cap \,\mathrm{seen\_st}$\;
            $\mathrm{useful\_post} \gets$ \PostImg{$\mathrm{useful\_pre}$}$ \cap \,\mathrm{reached}[i]$\;
            $\bias[i] \gets (\mathrm{useful\_pre}, \mathrm{useful\_post})$\;
            $\mathrm{reached}[i+1] \gets$ \PreImg{$\mathrm{reached}[i]$}\;
            $\mathrm{all\_reached} \gets \mathrm{all\_reached} \cup \mathrm{reached}[i+1]$\;
            $i \gets i + 1$\;
        }
        \Return{$\bias$}
    }
    \caption{Bias}
    \label{alg:bias}
\end{algorithm}

The \emph{bias} procedure computes which transitions would more quickly bring the system to an accepting state of the LDBA. It returns a hierarchical list of transitions according to how far they are from an accepting state in MM-P; the closer a transition is to an accepting state, the better. These transitions can then be used to bias the construction of the motion graph for faster convergence.

The procedure, described in Alg. \ref{alg:bias}, starts by initializing the variables $\bias$ and $\mathrm{reached}$, which store transitions and states, respectively. The first element of $\bias$ is the set of all transitions ending in accepting states of the MM-P (line 2). As for $\mathrm{reached}$, it keeps track of all backwards-reachable states from the accepting states; hence its first element is the set of accepting states (line 3), and the second element is the pre-image of the accepting states (line 4).
Then, until all the backward-reachable sets have been considered, $\bias$ is constructed iteratively based of the set of sampled states $\mathrm{seen\_st}$ (lines 7-13). In the end, $i$th element of $\bias$ will be the set of states that can reach an accepting state after exactly $i$ steps in the MM-P. 
%\todo{J: again, this paragraph is cryptic, needs more explanation or example - it's the new stuff; K:better? Explaining it line by line will make it even more cryptic}

Learn and Bias functions work in unison and help each other improve. The more \emph{may} transitions are learned, the better is the bias received. The better the bias, the more new transitions are learned and the faster it converges.

\subsection{Move procedure} \label{sec:move}

\begin{algorithm}[t]
    \small
    \SetKwFunction{Move}{Move}
    \DontPrintSemicolon
    \SetKwProg{Fn}{Function}{:}{}
    \SetKwFunction{Map}{FindBestMapFrontier}
    \SetKwFunction{Bias}{FindBestBiasFrontier}
    \SetKwFunction{Best}{Best}
    \Fn{\Move}{
        $p_1 \gets$ \Map{}\;
        $p_2 \gets$ \Bias{}\;
        \Return{\Best{$p_1, p_2$}}
    }
    \caption{Move}
    \label{alg:move}
\end{algorithm}

The idea behind the Move procedure, described in Alg.~\ref{alg:move}, is to decide where to move next: should we go towards a place that will provide more information about the map, or should we move according to the advice that has been given by Bias? In order to compare both options, we employ the concept of \emph{information gain} (IG). Given a map frontier, its information gain is defined as $\mathrm{IG_{map}} = \mathrm{size} \times f(d)$, where $\mathrm{size}$ is the size of the frontier and $f(d)$ is a strictly decreasing function (for $d > 0$) of the distance from the robot to the center of the frontier.

%Move procedure (described in Alg. \ref{alg:move}) maintains a grid-like discretization of the environment, which stores if each cell is \emph{free}, contains \emph{obstacle}, or \emph{unknown} yet. 
%We assume that as soon as the robot moves to a new location, it can update the discretization for each visible (path joining the robot and the cell is obstacle free) cell in the sensing radius. 
%A cell is called a \emph{frontier cell} if it is marked \emph{free} and has a neighbouring cell which is marked \emph{unknown}. A (usual) \emph{frontier} is a connected group of frontier cells and its size is the number of cells it has. 
%We define \emph{information gain} (IG) of these frontiers as $\frac{size}{d}$, where $d$ is the distance robot has to travel to reach a point in $G$ closest to the center of the frontier.

In a similar fashion, we define the information gain of a \emph{bias} frontier. Note that \emph{bias} frontiers were introduced in Alg.~\ref{alg:SAGRRG} (lines \ref{sagrrg:biasfront1} and \ref{sagrrg:biasfront2}) as a means to keep track of the vertices in $V$ that correspond to advices given by Alg.~\ref{alg:bias}. Since $\bias$ is a list of transitions, we can associate a \emph{rank} $r$ with each transition from $\bias$ equal to $\mathrm{index}+1$, where $\mathrm{index}$ is the index of the sampled transition. We define IG of these frontiers as $\mathrm{IG_{bias}} = g(r,d)$, where $g$ is some function such that both $g(r, \cdot)$ and $g(\cdot, d)$ are strictly decreasing, where $d$ is again the distance from the robot to the frontier.

The intuition behind the IG of the \emph{map} frontiers is to have a larger value the larger the frontier is, but penalise it according to its distance to the robot, so as to motivate exploration of smaller frontiers that are nearby. Similarly, with the IG of the \emph{bias} frontiers, we want to motivate movement towards low-rank frontiers, since these are closer to satisfying the formula.

### 2021-03-03 Paper/5_Analysis.tex
After RSS submission

\subsection{Analysis}\label{sec:analysis}

%\begin{remark}
%    We assume that there are no prohibited\todo{Jan:I want to polish this (it's not even defined what we mean by prohibited, right? the example illustrates this but we need to clarify this)} regions in the environment since most of the applications require reaching a bunch of regions in some order and repeating this infinitely often. However, even if there were any prohibitions, we can still solve it by treating them like obstacles. 
%\end{remark}
%To illustrate the last remark, we give an example:
%\begin{example}
%    Let's assume there are three regions labelled with $a,b$ and $c$ as shown in fig \ref{}. If the property says that reach $c$ from $a$ while avoiding $b$, our algorithm can sample a point in $c$ and possibly add an edge from $a$ to $c$ directly but that edge will certainly go over $b$. To deal with this, we can simply treat labelled regions as obstacles and when we do collision checking, we also check if the edge crosses any of these regions in between.
%\end{example}

\begin{theorem}
The algorithm is sound, i.e. any trajectory returned by SAG-RRG satisfies the given scLTL formula $\spec$.
\end{theorem}
\begin{proof}
(Sketch) The proposed algorithm iteratively constructs a product MM-P (Def.~\ref{def:product}) between a semantic abstraction of the RRG graph and the automaton $\ba$, which accepts exactly the language of the specification $\spec$. Paths in the product that visit accepting states 
%infinitely many times 
project directly onto accepting runs of the automaton and runs of MM-TS, which in turn project directly onto paths in the RRG graph $G$ and further onto trajectories of the robot in the workspace. Altogether, these trajectories necessarily satisfy~$\spec$.
%Our algorithm attempts to bias the growth of the RRG graph towards faster convergence to an accepting state of the BA $\ba$ (Def.~\ref{def:buchi}); however, in the worst-case scenario, it performs as bad as exhaustive search. Once such an accepting state is reached, the undirected graph generated by the RRG guarantees the existence of a cyclic path that generates an (infinite) accepting word of BA $\ba$.
\end{proof}

\begin{theorem}
SAG-RRG is asymptotically complete.
\end{theorem}
\begin{proof}
(Sketch) Follows directly from the convergence and completeness properties of the original RRG \cite{karaman2011sampling} and the fact that the biasing we introduced allows to eventually sample the whole space. Regardless of the scenario, including the one with no regularity in the environment that can be learned and exploited for guiding the search, the worse-case scenario will see the proposed approach perform an exhaustive search of the environment.%\todo{J:I don't understand this sentence. F: Better?} %By using an exploration algorithm as a fallback to learn the labeling function $\labelling$ as well as the obstacle space, we allow the RRG graph to eventually cover the whole space. 
%Therefore, if there exists a trajectory in the environment that satisfies the LTL specification $\spec$, the proposed algorithm will find it eventually.
\end{proof}

%Since the path that algorithm returns is an accepting path in the product automaton, it is accepted in the property automaton. Hence it is correct. If an accepting path exists, the algorithm will eventually find it because we are generating an RRG graph which will contain all possible paths \todo{Jan: not really but this musy have occurred with standard RRG already; imagine a one dimensional line between two obstacles; also imagine reachability (F x) where x is a single point  - again, classical issue, Sweden probably knows how to comment on this, my view is to require `robust' solutions exist, i.e. after adding \emph{any} $\vec\varepsilon$ with $|\vec\varepsilon|<\delta$ it is still a solution} in the limit and if a correct path exists it will be found.
%\todo{Write that ldba only needs repeated reachability, which is doable in this setting. Jan: we need a lasso but we're never coming back exactly (must have occurred as an issue already with surveillance - GF)}

% The proof of this algorithm can be extended from the proof of correctness of the RRG algorithm. Since the bias is always finite, it can sample from bias for a finite number of times, after that the bias will become null and it will sample randomly.

### 2021-03-03 Paper/6_Implementation.tex
After RSS submission

\section{Implementation and Experiments}\label{sec:cases}
The proposed approach was implemented in Java and run on a consumer grade hardware (2.60GHz Intel~i7-9750H CPU, 32 GB RAM). Binary Decision Diagrams (BDDs), which are very efficient for manipulating sets of Boolean variables, are used for storing and manipulating the product automaton, the labels of each node in the RRG, and the $\bias$. We encode the RRG as an undirected graph whose nodes also store the labels that hold true at that state. We use the Java Spatial Index RTree library for spatial indexing and faster querying of nearest neighbours. JavaBDD and JGraphT are the libraries used for encoding the BDDs and the graph, respectively. We also use \texttt{jhoafparser} library to parse the automaton file.

An example of the office-like environment used for the case study is presented in Fig.~\ref{fig:example}. In order to draw statistically-meaningful results, 100 different instantiations of the environment were randomly generated, in which the footprint (i.e. walls and doors) of the office space remains unchanged, but desks and wastebins are randomly positioned within the rooms (without blocking the door). %This randomization is performed to ensure that the desks and wastebins are not purposely positioned in a way that benefits our approach.
%We ran it on the example showed in Fig. \ref{fig:example}, all the black colored lines are the boundaries and the walls. There are 6 rooms, labelled with atomic propositions $r_1,r_2,\dots r_6$ and the hall (middle part) is labelled with $h$. All the red coloured boxes are tables in each room and green colored boxes are the bins. The area near the tables and bins shown by dashed lines are labelled with $t$ and $b$ respectively.

\input{figures/example}

The scLTL specification is inspired by a realistic scenario, common in every office environment: reach a wastebin in the office rooms. We translate such a specification to the following scLTL formula:
\begin{equation}
    \varphi = \F(r_1 \wedge b) \wedge \F (r_2 \wedge b)\wedge \dots \F (r_6 \wedge b) \label{eq: spec}
\end{equation}
%
Note that such a specification does not impose any ordering of events.

For information gain, we use the following functions in our simulations
% Regarding the calculation of the information gains $\mathrm{IG_{map}}, \mathrm{IG_{bias}}$, given a \emph{map} frontier $m$, a \emph{bias} frontier $x_s$, and the current position $p$ of the robot, we use the following equations in our simulations:
%
\begin{align}
    \mathrm{IG_{map}}(m, p) &= \frac{\mathrm{size}_m}{d_{m, p}} \\
    \mathrm{IG_{bias}}(x_s, p) &= \frac{a}{r^{b}_{x_s} \, d_{x_s, p}}
\end{align} % this was that way so it would occupy less space
%
where $\mathrm{size}_m$ is the size of frontier $m$, $d_{m, p}$ and $d_{x_s, p}$ are the length of the shortest path between $p$ and $m$ and $x_s$, respectively, $a,b > 0$ are  user-defined parameters, and $r_{x_s}$ is the rank of $x_s$. Adjusting $a,b$ is intuitive: i) suppose $m$ and $x_s$ are equidistant from $p$; ii) fix $r_{x_s}$ to 1 and choose $a$ to reflect how a \emph{bias} frontier compares to a \emph{map} frontier; iii) now suppose $x_{s,1}, x_{s,2}$ equidistant from $p$, such that $r_{x_{s,1}}~=~1$ and $r_{x_{s,2}} = 2$; iv) choose $b$ as to reflect how the importance of \emph{bias} frontier decays with its rank (e.g. linearly, quadratic). For our case study, we chose $a=100$ and $b=2$.

\begin{table*}%[th]
    \centering
    \caption{Mean and standard deviation of the total trajectory length to satisfy the scLTL mission \eqref{eq: spec}, along with total runtime and RRG size. The results were drawn from simulating each approach 3 times in each of the 100 randomly-generated environments.}\label{tab:my_label2}
    \setlength{\tabcolsep}{3pt}
    \footnotesize
    \begin{tabular}{l c c c | c c c }
        \toprule
        \textbf{} & \multicolumn{3}{c}{\textbf{See-through Desks}} & \multicolumn{3}{c}{\textbf{Opaque Desks}} \\
        \cmidrule(lr{.75em}){2-4} \cmidrule(l{.75em}r){5-7}
		&\textbf{\sep}&\textbf{\tog}&\textbf{\bia}&\textbf{\sep}&\textbf{\tog}&\textbf{\bia}\\
		\midrule
		\textbf{Total length}     & 77.3 (7.5) & 56.6 (8.0) & 29.4 (5.0) & 79.1 (7.1) & 62.9 (16.5) & 32.3 (11.8) \\
		%%
		\multicolumn{1}{l}{Exploration length}  & 57.1 (3.2) & 37.5 (7.1) & 28.0 (4.9) & 57.8 (4.9) & 44.4 (16.6) & 31.3 (12.1) \\
		%\hline
		\multicolumn{1}{l}{Remaining plan l.} & 20.2 (7.0) & 19.1 (3.6) & 1.3 (1.8) & 21.3 (5.1) & 18.5 (3.4) & 1.1 (1.8)\\
		%%
		\textbf{Total Time}       & 7.8 (2.0) & 6.4 (2.3) & 7.3 (1.9) & 9.6 (2.5) & 8.3 (3.2) & 9.1 (2.4)\\
		%%
		\textbf{RRG size}         & 1931.2 (460.9) & 1938.6 (559.5) & 1793.6 (312.1) & 2313.8 (550.9) & 1868.7 (498.2) & 1901.4 (301.2)\\
		\bottomrule
    \end{tabular}
\end{table*}

\begin{figure*}
    \centering
    \subfloat[\label{fig:tog0}]{\includegraphics[width = 0.33\linewidth]{figures/tog_firstMove.png}}
    \subfloat[\label{fig:tog1}]{\includegraphics[width = 0.33\linewidth]{figures/tog_room.png}}
    \subfloat[\label{fig:tog2}]{\includegraphics[width = 0.33\linewidth]{figures/tog_final.png}} \\
    \subfloat[\label{fig:sep0}]{\includegraphics[width = 0.33\linewidth]{figures/sep_bin.png}}
    \subfloat[\label{fig:sep1}]{\includegraphics[width = 0.33\linewidth]{figures/sep_room.png}}
    \subfloat[\label{fig:sep2}]{\includegraphics[width = 0.33\linewidth]{figures/sep_final.png}}
    \caption{Snapshots of the robot navigating the office environment in the attempt to satisfy the scLTL mission \eqref{eq: spec} with two different approaches. The yellow semi-circle in (a) corresponds to the robot's sensing radius. The top-row figures (a-c) display the trajectory (green) when using the approach proposed by us (SAG-RRG), where exploration and planning are done together; the bottom-row figures (d-f) show the case where the robot first explores the environment, and only then it plans a path that satisfies the mission. The RRG graph at the time of the snapshot is shown in blue, and the path that leads to satisfaction of the mission is in red (c,f).}
    \label{fig:snaps}
\end{figure*}

The results are presented in Table~\ref{tab:my_label2} for 100 randomly-generated environments. 
The solution presented in this paper can be seen as an approach that performs exploration of the environment and planning to satisfy the scLTL mission concurrently, without or with bias in building the RRG tree (`Simultaneous' and `Simult. biased' columns in Table~\ref{tab:my_label2}). We compare this integrated solution to the trivial sequential approach (`Explore, then plan' column in Table~\ref{tab:my_label2}), which consists of first exploring the whole environment, and then planning a trajectory that satisfies the mission. Lastly, in a more technical variant regarding the sensing capabilities of the robot, we analyse two cases, one where the robot can ``see through'' the desks (e.g., a flying robot), and another where they are considered to be Opaque. A few snapshots of the robot trajectory are shown in Fig.~\ref{fig:snaps}. Each approach is run three times in each of the 100 environments, totalling 1800 runs of the experiment.

The rows in Table~\ref{tab:my_label2} display the length (`Total length') of the trajectory traversed by the robot, from its initial state (common to all cases) until mission satisfaction, as well as the total computation time (`Total time') and number of nodes in the RRG graph (`RRG size'). Additionally, we also display the `Exploration length' which, for the sequential approach, represents the length of the trajectory traversed only during the exploration phase, while for our approach it represents the length traversed until the system realises that a trajectory that satisfies the mission already exists. `Remaining plan l.' is the length of the remaining trajectory that needs to be followed in order to satisfy the desired specification at the moment when exploration phase ends in the `Explore, then plan' case, or the moment when the trajectory is found in the `Simultaneous' and ´Simult. biased' cases. In Table~\ref{tab:my_label2} we see that having ``see-through'' desks makes the performance (both total length and time) slightly better in all the cases, which is to be expected as there are not as many occlusions in the map as with ``opaque'' desks. 
We also see that exploration and planning together performs better in general and including the bias makes it more than 2.5 times better than the naive approach in terms of the path length.

In the `Explore, then plan' case, the robot's visits to wastebins during the exploration do not count towards the mission satisfaction, in contrast to the `simultaneous unbiased' case. This is one of the reasons the latter performs better, as expected. The `simultaneous biased' version performs a lot better because it was able to visit a lot of wastebins (with the help of biasing) already during the exploration.

%There are two approaches to solve the problem (i) first do pure exploration then do planning in a known environment, explore and plan together, (ii) trying to satisfy the property while exploring. There can be two variants of both the approaches, with biasing and without biasing. We compare all 4 algorithms on the office like environment described earlier. We ran all 4 variants 100 times and present the average results in table \ref{tab:resultsseethrough} along with the standard deviation.

%Let us start by evaluating the case where the environment is completely known \emph{a priori}. Even though it might seem that the average time taken to find a solution when using the proposed approach is not a large improvement over the unbiased version, we highlight here the difference on the number of iterations taken until a solution was found. In the unbiased version, the algorithm spends a long time expanding the RRG graph randomly, creating a large graph that leads to unnecessary locations in the environment. On the other hand, our algorithm does not grow such a large graph (a third the size of the unbiased version), but invests time learning about the environment and biasing its search towards relevant regions.

%Such a difference becomes even more apparent if the environment is completely unknown \emph{a priori}. Even though both biased and unbiased versions are using the same underlying approach to explore the environment (frontier exploration \cite{yamauchi1997frontier}), our algorithm performs on average eight times faster, with an eighth of the nodes required by the unbiased version to find a satisfying path.

### 2021-03-03 Paper/figures/example.tex
After RSS submission

\begin{figure}
    \centering
    \resizebox{0.75\linewidth}{!}{
    \begin{tikzpicture}

        % Boundary
        \draw[very thick, fill = white] (0, 0) rectangle (8.4, 6);
        \draw[dashed] (0,2) -- (8.4,2) -- (8.4, 4) -- (0,4);

        % Vertical walls
        \filldraw (2.75, 0) rectangle (2.85, 2.05);
        \filldraw (2.75, 3.95) rectangle (2.85, 6);
        \filldraw (5.55, 0) rectangle (5.65, 2.05);
        \filldraw (5.55, 3.95) rectangle (5.65, 6);

        % Horizontal walls
        \filldraw (0, 1.95) rectangle (1.1, 2.05);
    	\filldraw (0, 3.95) rectangle (1.1, 4.05);
        \filldraw (1.7, 1.95) rectangle (3.9, 2.05);
    	\filldraw (1.7, 3.95) rectangle (3.9, 4.05);
        \filldraw (4.5, 1.95) rectangle (6.7, 2.05);
    	\filldraw (4.5, 3.95) rectangle (6.7, 4.05);
        \filldraw (7.3, 1.95) rectangle (8.4, 2.05);
        \filldraw (7.3, 3.95) rectangle (8.4, 4.05);

    	% Table label lines
    	\filldraw[dashed, fill = red!10!white] (0.25,0.25) rectangle (1.55,1.15); % r1
    	\filldraw[dashed, fill = red!10!white] (1.65,4.45) rectangle (2.55,5.75); %r2
    	\filldraw[dashed, fill = red!10!white] (3.1,0.25) rectangle (4,1.55); % r3
    	\filldraw[dashed, fill = red!10!white] (3.1,4.85) rectangle (4.4,5.75); %r4
    	\filldraw[dashed, fill = red!10!white] (7.25,0.25) rectangle (8.15,1.55); % r5
    	\filldraw[dashed, fill = red!10!white] (7.25,4.45) rectangle (8.15,5.75); %r6

    	% Bin label lines
    	\filldraw[dashed, fill = green!10!white] (0,0) rectangle (0.35,0.35); % r1
    	\filldraw[dashed, fill = green!10!white] (2.4,5.65) rectangle (2.75,6); % r2
    	\filldraw[dashed, fill = green!10!white] (2.85,0) rectangle (3.2,0.35); % r3
    	\filldraw[dashed, fill = green!10!white] (2.85,5.65) rectangle (3.2,6); % r4
    	\filldraw[dashed, fill = green!10!white] (5.65,1.6) rectangle (6,1.95); % r5
    	\filldraw[dashed, fill = green!10!white] (8.05,5.65) rectangle (8.4,6); % r6

    	% Bins
    	\filldraw[fill = green!40!white] (0, 0) rectangle (0.2, 0.2) node[midway] {\tiny $b$}; % r1
    	\filldraw[fill = green!40!white] (2.55, 5.8) rectangle (2.75, 6) node[midway] {\tiny $b$}; % r2
    	\filldraw[fill = green!40!white] (2.85, 0) rectangle (3.05, 0.2) node[midway] {\tiny $b$}; % r3
    	\filldraw[fill = green!40!white] (2.85, 5.8) rectangle (3.05, 6) node[midway] {\tiny $b$}; % r4
    	\filldraw[fill = green!40!white] (5.65, 1.75) rectangle (5.85, 1.95) node[midway] {\tiny $b$}; % r5
    	\filldraw[fill = green!40!white] (8.2, 5.8) rectangle (8.4, 6) node[midway] {\tiny $b$}; % r6

    	% Tables
    	\filldraw[fill = red!50!white] (0.4, 0.4) rectangle (1.4, 1) node[midway] {$t$}; % r1
    	\filldraw[fill = red!50!white] (1.8, 4.6) rectangle (2.4, 5.6) node[midway] {$t$}; % r2
    	\filldraw[fill = red!50!white] (3.25, 0.4) rectangle (3.85, 1.4) node[midway] {$t$}; % r3
    	\filldraw[fill = red!50!white] (3.25, 5) rectangle (4.25, 5.6) node[midway] {$t$}; % r4
    	\filldraw[fill = red!50!white] (7.4, 0.4) rectangle (8.0, 1.4) node[midway] {$t$}; % r5
    	\filldraw[fill = red!50!white] (7.4, 4.6) rectangle (8.0, 5.6) node[midway] {$t$}; % r6

    	\filldraw[black] (0.2, 3) circle (0.4mm);
    \end{tikzpicture}
    }
    \caption{Example of an office-like environment used in the case study. Black solid lines represent walls. There are six rooms, each labeled with one atomic proposition $r_i$, for $i \in [1,6]$, and a hallway labeled $h$. Each room contains a table (red) and a bin (green), labeled $t$ and $b$, respectively. The labels of tables and bins hold true within the corresponding dashed and shaded areas surrounding it. The initial position of the robot is marked with a black dot on the left side of the hallway.}
    \label{fig:example}
\end{figure}

### 2021-03-03 Paper/root.tex
After RSS submission

%%%%%%%%%%%%%% IROS %%%%%%%%%%%%%%%%%%
%\documentclass[letter, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
%
%\IEEEoverridecommandlockouts                              % This command is only needed if 
%                                                          % you want to use the \thanks command
%
%\overrideIEEEmargins                                      % Needed to meet printer requirements.

%%%%%%%%%%%%% RSS %%%%%%%%%%%%%%%%%%%%
\documentclass[conference]{IEEEtran}
\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{multicol}
\usepackage[bookmarks=true]{hyperref}

\input{macros}
\input{packages}

% RSS: remove \LARGE \bf
\title{
Semantic Abstraction-Guided Motion Planning \\ for scLTL Missions in Unknown Environments
%LTL Motion Planning Guided by Semantic Abstractions
}

% RSS: 
\author{Author Names Omitted for Anonymous Review.}
% IROS:
%\author{Kush Grover$^{1}$ \and Fernando S. Barbosa$^{2}$ \and Jana Tumova$^2$ \and Jan K\v ret\'insk\'y$^1$% <-this % stops a space
% \thanks{*This work was not supported by any organization}% <-this % stops a space
%\thanks{$^{1}$ Technical University of Munich, Germany. 
%        {\tt\small \{grover, jan.kretinsky\}@in.tum.de}}%
%\thanks{$^{2}$KTH Royal Institute of Technology, Stockholm, Sweden.
%        {\tt\small \{fdsb, tumova\}@kth.se}}%
%}

\begin{document}

% \theoremstyle{theorem}
% \theoremstyle{definition}
% \theoremstyle{problem}

\maketitle

% RSS: remove the following 2 lines
%\thispagestyle{empty}
%\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Complex mission specifications can be often specified through temporal logics, such as Linear Temporal Logic and its syntactically co-safe fragment, scLTL. Finding trajectories that satisfy such specifications becomes hard if the robot is to fulfil the mission in an initially unknown environment, where neither locations of  regions or objects of interest in the environment nor the obstacle space are known a priori. We propose an algorithm that, while exploring the environment, learns important semantic dependencies in the form of a semantic abstraction, and uses it to bias the growth of an RRG graph towards faster mission completion. Our approach leads to finding trajectories that are much shorter than those found by the sequential approach, which first explores and then plans. Simulations comparing our solution to the sequential approach, carried out in 100 randomized office-like environments, show more than 50\% reduction in the trajectory length.
\end{abstract}

% RSS:
\IEEEpeerreviewmaketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{1_Introduction}
\input{2_Prelims}
\input{3_ProblemFormulation}
\input{4_Solution}
\input{5_Analysis}
\input{6_Implementation}

\section{CONCLUSIONS AND FUTURE WORK}\label{sec:conclusion}
We presented an online sampling-based algorithm capable of finding a trajectory in an \emph{a priori} unknown environment that satisfies an scLTL specification. We enrich the RRG algorithm with functions that attempt to learn possible relations between labels of the environment and use such relations for biasing the search for a satisfying trajectory. The resulting paths are significantly shorter than in straightforward sequential exploration followed by planning in a known space.

A few topics to be considered for future work include extending the approach to consider probabilistic relationships in the semantic abstraction of the system, as well as the extension to multi-agent systems.

%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section*{APPENDIX}

%\section*{ACKNOWLEDGMENT}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% IROS:
%\bibliographystyle{IEEEtran}
% RSS:
\bibliographystyle{plainnat}

\bibliography{literature}

\end{document}

### 2021-06-03 Presentation/main.tex
Uploaded Presentation

\documentclass{if-beamer}

\usepackage{pgf}
\usepackage{pgfplots}
\usepackage{amsmath}

\usepgfplotslibrary{fillbetween}

\usepackage{amssymb}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}

%\newcommand{\xmark}{\ding{55}}%
%\newcommand{\cmark}{\ding{51}}
\newcommand{\sep}{Explore, then plan}
\newcommand{\tog}{Simultaneous}
\newcommand{\bia}{Simult.\ biased}

% --------------------------------------------------- %
%                  Presentation info	              %
% --------------------------------------------------- %
%\title[An anytime algorithm for reachability in uncountable MDPs]{An anytime algorithm for reachability in uncountable MDPs}
%\subtitle{dsh}
%\author{Kush Grover}

\title[Robot Motion Planning]{Robot Motion Planning}
%\author{\authorblockN{Kush Grover\authorrefmark{1},
%		Fernando S. Barbosa\authorrefmark{2},
%		Jana Tumova\authorrefmark{2} and
%		Jan K\v ret\'insk\'y\authorrefmark{1}}
%	\authorblockA{\authorrefmark{1}Technical University of Munich, Germany.\\ Emails: \{grover, jan.kretinsky\}@in.tum.de}
%	\authorblockA{\authorrefmark{2}KTH Royal Institute of Technology, Stockholm, Sweden.\\ Emails: \{fdsb, tumova\}@kth.se}
%}

\author{K. Grover, F. Barbosa, J. T\r{u}mov\'a, J. K\v{r}et\'{i}nsk\'{y}}
\logo{
	\includegraphics[scale=0.0065]{figures/tum-logo.png}
}

\institute[]{
	Technical University of Munich, KTH Royal Institute of Technology
}

% \subject{Presentation subject} % metadata

\graphicspath{{figures/}}
% --------------------------------------------------- %
%                    Title + Schedule                 %
% --------------------------------------------------- %

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\arrayrulecolor{tableColor1}
% --------------------------------------------------- %
%                      Presentation                   %
% --------------------------------------------------- %

\section{Motion Planning Problem}

\begin{frame}
	\frametitle{Motion Planning Problem}
	\begin{center}
		\begin{tikzpicture}
		\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
		\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
		\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
		\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
		\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
		\filldraw[black] (0.2, 3) circle (0.4mm);
		\end{tikzpicture}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{RRT/RRG}
	\begin{center}
		\begin{tikzpicture}
		\onslide<0->{
			\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
			\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
			\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
			\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
			\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
			\filldraw[black] (0.2, 3) circle (0.4mm);
		}
		\onslide<1-2>{
			\filldraw[red] (1.8, 3.8) circle (0.4mm);
		}
		\onslide<2->{
			\filldraw[black] (0.6, 3.2) circle (0.4mm);
			\draw (0.2,3) -- (0.6, 3.2);
		}
		\onslide<3->{
			\filldraw[black] (0.45, 2.6) circle (0.4mm);
			\draw (0.2,3) -- (0.45, 2.6);
			\filldraw[black] (0.8, 3.5) circle (0.4mm);
			\draw (0.6,3.2) -- (0.8, 3.5);
			\filldraw[black] (0.9, 2.9) circle (0.4mm);
			\draw (0.6,3.2) -- (0.9, 2.9);
		}
		\onslide<4>{
			\filldraw[red] (0.8, 2.5) circle (0.4mm);
		}
		%			\onslide<5->{
		%				\filldraw[black] (0.8, 2.5) circle (0.4mm);
		%				\draw (0.45, 2.6) -- (0.8, 2.5);
		%			}
		%			\onslide<6->{
		%				\draw (0.9, 2.9) -- (0.8, 2.5);
		%			}
		\end{tikzpicture}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{RRT}
	\begin{center}
		\begin{tikzpicture}
		\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
		\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
		\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
		\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
		\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
		\filldraw[black] (0.2, 3) circle (0.4mm);
		\filldraw[black] (0.6, 3.2) circle (0.4mm);
		\draw (0.2,3) -- (0.6, 3.2);
		\filldraw[black] (0.45, 2.6) circle (0.4mm);
		\draw (0.2,3) -- (0.45, 2.6);
		\filldraw[black] (0.8, 3.5) circle (0.4mm);
		\draw (0.6,3.2) -- (0.8, 3.5);
		\filldraw[black] (0.9, 2.9) circle (0.4mm);
		\draw (0.6,3.2) -- (0.9, 2.9);
		\filldraw[black] (0.8, 2.5) circle (0.4mm);
		\draw (0.45, 2.6) -- (0.8, 2.5);
		\end{tikzpicture}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{RRG}
	\begin{center}
		\begin{tikzpicture}
		\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
		\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
		\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
		\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
		\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
		\filldraw[black] (0.2, 3) circle (0.4mm);
		\filldraw[black] (0.6, 3.2) circle (0.4mm);
		\draw (0.2,3) -- (0.6, 3.2);
		\filldraw[black] (0.45, 2.6) circle (0.4mm);
		\draw (0.2,3) -- (0.45, 2.6);
		\filldraw[black] (0.8, 3.5) circle (0.4mm);
		\draw (0.6,3.2) -- (0.8, 3.5);
		\filldraw[black] (0.9, 2.9) circle (0.4mm);
		\draw (0.6,3.2) -- (0.9, 2.9);
		\filldraw[black] (0.8, 2.5) circle (0.4mm);
		\draw (0.45, 2.6) -- (0.8, 2.5);
		\draw (0.9, 2.9) -- (0.8, 2.5);
		\end{tikzpicture}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{What about LTL?}
	\begin{itemize}
		\item Algorithms already exists for LTL specifications.\\
		\pause
		\item But they only work if the robot knows the whole environment.\\
		\pause
		\item In reality, the robot know only about things inside a sensing radius and it cannot see beyond walls or obstacles.
		\pause
		\item To know the whole environment, the robot has to explore before start planning. This is usually done by a ``frontier-based" exploration.\\
		%			\pause
	\end{itemize}
	%		\vspace{10pt}
	%		\textbf{Can we do exploration and planning together?}
\end{frame}

%\begin{frame}
%	\frametitle{Frontier Exploration}
%	\begin{itemize}
%		\item Discretize the environment into a grid. \pause
%		\item Mark each cell of the grid as \textit{free}, \textit{obstacle} or \textit{unknown}. Initially all cells are \textit{unknown}. \pause
%		\item An \textit{unknown} cell is a `frontier cell' if one of its neighbours is free. \pause
%		\item A chain of frontier cells is called a \textit{frontier}. \pause
%		\item \textit{Information Gain} of a frontier depends on its size and distance from the current position of the robot.
%	\end{itemize}
%\end{frame}

\begin{frame}
	\frametitle{LTL Motion Planning}
	\begin{center}
		\begin{tikzpicture}
		\filldraw[dashed, fill = gray!10!white] (0.25,0.25) -- (1.55,0.25) -- (1.55,1.15) -- (0.25,1.15) -- (0.25, 0.25);
		\filldraw[dashed, fill = green!10!white] (0,0) -- (0.35,0) -- (0.35,0.35) -- (0,0.35) -- (0,0);
		\filldraw[dashed, fill = gray!10!white] (1.85,4.45) -- (2.75,4.45) -- (2.75,5.75) -- (1.85,5.75) -- (1.85, 4.45);
		\filldraw[dashed, fill = green!10!white] (2.65,5.65) -- (3,5.65) -- (3,6) -- (2.65,6) -- (2.65,5.65);

		\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
		\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
		\filldraw[fill = green!40!white, draw = black] (2.8, 5.8) rectangle (3, 6);
		\filldraw[fill = black!10!white, draw = black] (2, 4.6) rectangle (2.6, 5.6);

		\draw[orange] (0,1.7) arc (-95:95:1.3);

		\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
		\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
		\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
		\filldraw[black] (0.2, 3) circle (0.4mm);
		\end{tikzpicture}
%		\vspace{5pt}
		\\ \emph{Specification:} $F (r_1 \wedge b) \wedge F (r_2 \wedge b)$
	\end{center}
\end{frame}

\section{Naive Algorithm}

\begin{frame}
	\frametitle{Naive Solution}
	\begin{itemize}
		\item Explore the whole environment initially using frontier based approaches. \pause
		\item Start building the RRG graph and construct an abstraction of the system on-the-fly. \pause
		\item Construct the product automaton with the property automaton. \pause
		\item Find an accepting path in the product. \pause
		\item Lift this path to the original environment.
	\end{itemize}
\end{frame}

%	\begin{frame}
%		\frametitle{Better Solution}
%		Can we do better?\\ \pause \vspace{10pt}
%		\textbf{Jan's idea:} Get help in sampling from current abstraction. \\
%		Depending on previous experiences. In other words, predict which samples would be better from previous samples. \\ \pause \vspace{10pt}
%		For e.g: The bin was close to the table in first room so try to look close to the table in the other room to find the bin.
%	\end{frame}

\begin{frame}
	\frametitle{Can we do better?}
	\begin{itemize}
		\item Too much effort is wasted in exploration. \pause
		\item Can we do exploration while trying to satisfy the path simultaneously? \pause
		\item Identify places to go to, so that there is some progress in satisfying the property.
	\end{itemize}
\end{frame}

\section{Better Solution}

\begin{frame}
	\frametitle{Our solution}
	\begin{itemize}
		\item Sample a batch in the known area (initially, sensing radius) using biasing (initially, no biasing).
		\only<2-3>{
			\item Learn from the newly added edges and improve the bias.
		}
		\only<4>{
			\item \textcolor{red}{Learn from the newly added edges and improve the bias.}
		}
		\onslide<3->{
			\item Find the best move and go there.
			\vspace{10pt}
		}
		\onslide<2>{
			\begin{figure}
				\centering
				\scalebox{.5}{
					\begin{tikzpicture}[->,>=stealth', minimum width = 32mm,text width = 30mm, state/.style={
						rectangle,
						%   rounded corners,
						draw=black, very thick,
						minimum height=2em,
						inner sep=2pt,
						text centered,
					},]
					\node[state=initial, color = blue, anchor=center, fill = blue!10!white,thin] (A) {System};
					\node[state, color = blue, below of = A, fill = blue!10!white, node distance =1.5cm,thin] (B) {RRG graph};
					\node[state, color = blue, below of = B, fill = blue!10!white, node distance =1.5cm,ultra thick] (C) {Semantic abstraction};
					\node[state, color = red, below of = C, fill = red!10!white, xshift = -2cm, node distance =1.5cm,thin] (D) {Product};
					\node[state, color = green!50!black, left of = A, fill = green!10!white, node distance =4cm,thin] (E) {scLTL property};
					\node[state, color = green!50!black, below of = E, fill = green!10!white, node distance =2.25cm,thin] (F) {Automaton};

					\path (A) edge[] (B)
					(B) edge[ultra thick] node[right,pos=0.5]{\texttt{Learn}} (C)
					(C) edge (D)
					(E) edge (F)
					(F) edge (D)
					(D.east) edge[bend right=70,looseness=1.5, ultra thick] node[right,pos=0.9]{\texttt{Bias}} (B.east)
					%guidance $\rightarrow$ better paths and learning w.r.t. property
					;
					\end{tikzpicture}
				}
				%					\caption{Scheme of our model-checking-inspired approach with novel elements drawn thickly. }
				\label{fig:guide}
			\end{figure}
		}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Learn and Bias}
	\begin{itemize}
		\item Learn from past experiences i.e. look for transitions that are similar to transitions that we have seen till now. \pause
		\item For each transition sampled in the current batch, add similar `maybe' transitions in the abstraction as well. \pause
		\item Use these `maybe' transitions to bias the search.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Learn and Bias}
	\begin{itemize}
		\item What does similar mean here?\\
		$(r_1, t) \rightarrow (r_1, b) \implies (r_2, t) \rightarrow (r_2,b)$. \\
		\begin{itemize}
			\item Compute domain of changes: Set of APs changing in a transition.\\
			$DOC = \{t,b\}$, $(s_1 \oplus s_2)$
			\item Add transitions $s_1' \rightarrow s_2'$ where $s_1'$ and $s_2'$ are states which agree with $s_1$ and $s_2$ on $DOC$ respectively.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Example}
	\begin{center}
		\begin{tikzpicture}[scale = 0.9]
		\filldraw[dashed, fill = gray!10!white] (0.25,0.25) -- (1.55,0.25) -- (1.55,1.15) -- (0.25,1.15) -- (0.25, 0.25);
		\filldraw[dashed, fill = green!10!white] (0,0) -- (0.35,0) -- (0.35,0.35) -- (0,0.35) -- (0,0);
		\filldraw[dashed, fill = gray!10!white] (1.85,4.45) -- (2.75,4.45) -- (2.75,5.75) -- (1.85,5.75) -- (1.85, 4.45);
		\filldraw[dashed, fill = green!10!white] (2.65,5.65) -- (3,5.65) -- (3,6) -- (2.65,6) -- (2.65,5.65);

		\filldraw[fill = green!40!white, draw = black] (0, 0) rectangle (0.2, 0.2);
		\filldraw[fill = black!10!white, draw = black] (0.4, 0.4) rectangle (1.4, 1);
		\filldraw[fill = green!40!white, draw = black] (2.8, 5.8) rectangle (3, 6);
		\filldraw[fill = black!10!white, draw = black] (2, 4.6) rectangle (2.6, 5.6);

		\draw (0, 0) -- (3, 0) -- (3, 6) -- (0, 6) -- (0, 0);
		\draw (0, 2) -- (1.2, 2) (1.8, 2) -- (3, 2);
		\draw (0, 4) -- (1.2, 4) (1.8, 4) -- (3, 4);
		\filldraw[black] (0.2, 3) circle (0.4mm);

		%			\onslide<2->{
		%				\node[state,initial,initial text=,scale=0.8] (q1) at (5,5) {$0$};
		%				\node[state,scale=0.8] (q2) at (8,5) {$1$};
		%				
		%				\draw (q1) edge[loop above] node{$r_1 \wedge r_2 \wedge b\ \ast$} (q1);
		%				\draw (q1) edge[loop below] node{$(r_1 \wedge \neg b) \vee \neg r_1$} (q1);
		%				\draw (q1) edge[bend left, above, ->] node{$r_1 \wedge \neg r_2 \wedge b$} (q2);
		%				\draw (q2) edge[loop above] node{$(r_2 \wedge b) \vee \neg r_1$} (q2);
		%				\draw (q2) edge[bend left, below, ->] node{$r_2 \wedge b\ \ast$} (q1);
		%			}

		\onslide<2->{
			\filldraw[black] (0.6, 3.2) circle (0.4mm);
			\draw (0.2,3) -- (0.6, 3.2);
			\filldraw[black] (0.45, 2.6) circle (0.4mm);
			\draw (0.2,3) -- (0.45, 2.6);
			\filldraw[black] (0.8, 3.5) circle (0.4mm);
			\draw (0.6,3.2) -- (0.8, 3.5);
			\filldraw[black] (0.9, 2.9) circle (0.4mm);
			\draw (0.6,3.2) -- (0.9, 2.9);
			\filldraw[black] (0.8, 2.5) circle (0.4mm);
			\draw (0.45, 2.6) -- (0.8, 2.5);
			\draw (0.9, 2.9) -- (0.8, 2.5);
			\filldraw[black] (1.1, 3.8) circle (0.4mm);
			\draw (0.8, 3.5) -- (1.1, 3.8);
			\filldraw[black] (1.4, 3.9) circle (0.4mm);
			\draw (1.1, 3.8) -- (1.4, 3.9);

			\node[state,initial,initial text=] (h) at (5,3) {$h$};
			%				\draw (h) edge[loop above, color = green, very thick] node{} (h);
		}

		\onslide<3->{
			\filldraw[black] (1.3, 4.3) circle (0.4mm);
			\draw (1.4, 3.9) -- (1.3, 4.3);

			\node[state] (r1) at (7,4) {$r_1$};
			\draw (h) edge[color = green, bend left, ->, very thick] node{} (r1);
		}
		\onslide<4->{
			\filldraw[black] (1.9, 5.7) circle (0.4mm);
			\draw (1.3, 4.3) -- (1.9, 5.7);

			\node[state] (r1t) at (9, 4) {$r_1, t$};
			\draw (r1) edge[color = green, bend left, ->, very thick] node{} (r1t);

		}
		\onslide<6->{
			\node[state] (r2) at (7,2) {$r_2$};
			\node[state] (r2t) at (9, 2) {$r_2, t$};
			\draw (r2) edge[color = blue, bend left, ->] node{} (r2t);

		}
		\onslide<7->{

			\filldraw[black] (2.9, 5.9) circle (0.4mm);
			\draw (1.9, 5.7) -- (2.9, 5.9);
			\filldraw[black] (1.5, 1.9) circle (0.4mm);
			\draw (0.8, 2.5) -- (1.5, 1.9);

			\node[state] (r1b) at (11, 4) {$r_1, b$};
			\node[state] (r2b) at (11, 2) {$r_2, b$};
			\draw (r1) edge[color = green, bend left, ->, very thick] node{} (h);
			\draw (r1t) edge[color = green, bend left, ->, very thick] node{} (r1);
			\draw (r1t) edge[color = green, bend left, ->, very thick] node{} (r1b);
			\draw (r1b) edge[color = green, bend left, ->, very thick] node{} (r1t);
			\draw (h) edge[color = green, ->, very thick] node{} (r2);

			%				\draw (r2) edge[color = blue, bend left, ->] node{} (h);
			\draw (r2t) edge[color = blue, bend left, ->] node{} (r2);
			\draw (r2t) edge[color = blue, bend left, ->] node{} (r2b);
			\draw (r2b) edge[color = blue, bend left, ->] node{} (r2t);
		}
		\onslide<8->{
			\draw (r2) edge[color = red, bend left, ->, thick] node{} (r2t);
			\draw (r2t) edge[color = red, bend left, ->, thick] node{} (r2b);

		}

		\end{tikzpicture}
		\onslide<5-6>{
			$DOC = \{t\}$
		}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Our solution}
	\begin{itemize}
		\item Sample a batch in the known area (initially, sensing radius) using biasing (initially, no biasing).
		\item Learn from the newly added edges and improve the bias. \textcolor{green}{\cmark}
		\only<1>{
			\item Find the best move and go there.
		}
		\only<2>{
			\item \textcolor{red}{Find the best move and go there.}
		}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Where to move?}
	Possible options:
	\begin{itemize}
		\item One of the frontiers (as in frontier exploration).
		\item Points sampled in the current batch which were according to the bias. Define IG for these based on how far they are from the accepting state in the product automaton.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Our solution}
	\begin{itemize}
		\item Sample a batch in the known area (initially, sensing radius) using biasing (initially, no biasing).
		\item Learn from the newly added edges and improve the bias. \textcolor{green}{\cmark}
		\item Find the best move and go there. \textcolor{green}{\cmark}
	\end{itemize}
\end{frame}

\section{Experiments}

\begin{frame}
	\frametitle{A run of ``First Explore Then Plan"}
	\only<1>{
		\includegraphics[width=\textwidth]{Fig/sep_bin.png}
	}
	\only<2>{
		\includegraphics[width=\textwidth]{Fig/sep_room.png}
	}
	\only<3>{
		\includegraphics[width=\textwidth]{Fig/sep_final.png}
	}
\end{frame}

\begin{frame}
	\frametitle{A run of ``Simultaneously with biasing"}
	\only<1>{
		\includegraphics[width=\textwidth]{Fig/tog_firstMove.png}
	}
	\only<2>{
		\includegraphics[width=\textwidth]{Fig/tog_bin.png}
	}
	\only<3>{
		\includegraphics[width=\textwidth]{Fig/tog_room.png}
	}
	\only<4>{
		\includegraphics[width=\textwidth]{Fig/tog_final.png}
	}
\end{frame}

\begin{frame}
	\frametitle{Experiments: See Through Desks}
	Compared different approaches on 100 randomly generated office like environments.
	\vspace{15pt}
	\begin{table}%[th]
		\centering
		\scalebox{0.8}{
			\begin{tabular}{l c c c}
				\hline
				\textbf{} & \multicolumn{3}{c}{\textbf{See-through Desks}}  \\
				&\textbf{\sep}&\textbf{\tog}&\textbf{\bia}\\
				\hline
				\textbf{Total length}     & 77.3 (7.5) & 56.6 (8.0) & 29.4 (5.0) \\
				%%
				\multicolumn{1}{l}{Exploration length}  & 57.1 (3.2) & 37.5 (7.1) & 28.0 (4.9)  \\
				%\hline
				\multicolumn{1}{l}{Remaining plan l.} & 20.2 (7.0) & 19.1 (3.6) & 1.3 (1.8)\\
				%%
				\textbf{Total Time}       & 7.8 (2.0) & 6.4 (2.3) & 7.3 (1.9) \\
				%%
				\textbf{RRG size}         & 1931.2 (460.9) & 1938.6 (559.5) & 1793.6 (312.1) \\
				\hline
		\end{tabular}}
	\end{table}
\end{frame}

\begin{frame}
	\frametitle{Experiments: Opaque Desks}
	Compared different approaches on 100 randomly generated office like environments.
	\vspace{15pt}
	\begin{table}%[th]
		\centering
		\scalebox{0.8}{
			\begin{tabular}{l c c c}
				\hline
				\textbf{} & \multicolumn{3}{c}{\textbf{Opaque Desks}} \\
				&\textbf{\sep}&\textbf{\tog}&\textbf{\bia}\\
				\hline
				\textbf{Total length}    & 79.1 (7.1) & 62.9 (16.5) & 32.3 (11.8) \\
				%%
				\multicolumn{1}{l}{Exploration length} & 57.8 (4.9) & 44.4 (16.6) & 31.3 (12.1) \\
				%\hline
				\multicolumn{1}{l}{Remaining plan l.} & 21.3 (5.1) & 18.5 (3.4) & 1.1 (1.8)\\
				%%
				\textbf{Total Time}       & 9.6 (2.5) & 8.3 (3.2) & 9.1 (2.4)\\
				%%
				\textbf{RRG size}         & 2313.8 (550.9) & 1868.7 (498.2) & 1901.4 (301.2)\\
				\hline
			\end{tabular}
		}
	\end{table}
\end{frame}

\section{Conclusion and Future Work}

\begin{frame}
	\frametitle{Conclusion and Future Work}

	\textbf{Conclusion}
	\begin{itemize}
		\item Gave an algorithm for Motion Planning of scLTL missions in an unknown environment. \pause
		\item Use the previous experiences to help in the future. \pause
		\item Reduce the path traversed by more than 50\% compared to the naive approach.
	\end{itemize}
	\vspace{20pt}

	\textbf{What's next?}
	\begin{itemize}
		\item Can we do full LTL?. \pause
		\item Experiments on actual robots. \pause
		\item Figure out atomic propositions on-the-fly to make it work in a completely unknown environment.
	\end{itemize}
	\vspace{20pt}
	Thank You!
\end{frame}

\end{document}
